### 2014-10-19 Rmarkdown/avy.Rmd
Pre-break commit

---
title: "Identifying epialleles in isogenic mice"
author: "Peter Hickey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

# Chapter overview

The agouti viable yellow ($A^{vy}$) mouse model is the archetypical _epiallele_, a genetically identical region where epigenetic variability result in phenotypic variation. The research group of Professor Emma Whitelaw, one of the researchers to discover the epigenetic variant that underpins the $A^{vy}$ model, recently performed whole-genome bisulfite-sequencing of five mice, including two with the $A^{vy}$ locus, with the aim of identifying additional epialleles.

Due to a somewhat unusual experimental design, Professor Whitelaw requested the assistance of Professor Terry Speed and I in analysing the data. As with many experiments, there was pressure to quickly develop methods and implement these as software to analyse the data. This chapter describes the experiment and how we addressed these challenges. It also discusses general considerations for identifying epialleles.

# Introduction + literature review

"The Agouti gene is responsible for determining whether a mammal's coat is banded (_agouti_) or of a solid color (_non-agouti_)" [TODO: re-word. Wikipedia!]. In mice, there are several well-studied mutants where differences in the expression of the agouti protein influences multiple, seemingly unrelated phenotypic traits (_pleiotropy_). Specifically, yellow mice, those with high expression of the agouti protein, are obese, diabetic and have an increased susceptibility to tumours.

One of these agouti-associated obesity mutations is the agouti viable yellow allele ($A^{vy}$), discovered by (__TODO: Either 1960s [according to http://books.google.com.au/books?id=V6wNB2cGxMwC&lpg=PT250&dq=IAP%20near%20avy&pg=PT250#v=onepage&q=IAP%20near%20avy&f=false] or \cite{Duhl:1994bx}). The cause of the $A^{vy}$ mutation in the agouti protein was identified by a team led by Professor Emma Whitelaw \cite{Morgan:1999ds} as being due to methylation differences upstream of the agouti gene. Moreover, these epigenetic-driven phenotypic differences display variable expressivity (differences between individuals), variegation (mosaicism between cells of the same individual), are mitically heritable and display parent-of-origin effects (distinct from classic parental imprinting).

For the $A^{vy}$ allele, there is an intracisternal A particle (IAP) insertion  approximately 100 kb upstream of the agouti gene. The IAP contains a cryptic promoter that can induce ectopic (unusual) expression of the agouti gene \cite{Rakyan:2002wz}. Methylation of this IAP prevents activation of the cryptic promoter and so the agouti gene is inactive, ultimately resulting in a mouse with a _pseudoagouti[^pseudoagouti_vs_agouti]_ coat. Conversely, when the IAP is unmethylated the mouse has a yellow coat since the cryptic promoter is activated and the agouti protein is expressed. Mice with intermediate levels methylation[^intermediate_methylation] at the IAP have a mottled coat due to mosaicism between cells.

[^pseudoagouti_vs_agouti]: These mice are properly described as pseudoagouti since they are heterozygous for the wildtype agouti allele ($A^{vy}/a$) but are phenotypically indistinguishable from true agouti mice, which are homozygous for the wildtype gene ($a/a$).

[^intermediate_methylation]: Intermediate methylation referring to the average methylation of the IAP across a pool of cells from the individual mouse.

The $A^{vy}$ allele is homozygous lethal, only heterozygous mice ($A^{vy}/a$) and homozygous wildtype mice ($a/a$) are viable, and so the mouse model is properly referred to as the agouti viable yellow mouse (__TODO: Raykan (2012) refer to homozygous $A^{vy}/A^{vy}$ mice, but I thought this was embryonically lethal???__).

An epiallele is a region of the genome that is genetically identical yet epigenetically variable. Epialleles are of most interest when they also confer phenotypic variabilitiy. The $A^{vy}$ allele is the archetype of a _metastable_ epiallele, which \cite{Rakyan:2002wz} define as an allele "at which the epigenetic state can switch and establishment is a probabilistic event". Importantly, such metastable epialleles are mitotically heritable (__TODO: Are there non-metastable epialleles?__).

The $A^{vy}$ epiallele was discoverable because much was known about its probable location and it could be verified through functional assays applied to the offspring of extensive mouse breedings (__TODO: Clarify__). The agouti gene was known to cause differences in coat colour and so it was the obvious candidate. Differences in expression of this gene were not due to genetic variability since the mice were isogenic, and environmental differences were minimal since all mice were raised in a similar laboratory environment. This pointed to the possibility that an epiallele was responsible for the phenotypic variability. I know describe a general strategy for identifying epialleles when there is a strong candidate gene. This is not exactly the strategy employed by \cite{Morgan:1999ds} but is a

## A general strategy for identifying epialleles of known genes

Suppose we have a gene whose expression is known to cause the phenotypic variation we are interested in. Furthermore, suppose that we suspect this difference in gene expression is due to epigenetic variation, such as when the variation in gene expression is observed in genetically identical animals raised in similar environments. These days we could perform bisulfite-sequencing of several samples with different values of the phenotype and look for methylation differences that associate with the phenotype.

The obvious first place to look is at the promoter of the gene, to look for the classical methylation/expression relationship described in Chapter \ref{chap:intro}. Things are rarely so simple and we must often widen our search. Cryptic promoters, such as that in the IAP insertion of the $A^{vy}$ locus, are good candidates (__TODO: How easy is it to identify cryptic promoters__). Enhancers, which can be many kilobases upstream or downstream of a gene, are more challenging to identify since the search must be further widened.

While challenging, this is the "easiest" experimental design from which to identify epialleles since we have a known phenotype and a candidate region(s) in which to search. But what if we want to perform a screen for _putative_ epialleles? A putative epiallele is any region of the genome where the samples display differential methylation. It is identified based solely on differential methylation and not on any association with phenotypic differences. Rather, these putative epialleles may be followed-up in larger cohorts to see whether they associate with a yet-to-be-determined phenotype. This is the type of experiment that Emma Whitelaw came to Terry and I with to ask for assistance in analysing.

# Description of experiment

Five mice on a C57Bl/6 background underwent whole-genome bisulfite-sequencing. Two of the five mice are heterozygous for the $A^{vy}$ locus, one yellow and one pseudoagouti, while the remaining mice are homozygous wildtype agouti (Table \ref{tab:whitelaw_data}).

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
Sample code & Mouse number & Agouti genotype           \\ \midrule
sample1     & mouse \#56   & $A^{vy}/a$ (yellow)       \\
sample2     & mouse \#57   & $a/a$                     \\
sample3     & mouse \#58   & $a/a$                     \\
sample4     & mouse \#65   & $A^{vy}/a$ (pseudoagouti) \\
sample5     & mouse \#66   &                           \\ \bottomrule
\end{tabular}
\caption{Description of mice in Whitelaw experiment.}
\label{tab:whitelaw_data}
\end{table}

The aim of the study is identify putative epialleles, regions that differ in methylation between individuals but without any genetic cause (the mice are inbred and isogenic[^isogenic], so we assume methylation is not affected by genetic variants).

[^isogenic]: Technically, the mice are isogenic except for a 3 Mb region surrounding the $A^{vy}$ allele. The $A^{vy}$ allele is inherited from a distant ancestor who was the offspring of mating between a C57Bl/6 mouse ($a$) with a C3H mouse ($A^{vy}$). It is inherited as part of the 3 Mb chunk of DNA from the C3H genome.

While any differential methylation identified in this analysis is by definition a candidate epiallele, certain features are make stronger candidates. The Whitelaw are particularly interested in differential methylation of retrotransposons. Two of the best examples of metastable epialleles, $A^{vy}$ and $Axin^{Fu}$, are due to associations with retroviral elements \cite{Rakyan:2002wz}. To be of interest, the epiallele should also have a functional consequence, but this cannot be determined from this experiment alone. Rather, this will be determined by downstream analyses incorporating additional data and experiments of the putative epialleles.

My involvement in the project was at a relatively late stage and focused on developing a method to analyse these data. I was to write a software implemention of this method but the final analysis would be performed by Harald Oey, a bioinformatician in the Whitelaw group. Consequently, this chapter focuses on the statistical details of the analysis rather than the biological results and their interpretation.

## Data

Harald aligned the reads to the mouse reference genome (mm9) using Bismark \cite{Krueger:2011eb}. __TODO: Check pre- and post-processing of reads.__ I received from Harald five tables, one for each sample, containing the number of methylated and unmethylated reads at each CpG. An example is shown below:

```
chr     start   end     beta    M       U
chr2    3000163 3000164 0.7     7       3
chr2    3000214 3000215 0.916666666666667       11      1
chr2    3000221 3000222 1       12      0
```

I only received the data for chromosome 2, since these were sufficient for me to develop the methodology and its implementation.

The mouse reference genome is based on the C57Bl/6 strain and does not contain the $A^{vy}$ allele. Therefore we cannot measure methylation at the IAP insertion in the $A^{vy}$ allele from the sequencing data. However, because the $A^{vy}$ allele is inserted in chromosome 2, there may be some "bleedover" into the region surrounding the insertion (chr2:154,760,839-154,760,969) (__TODO: This is far shorter than 3 Mb, why?__)

__TODO: Comment on sequencing coverage__

# Methodology

The basic question is to identify loci where not all of the five mice have the same methylation level. There is no grouping structure of the samples. While we could conceivably group the mice by coat colour (yellow vs. agouti and pseudoagouti) or by genotype at the agouti locus ($a/a$ vs. $A^{vy} vs. a$) or by _epigenotype_ at the agouti locus ("yellow" vs. pseudoagouti vs. agouti), the Whitelaw group were interested in _any_ methylation differences between _any_ subset of the five mice[^n=1]. This effectively makes this a one group experiment.

[^n=1]: Both the "coat colour" and "agouti epigenotype" contrasts result in groups with $n = 1$, meaning we would be unable to estimate their corresponding within-group variances.

This is the first example I had seen of a bisulfite-sequencing experiment with one-group. Surprisingly, this rules out all(?) published methods for identifying differential methylation, since these are all designed for identifying differential methylation in experiments with two or more groups.

There is also the question of what sized "features" we are interested in testing for differential methylation. Harald and Emma told me that they were interested in both differentially methylated CpGs and differentially methylated regions, where the regions were not defined _a priori_. I took the commonly used approach of testing individual CpGs for differential methylation and then building runs of DMCs to define differentially methylated regions.

My algorithm for calling differentially methylated regions is adapted from that of \cite{Hansen:2011gu}. I call my implementation `runDMC` (runs of differentially methylated cytosines) and it proceeds as follows:

1. Identify candidate differentially methylated cytosines (_cDMCs_).
2. Declare contiguous runs of _cDMCs_ where the $IPD$ between _cDMCs_ is less than some threshold to be candidate differentially methylated regions (_cDMRs_).
3. Summarise _cDMRs_ and apply filters.

## Step 1: Identifying candidate differentially methylated cytosines

At each CpG, $i$, we are testing the following hypothesis:

\begin{equation*}

H_{0}:& \beta_{i, j} = \beta_{i, 0}, j = 1, \ldots, 5 \\
  vs.& \\
H_{1}:& \exists j = 1, \ldots, 5 s.t. \beta_{i, j} \neq \beta_{i, 0}

\end{equation*}

where $\beta_{i, 0}$ is estimated from the data. For each CpG we can summarise the data by a $5 \times 2$ contingency table:

\begin{table}[h]
\centering
\begin{tabular}{lll}
\hline
Sample code & Methylated reads & Unmethylated reads \\ \hline
sample1     & $m_{i, 1}$       & $u_{i, 1}$         \\
sample2     & $m_{i, 2}$       & $u_{i, 2}$         \\
sample3     & $m_{i, 3}$       & $u_{i, 3}$         \\
sample4     & $m_{i, 4}$       & $u_{i, 4}$         \\
sample5     & $m_{i, 5}$       & $u_{i, 5}$         \\ \hline
\end{tabular}
\caption{A $5 \times 2$ contingency table summarising the number of methylated and unmethylated reads for each sample, $j$, at the $i^{th}$ CpG.}
\label{tab:5x2}
\end{table}

This amounts to the classical test of differences in proportions for a group of samples. I used the simple $\chi^{2}$ test, as implemented by the `prop.test` function in `R` (__CITE__). This compares the observed and expected counts of methylated and unmethylated reads under the assumption that all 5 samples have equal methylation levels at this CpG. In fact, I wrote a simple wrapper around `prop.test`, called `propTest` (__TODO: Include code in appendix__), so that only the the value of the $\chi^{2}$ test statistic, the degrees of freedom and the resulting P-value were retained, since the rest of the object returned by `prop.test` was not required. `propTest` can also parallelise the computation by chromosome for a faster runtime.

__TODO: Plot and comment on the distribution of P-values__

A CpG is declared a _cDMC_ if the P-value is in the lower $5\%$ of all P-values across the genome. A stricter criteria is required if the aim of the analysis is to identify differentially methylated cytosines. However, this first step is rather designed to identify sites for the `runDMC` algorithm (described below).

The question of

## Identifying differentially methylated regions

The `runDMC` algorithm is similar to one proposed in \cite{Hansen:2011gu}, which is implemented in the `bsseq` R/Bioconductor package (__CITE__). `bsseq` defines differentially methylated region as runs of CpGs with a t-statistic greater than an empirically chosen threshold. Because `bsseq` is designed for two-group experiment, it also requires that the t-tatistic has the same sign for all CpGs in the region, indicating that the effect is in the same direction for all CpGs in the DMR.

Rather than using the $\chi^{2}$ test-statistics, `runDMC` uses the corresponding P-values. The reason is that for some CpGs only a subset of the samples have sequencing data, hence the test statistic has fewer degrees of freedom. By using the P-value rather than the raw $\chi^{2}$ test statistic, we effectively normalise for the degrees of freedom.

`runDMC` is written in `C++` and interfaces with `R` via the `Rcpp` package (__CITE__). The `C++` version is some $25$ times faster than the equivalent code written in `R` (__TODO: Include benchmark data__). The user passes it an object containing the set of CpGs, each annotated by whether it is a _cDMC_, and specifies __UP TO HERE__

# Results

# Discussion

# Conclusion