### 2014-11-15 Rmarkdown/a_statistical_model_of_methlyC-seq.Rmd
Moved stat model chapter to Rmarkdown.

---
title: "A statistical framework for analysing bisulfite-sequencing data"
author: "Peter Hickey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

\section{Notes from outline.md (plus other notes added along the way)}

* Chapter Overview
* Notation
  * See `notation.Rmd`
* Estimating $Z$, $M$, $U$ and $\beta$
  * See `notation.Rmd`
  * Current approaches
* Statistical properties of $z$, $m$, $u$ and $\beta$
  * See `notation.Rmd`
  * Marginal distributions
  * Dependence/correlation over $i$ (see also __Chapter X__)
    * Li _et al._ compute $\beta$ values and compute correlations within a sample as a function of genomic distance.
    * Arabidopsis compute $\beta$ values and compute correlations within a sample as a function of genomic distance.
    * Lister (2009) Fig. 3 and Sup. Fig. 9
    * Lister _et al._ compute both within-read and across-read correlations or dependencies of DNA methylation.
    * Akulenko, R., _et al._ compute gene-level $\beta$ values and compute the correlations between pairs of genes across samples as a function of genomic distance.
  * Dependence/correlation over $j$ (?)
    * Akulenko, R., _et al._ compute gene-level $\beta$ values and compute the correlations between pairs of genes across samples as a function of genomic distance.

\section{Chapter overview}\label{sec:statmodel_chapter_overview}

- [x] DONE

In this chapter I set out a statistical framework for analysing bisulfite sequencing data. I begin by explaining the various levels of stochasticity in a methylC-seq experiment. Following this, I define the mathematical notation that I use throughout my thesis. I describe key variables, common estimators of these and their statistical properties, and statistical models used in downstream analyses.

To begin, I consider the simple experiment of performing methylC-seq on a single sample. Once I have described the framework for a single sample, I extend it to multiple samples. The ideas here are simple, although when extended to their full generality the notation becomes messy.

\section{One sample}\label{sec:one_sample}

- [x] DONE

Even with only a single sample, there are several levels of variation to consider, which have a hierarchical structure. I separate this structure into two stages, as illustrated in __CARTOON FIGURE OF METHYLC-SEQ EXPERIMENT__:

1. Pre-sequencing
2. Post-sequencing

\subsection{Single locus analysis}\label{sec:one_sample_single_locus}

- [x] DONE

As described in Chapter \ref{chap:wgbs_analysis}, most analyses to date of WGBS data have focused on the methylation level at individual loci averaged across all cells in the sample. For this reason, I initially describe the framework in terms of single loci. However, bisulfite sequencing data contain much more information than provided by these univariate, marginal summaries of the data. In Section \ref{sec:one_sample_mtuples} I introduce _m-tuples_ that can summarise some of this extra information and which I use extensively in Chapter \ref{chap:comethylation}.

\subsubsection{Pre-sequencing}\label{sec:one_sample_pre_sequencing}

- [x] DONE

A methylation locus is a single cytosine, that is, a CpG, CHG or CHH. The set of these loci is labelled $\mathcal{I} = \{pos_{i}: i = 1, \ldots, N_{loci} \}$, where $pos_{i}$ is the genomic co-ordinates of the $i^{th}$ locus with respect to the forward strand, e.g. chr1:723,461. I frequently refer to loci by the subscript $i$ rather than by $pos_{i}$. This means that the distance between the $i^{th}$ and $(i + 1)^{th}$ methylation loci varies along the genome and, for a small number of instances, that the $i^{th}$ and $(i + 1)^{th}$ methylation loci are on separate chromosomes. Generally, $N_{loci}$ is not known exactly, although estimates may be made based from the sample or a reference genome, but this is no great concern.

The methylation state of a locus can vary within a sample due to the fact that DNA for each sample is extracted from hundreds or thousands of cells and each cell can have a slightly different methylation profile. Furthermore, within a diploid cell there are two copies of each chromosome, and therefore two copies of each methylation locus, and these two copies can have different methylation states. Therefore, it is also necessary to consider the next level down in the hierarchy; the DNA fragments within the sample.

Suppose that in the pool of DNA fragments for the sample that there are $H_{i}$ fragments containing the $i^{th}$ methylation locus. In general, $H_{i}$ is unknown and will vary from locus to locus within a sample[^H_i]. Note that the value of $H_{i}$ is determined following the library preparation, including PCR amplificatin of the DNA; therefore, it can give a grossly distorted picture of the true representation of the cells. I denote by $\mathcal{H}_{i}$ the set of all fragments containing the $i^{th}$ locus.

[^H_i]: Knowing $H_{i}$ would require knowing: (1) the number of cells used as input (which might only be known to within an order of magnitude), (2) the ploidy of each cell (generally known) and (3) the number of PCR cycles (generally known). But the real problem is that none of the steps in creating the pool of DNA fragments is perfect, in particular, PCR introduces biases -- some molecules are preferrentially amplified while others "drop out". So even if we knew (1), (2) and (3) we cannot simply multiply these together to compute $H_{i}$, although this might at least give us a rough estimate.

Although we do not know the number of fragments in the pool, we can define (and measure) the methylation state of a locus on a single DNA fragment. I denote by the indicator random variable, $Z_{h, i}$, the methylation state of $i^{th}$ methylation locus on the $h^{th}$ DNA fragment containing the $i^{th}$ methylation locus:

\begin{equation*}
Z_{h, i} = \left\{
  \begin{array}{l l}
    1 & \quad \text{if methylated on the } h^{th} \text{ fragment}\\
    0 & \quad \text{if unmethylated on the } h^{th} \text{ fragment}
  \end{array} \right.
\end{equation*}

By summing over the number of fragments containing the $i^{th}$ locus, we obtain the number of fragments that are methylated at the $i^{th}$ locus ($M_{i}$) and unmethylated at the $i^{th}$ locus ($U_{i}$):

\begin{align*}
  M_{i} &= \sum_{h = 1}^{H = H_{i}} Z_{h, i} = |\{Z: Z \in \mathcal{H}_{i}, Z = 1 \}| \\
  U_{i} &= \sum_{h = 1}^{H = H_{i}} (1 - Z_{h, i}) = |\{Z: Z \in \mathcal{H}_{i}, Z = 0 \}|
\end{align*}

Related to these is the proportion of fragments that methylated at the $i^{th}$ locus:
\begin{equation*}
  B_{i} = \frac{M_{i}}{M_{i} + U_{i}}
\end{equation*}

Again, I emphasise that $H_{i, j}, Z_{h, i, j}, M_{i, j}, U_{i, j} \text{ and } B_{i, j}$ are unobservable. However, by sequencing the pools of DNA fragments we aim to estimate these variables.

\subsubsection{Post-sequencing}\label{sec:one_sample_post_sequencing}

- [x] DONE

We do not sequence every fragment in the pool. Rather, sequencing can be thought of as sampling without replacement from the pool of DNA fragments. We have a large number(~$10^{10}$) of fragments in the pool and each methylation locus is only present on a small number of those fragments. Therefore, we can approximate this sampling by Poisson sampling, where the rate parameter for locus $i$ is proportional to the number of fragments in the pool and inversely proportional to $H_{i}$.

At this point I ignore reads containing no methylation loci as these are not relevant to this discussion and I make three simplifying assumptions:

1. We perform single-end sequencing.
2. Sequencing is performed without error.
3. Read mapping is perfect.

The effect of the first assumption is minor. When using single-end sequencing, the methylation loci from a single read will always form a positively ordered sequence without "gaps", i.e., $(i, i + 1, i + 2)$ and not $(i, i - 1, i - 2)$ nor $(i, i + 1, i + 3)$. However, when using paired-end sequencing the methylation loci from a paired-end read will still be an ordered sequence but one of the following may occur (__FIGURE__):

1. There may be gaps due to the insert size being longer than the sum of the read lengths, e.g. $(i, i + 1, i + 3, i + 4))$. In effect, we have missing data for any intervening methylation loci; the $(i + 2)^{th}$ loci in this example.
2. Loci may be measured twice if the insert size is less than the sum of the read lengths, e.g. read_1 gives us $(i, i + 1)$ and read_2 gives us $(i + 1, i + 2, i + 3)$. In this example we must use only one of read_1 or read_2 as the measurement of the $(i + 1)^{th}$ locus because otherwise we are 'double-counting'.

I discuss the implications of the latter two assumptions in Section \ref{sec:sequencing_and_read_mapping_errors}.

Each read measures the methylation state of one or more loci from a single DNA fragment. I denote by $\mathcal{R}_i$ the set of all reads containing the $i^{th}$ locus. Therefore, the number of reads containing the $i^{th}$ locus is $|\mathcal{R}_{i}|$, where $|\mathcal{R}_{i}| \leq H_{i}$ with strict inequality for almost all $i$.

A single read containing the $i^{th}$ locus is denoted by $z: z \in \mathcal{R}_{i}$; the observed methylation state is given by:

\begin{equation*}
z: z \in \mathcal{R}_{i} = \left\{
  \begin{array}{l l}
    1 & \quad \text{if methylated at the } i^{th} \text{ locus}\\
    0 & \quad \text{if unmethylated at the } i^{th} \text{ locus}
  \end{array} \right.
\end{equation*}

By summing over the number of reads containing the $i^{th}$ locus we obtain the number of reads that are methylated at the $i^{th}$ locus ($m_{i}$) and unmethylated at the $i^{th}$ locus ($u_{i}$):

\begin{align*}
  m_{i} &= \sum_{z: z \in \mathcal{R}_{i}} z \\
        &= |\{z: z \in \mathcal{R}_{i}, z = 1 \}| \\
  u_{i} &= \sum_{z: z \in \mathcal{R}_{i}} (1 - z) \\
      &= |\{z: z \in \mathcal{R}_{i}, z = 0 \}
\end{align*}

Similarly, we obtain the proportion of reads that are methylated at the $i^{th}$ locus as:
\begin{equation*}
  \beta_{i} = \frac{m_{i}}{m_{i} + u_{i}}
\end{equation*}

These are the so-called $\beta$-values, which are commonly interpreted as an estimate of the proportion of cells in the sample that are methylated at the $i^{th}$ locus. I discuss this interpretation, and other estimators of the  methylation level at a locus, in __SECTION__. I now move from single locus summaries to multi-loci summaries using m-tuples.

\subsection{m-tuples: Multiple loci analysis}\label{sec:one_sample_mtuples}

I define an m-tuple to be a tuple of methylation loci, where m = $1, 2, \ldots$ is the size of the tuple. For example, a CpG 3-tuple is 3 CpGs. A CHH 1-tuple is a single CHH; 1-tuples are the basis of single locus analyses of bisulfite-sequencing data. We can learn more from bisulfite-sequencing data by using m-tuples with m $> 1$.

An observation on an m-tuple is the methylation pattern from a single read that overlaps the m-tuple. I require that the observation is from a single read as this ensures that each observation ultimately comes from a single cell[^chimeric_reads]. For example, suppose we have a read containing 3 CpGs -- the first two CpGs are methylated while the last one is unmethylated. From this read we can observe 3 $\times$ CpG 1-tuples, 2 $\times$ CpG 2-tuples and 1 $\times$ CpG 3-tuple. We can have multiple observations on an m-tuple by observing multiple reads, each containing the m-tuple. __FIGURE__ illustrates this example.

[^chimeric_reads]: This ignores chimeric reads, which depending on the sequencing protocol may be a problem.

Note that in __FIGURE__ I have not constructed a 2-tuple using the first CpG and the last CpG. In general, I focus on m-tuples where the m methylation loci are adjacent. That is, I focus on m-tuples where the _number of intervening loci_ is zero ($NIL = 0$). There are 3 reasons for this:

1. Quantity: From a sequence containing $l$ methylation loci there are $l - \text{m} + 1$ m-tuples, provided that we restrict ourselves to those m-tuples with $NIL = 0$. In contrast, if we allow $NIL \geq 0$ then there are $\binom{l}{m} \sim$ m-tuples.
2. Interpretability: Results for m-tuples with $NIL = 0$ are simpler to interpret than when allowing $NIL \geq 0$. This is discussed in Chapter \ref{chap:co-methylation}
3. Measurability: We cannot observe m-tuples where the methylation loci are far apart due to the read length limitations of the Illumina sequencing technology. This is true even when $NIL = 0$ but is especially the case if we allow $NIL \geq 0$.

Generally, when referring to m-tuples I implicitly mean $NIL = 0$; I will explicitly use the notation $NIL \geq 0$ when I wish to make clear that there may be intervening methylation loci in the m-tuple.

For each m-tuple, I define the _intra-pair distance_ (IPD) as vector containing the $m - 1$ pair-wise distances (measured in bp) between adjacent methylation loci in the m-tuple. For example, the 2-tuple (chr7:145, 163) has $IPD = (163 - 145) = (18)$. The 5-tuple (chr2:560, 570, 572, 588, 612) has $IPD = (570 - 560, 572 - 570, 588 - 572, 612 - 588) = (10, 2, 16, 24)$. The IPD vector of a 1-tuple is not defined. In Chapter \ref{chap:co-methylation} I use the IPDs, along with other features such as the genomic context, to define 'similar' m-tuples.

\subsubsection{Pre-sequencing}

- [x] DONE

Mathematically, an m-tuple is denoted by a sequence of methylation loci, $(i, i + 1, \ldots, i + m - 1)$. The remaining definitions are analogous to those for single methylation loci, that is, when m $= 1$.

I denote by the vector of indicator random variable, $Z_{h, (i, i + 1, \ldots, i + \text{m} - 1)}$, the methylation pattern of the m-tuple $(i, i + 1, \ldots, i + m - 1)$ on the $h^{th}$ DNA fragment containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$:

\begin{equation*}
Z_{h, i} = \left\{
  \begin{array}{l l}
    (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus on the } h^{th} \text{ fragment}\\
    (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th}) \text{ locus and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus on the } h^{th} \text{ fragment} \\
    \vdots \\
    (1, 1, \ldots, 1) & \quad \text{if methylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus on the } h^{th} \text{ fragment}
   \end{array} \right.
\end{equation*}

$\mathcal{H}_{(i, i + 1, i + m - 1)}$ denotes the set of all fragments containing the m-tuple $(i, i + 1, i + m - 1)$ and $\mathcal{R}_{(i, i + 1, i + m - 1)}$ denotes the set of all reads containing the m-tuple $(i, i + 1, i + m - 1)$.

There are $2^{\text{m}}$ possible methylation patterns at an m-tuple. Rather than denote a methylation pattern by a m-vector of zeros and ones, I also write these using $U$ and $M$; for example, the possible methylation patterns at a 2-tuple are $MM, MU, UM$ and $UU$.

Analogously to the definition of $M_{i}$ and $U_{i}$ and $m_{i}$ and $u_{i}$ for the case where $\text{m} = 1$, we have when $\text{m} = 2$:

\begin{align*}
  MM_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (1, 1)\}| \\
  MU_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (1, 0)\}| \\
  UM_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (0, 1)\}| \\
  UU_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (0, 0)\}| \\
\end{align*}

We could extend the $B_{i}$ values to m-tuples, although the intuitive interpretation of these as the average methylation level is lost. Instead, it reflects the relative frequencies of each methylation pattern. Here are the definitions for $\text{m} = 2$:

\begin{align*}
  B_{(i, i + 1)}^{MM} &= \frac{MM_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}} \\
  B_{(i, i + 1)}^{MU} &= \frac{MU_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}}  \\
  B_{(i, i + 1)}^{UM} &= \frac{UM_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}}  \\
  B_{(i, i + 1)}^{UU} &= \frac{UU_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}}
\end{align*}

The definitions for $\text{m} > 3$ follow in the obvious manner.

Again, I emphasise that $H_{(i, i + 1, i + \text{m} - 1), j}, Z_{h, (i, i + 1, i + \text{m} - 1), j}, B_{(i, i + 1, i + \text{m} - 1), j}$ and the set of methylation patterns are unobservable. However, by sequencing the pools of DNA fragments we aim to estimate these variables.

\subsubsection{Post-sequencing}

- [x] DONE

The set of all reads containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$ is denoted by $\mathcal{R}_i$. A single read containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$ is denoted by $z: z \in \mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)}$, and the observed methylation state is given by:

\begin{equation*}
z: z \in \mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)} = \left\{
  \begin{array}{l l}
    (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus}\\
    (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th}) \text{ locus and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus} \\
    \vdots \\
    (1, 1, \ldots, 1) & \quad \text{if methylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus}
      \end{array} \right.
\end{equation*}

Note that we do not know from which haplotype ($h$) each read came from, only that all methylation loci in the read came from the same DNA fragment.

By counting over the number of reads containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$ we obtain the number of reads contain each methylation pattern at that m-tuple. Here are the definitions for $\text{m} = 2$:

\begin{align*}
  mm_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (1, 1)\}| \\
  mu_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (1, 0)\}| \\
  um_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (0, 1)\}| \\
  uu_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (0, 0)\}|
\end{align*}

The definitions for $\text{m} > 2$ follow in the obvious manner.

We can extend the $\beta_{i}$ values to m-tuples. Here are the definitions for $\text{m} = 2$:

\begin{align*}
  \beta_{(i, i + 1)}^{mm} &= \frac{mm_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}} \\
  \beta_{(i, i + 1)}^{mu} &= \frac{MU_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}} \\
  \beta_{(i, i + 1)}^{um} &= \frac{UM_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}} \\
  \beta_{(i, i + 1)}^{uu} &= \frac{UU_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}}
\end{align*}

The definitions for $\text{m} > 3$ follow in the obvious manner.

Finally, \cite{Landan:2012kp} compute the average methylation of a read containing the m-tuple $(i, \ldots, i + m - 1)$. For each read, $z \in \mathcal{R}_{(i, i + 1, i + 2, i + 3)}$, the average methylation of the read, $\zeta_{z}$, is defined as the proportion of methylation loci in the read that are methylated. Thus, $\zeta_{z} = 0, \frac{1}{m}, \frac{2}{m}, \ldots, 1$.

\subsection{Some complications for a single sample}\label{sec:one_sample_complications}

- [x] DONE

I now discuss some complications and how this framework might accommodate these issues in practice.

\subsubsection{What is $\mathcal{I}$?}\label{sec:what_is_I}

- [x] DONE

As mentioned in Chapter \ref{chap:wgbs_analysis}, studies using bisulfite-conversion assays rely on either a reference genome or, less commonly, separate DNA sequencing of the sample that is assayed. Different analysis strategies lead to different definitions of $\mathcal{I}$, which are approximations to the 'true' set of methylation loci in the sample, $\mathcal{I}^{truth}$. Listed here are definitions of $\mathcal{I}$ from least closely matching to most closely matching $\mathcal{I}^{truth}$:

1. $\mathcal{I}^{ref}$: Defined by the set of methylation loci in the reference genome. This ignore all genetic variation between the sample and the reference.
2. $\mathcal{I}^{refFilter}$: Defined by filtering out problematic loci from $\mathcal{I}^{ref}$. A conservative approach that removes many sites of genetic variation between the sample and the reference as well as sites that do not display genetic variation between the sample and the reference. Misses sample-specific methylation loci.
3. $\mathcal{I}^{BisSNP}$: Defined by calling genetic variants from the bisulfite-sequencing data using `Bis-SNP` \cite{Liu:2012ge}. Identifies sample-specific methylation loci and removes reference-specific methylation loci. The best that can be done if only bisulfite-sequencing data is available. Genetic variation detection is not as good as that from DNA-seq data. This is the approach I favour.
4. $\mathcal{I}^{WGS}$: Defined by identifying all methylation loci from _whole-genome sequencing_ (WGS) of the sample's genome. The gold standard. All methylation loci are defined with respect to the sample's genome. The only differences between $\mathcal{I}^{WGS}$ and the "truth" are due to sequencing errors and variant calling errors of the WGS data.

\subsubsection{Genetic heterozygosity at a methylation locus}\label{sec:heterozygosity}

- [x] DONE

In a diploid cell there are are sites where, for example, the maternal chromosome is a CpG whereas the paternal chromosome is an ApG. In effect, the maternal and paternal chromosomes within the sample have different $\mathcal{I}^{truth}$. The number of these heterozygous methylation loci is often small enough not to worry about. However, in some studies, such as those of allele-specific methylation, these loci can be very important.

Parent-specific methylation loci can be identified by calling heterozygous genetic variants using `Bis-SNP` or from WGS of the sample. In practice, the existance of parent-specific methylation loci is often ignored.

\subsubsection{Sequencing and read mapping errors}\label{sec:sequencing_and_read_mapping_errors}

- [x] DONE

Sequencing errors from the Illumina technology are typically base substitutions, for example, a _G_ being read as a _T_. These errors occur more frequently at the 3' end of the reads. Other sequencing technologies have different error profiles. In DNA sequencing, the base quality scores are used to filter out or down-weight likely sequencing errors. However, because the base quality scoring algorithm is tuned for DNA-seq where all 4 nucleotides appear frequently, the base quality scores of methylC-seq data may not be as reliable since the relative frequencies of the 4 nucleotides is quite different.

There are two negative effects of sequencing errors. Firstly, if a methylated cytosine on the forward strand, which should be a C in the read, is erroneously read as another base -- A, G or T -- then the methylation signal is lost. If the erroneous call is an A or a G then at least we might be able to infer that a sequencing error occured. However, if the erroneous basecall is a T then we might instead mistakenly conclude that the cytosine was unmethylated.

Secondly, sequencing errors make read mapping more difficult. These issues are discussed in Chapter \ref{chap:wgbs_analysis}. In terms of the statistical framework described in this chapter, the effect of both sequencing errors and read mapping errors are to introduce errors into $m$ and $u$ and all resulting statistics.

\section{Multiple samples}\label{sec:multiple_samples}

- [x] DONE

To move from a single sample to multiple samples simply requires an additional subscript, $j = 1, \ldots, n$, where $n$ is the number of samples. For example, $\mathcal{I}_{j}$ is the set of methylation loci in the $j^{th}$ sample and $\beta_{i, j}$ is the beta value for the $i^{th}$ locus in the $j^{th}$ sample.

This defines the three levels in the hierarchy of a typical experiment -- individual molecules ($h$), individual methylation loci ($i$) and individual samples ($j$). A fourth level is how the samples relate in terms of an outcome of interest, such as phenotype. This fourth level might be defined up-front, such as in a designed experiment looking for differences in methylation between pre-defined groups of samples, or the aim of the experiment might be to discover this level, such as in a clustering analysis.

A common experiment of the first type is the two-group design in which $n_{1}$ samples are from group $1$ and $n_{2}$ samples are from group $2$ ($n_{1} + n_{2} = n$). This can be represented by a design matrix $X = [X_{j}]$, where $X_{j} = 1$ if the sample is from group $1$ and $X_{j} = 0$ if the sample is from group $2$.

More generally, we can study multi-group experiments via a suitable definition of the design matrix $X$. We can also include covariates in the standard way by allowing $X_{j}$ to be a row vector, $X_{j} = (x_{1, j}, \ldots, x_{P, j})$, where $x_{p, j}$ encodes the information on the $p^{th}$ covariate for the $j^{th}$ sample.

The second type of experiment is one that aims to cluster individuals based on methylation patterns. This can be thought of as an experiment where the outcome of interest is unknown and we wish to discover it by clustering together samples with 'similar' methylation patterns. We might also wish to then correlate these clusters with some external information, such as disease-status.

\subsection{Some complications with multiple samples}

- [x] DONE

In addition to the complications of the previous section, we now have sample-to-sample variation that must be addressed by this framework.

\subsubsection{}$\mathcal{I}$}

- [x] DONE

Each individual has their own set of methylation loci, that is, $\mathcal{I}_{j}$ differs for all $j$. Furthermore, sequencing coverage varies from sample-to-sample. This means that even if the samples have exactly the same $\mathcal{I}_{j}$, i.e., the samples are genetically identical, each sample will have a different set of loci with 'sufficient' sequencing coverage. Loci without sufficient sequencing coverage are effectively missing data.

In practice, we might choose to study $\mathcal{I}^{common} = \cap_{j} \mathcal{I}_{j}$ or some other intersection of the $\mathcal{I}_{j}$, such as all methylation loci present in at least some fraction of the $n$ samples.

A conservative analysis might only analyse those loci where at least some fraction of the $n$ samples have sufficient sequencing coverage. A less conservative analysis might try to impute the missing values based on methylation levels at neighbouring loci.

\section{Parameter estimation}\label{sec:parameter_estimation}

- [x] DONE

In this section I summarise current techniques for parameter estimation from WGBS data. I do not describe the processing of the raw data. When necessary, I have 'translated' the original work into my notation to make these methods more readily comparable. All methods estimate parameters from files containing the aligned reads.

\subsection{Estimating $M$, $U$}\label{sec:estimating_m_and_u}

- [x] DONE

The simplest and most commonly used estimators of $M_{i}$ and $U_{i}$ are $m_{i}$ and $u_{i}$, that is, the number of reads that are methylated and unmethylated at the $i^{th}$ locus (e.g., \cite{Cokus:2008fc, Lister:2008bh, Lister:2009hy, Lister:2011kg, Hansen:2011gu}). $M_{i}$ and $U_{i}$ may be estimated per-strand or aggregated across strands in the case of palindromic methylation loci. CpG methylation is typically aggregated across strands.

Not all reads, nor positions within reads, are equally treated in this counting process. Essentially, each sequenced nucleotide is assigned a weight. By far the most used weighting scheme is a set of filters that assigns each sequenced nucleotide a weight of 0 (observation excluded) or 1 (observation included). More complicated weighting schemes might include weighting by mapping quality and by base quality, although this would likely lead to fractional estimates of $M$ and $U$, which may not be desirable.

`methtuple` uses the same 'filter and count' method to estimate $MM_{(i, i + 1)}$, $MU_{(i, i + 1)}$, etc., that is, the frequency of methylation patterns at m-tuples.

\subsection{Estimating $B$}

- [ ] DONE

Most analyses of bisulfite-sequencing data have focused on estimating the average methylation level at individual methylation loci, $B_{i, j}$. The simplest estimator is $\beta_{i, j} = \frac{m_{i, j}}{m_{i, j} + u_{i, j}}$, which has been widely used \citet[e.g.,]{Cokus:2008fc, Lister:2008bh, Lister:2009hy, Lister:2011kg}). Under a Binomial model for the number of methylated reads at a locus, $M_{i, j} \eqd Binomial(n_{i, j}, B_{i, j})$, where $n_{i, j}$ is the sequencing coverage of the $i^{th}$ locus in the $j^{th}$ sample, $\beta_{i, j}$ is the maximum likelihood estimator of $B_{i, j}$.

Recently, more sophisticated methods have been proposed to model the average methylation level. These methods, which are based on $m_{i, j}$ and $u_{i, j}$, include beta-binomial models \citet{Feng:2014iq, Sun:2014fk, Dolzhenko:2014bo}, and the smoothing-methods proposed by \cite{Hansen:2011gu, Hansen:2012gr, Hebestreit:2013ko}. The aim of these methods isn't necessarily to estimate $B_{i, j}$ - it is typically to identify differential methylation (discussed below) - but estimates of $B_{i, j}$ can be obtained if so desired.

\subsubsection{Beta-binomial models}

- [x] DONE

Several papers have proposed the beta-binomial distribution as a natural model "for describing methylation levels of an individual site across replicates" \cite{Dolzhenko:2014bo}. The 'beta' component of the distribution models the underlying methylation level, $B_{i, j}$, while the 'binomial' component models the sampling of reads by sequencing. Another way to think of this is that the Beta component models the biological variability of the data, while the Binomial component models the sampling variability. This separation of biological and technical variability has proven successful in detecting differential gene expression from RNA-seq data. For example, the `edgeR` software \cite{Robinson:2010cw} uses the negative binomial distribution, which can be thought of as a gamma-poisson mixture distribution, to account for both the biological and sampling variability.

An attractive feature of the beta-binomial distribution is that it can be motivated by, and analysed, using Bayesian statistics, including empirical Bayes methods, or frequentist techniques such as maximum likelihood. For example, the software `DSS` \cite{Feng:2014iq} and `MOABS` \citet{Sun:2014fk} both use the beta-binomial distribution in an empirical Bayes analysis of differential methylation from bisulfite-sequencing data. In contrast, `RADmeth` \cite{Dolzhenko:2014bo} uses the beta-binomial model in a maximum likelihood framework to address the same problem.

\subsubsection{Smoothing $\beta$-values}

- [x] DONE

BSmooth, published in \citet{Hansen:2011gu, Hansen:2012gr} and available in the R/Bioconductor package `bsseq`, and `BiSeq`, published in \citet{Hebestreit:2013ko} and available in the R/Bioconductor package `BiSeq`, take a different approach to getting improved estimates of the $B_{i, j}$. Both `bsseq` and `BiSeq` use statistical smoothing of the 'raw' $\beta_{i, j} = \frac{m_{i, j}}{m_{i, j} + u_{i, j}}$. Smoothing is motivated and justified by the fact that the $B_{i, j}$ are spatially correlated within a sample, i.e., because of the presence of co-methylation, which is discussed in Chapters \ref{chap:co-methylation}.

Smoothing is particularly powerful for loci with low sequencing coverage, where the denominator $m_{i, j} + u_{i, j}$ is small and the corresponding standard error of $\beta_{i, j}$ is large. The smoothed $\beta$-values, and not the raw $\beta$-values, are then generally used in all downstream analyses.

__TODO: Discuss with Terry - is smoothing of the raw $\beta$-values still useful when you have high-coverage sequencing data?__

Both `bsseq` and `BiSeq` use a binomial local likelihood smoother. In each case this smoother was chosen because `BSmooth` and `BiSeq` model the number of methylated reads at the $i^{th}$ locus in the $j^{th}$ sample by $M_{i, j} \eqd Binom(n_{i, j}, B_{i, j})$. The smoothing is local to exploit co-methylation, which is a local phenomenon.

In both `bsseq` and `BiSeq` the raw $\beta$-values are weighted according to the binomial likelihood and the kernel function. The binomial likelihood weights points inversely to their standard error, $se(\beta_{i, j})$, and the kernel gives greater weight to those $\beta_{i, j}$ near the centre of the window. \citet{Lacey:2013iy} note that loci with very high sequencing coverage will strongly influence the smoother, potentially biasing estimates at neighbouring loci with low coverage.

`bsseq` assumes that for each sample that the underlying methylation level, $B_{i, j}$, is a smoothly varying function of the position in the genome, $i$. In contrast, `BiSeq` first creates clusters of CpGs and only assumes that the underlying methylation level is smooth at positions within each cluster.

Whenever smoothing is used, a key parameter is the bandwidth, which is the size of the window in which observations are included at each iteration of the smoother. `bsseq` uses a much larger window size than `BiSeq`; the default window size in `bsseq` is one that contains at least 70 CpGs and is at least 2000kb wide, whereas the default window size in `BiSeq` is 80bp, regardless of CpG-density. This is due to `BiSeq` being developed for RRBS data, which has a high CpG-density per window, whereas `bsseq` was developed for WGBS data, which has a lower and more variable CpG density per window.

Another 'parameter' choice when smoothing is the choice of kernel, although this is generally less important than the choice of bandwidth. `bsseq` uses a tricube kernel and `BiSeq` uses a triangular kernel.

\citet{Hebestreit:2013ko} and \citet{Lacey:2013iy} compare the smoothing results of `BiSeq` to `bsseq`. Both \citet{Hebestreit:2013ko} and \citet{Lacey:2013iy} provide instances where they claim `BiSeq` gives more 'reasonable' smoothed values than `bsseq`, but the comparison is selective and limited to a few regions. Furthermore, the comparisons are made using RRBS data, both real and simulated, and, `bsseq` is designed from WGBS[^rrbs_sim].

[^rrbs_sim]: Both \citet{Hebestreit:2013ko} and \citet{Lacey:2013iy} altered the default `bsseq` parameters to try to make them comparable to `BiSeq`. \citet{Hebestreit:2013ko} changed the default minimum window size to 80bp but still required at least 20 CpGs per window. \citet{Lacey:2013iy} kept the default minimum window size of 2000 bp but reduced the minimum number of CpGs per window to 50 from the default of 70.

\section{Statistical properties of $\beta$-values}\label{sec:beta}

- [x] DONE

Under the binomial model, $M_{i, j} \eqd Bin(m_{i, j} + u_{i, j}, B_{i, j})$ , $\beta_{i, j} = \frac{m_{i, j}}{m_{i, j} + u_{i, j}}$ is an unbiased estimator of $B_{i, j}$ with standard error $se(\beta_{i, j}) = \sqrt(\frac{\beta_{i, j}(1 - \beta_{i, j})}{n_{i, j}})$ \citep{Hansen:2012gr}. The natural interpretation of $\beta_{i, j}$ is then as an estimator of the average level of methylation at the $i^{th}$ locus in the $j^{th}$ sample. In this section I discuss this interpretation and statistical properties of this estimator.

\subsection{Distributions}

- [x] DONE

In a study involving multiple samples, the set of $\beta$-values can be summarised as a matrix where each row is a locus and each column is a sample. This matrix might be visualised to learn about the distribution of methylation levels, either row-wise, to learn about the variability across samples, or column-wise, to learn about the variability within samples.

Restricting our attention to CpGs, __FIGURE__ shows the column-wise summaries of the EPISCOPE data, the genome-wide distributions of $\beta$-values within each sample. What is immediately clear is that these distributions are bimodal. This bimodality is driven by the fact that CpGs in CpG islands are generally unmethylated whereas those outside of CpG islands are generally methylated, which we can see when we stratify CpGs by their CpG island status.

```{r}
# TODO: Show a bunch of plots of beta-values for different samples
```

```{r}
# TODO: Show a bunch of plots of beta-values stratified by CGI-status for different samples
```

To do the same plots for the row-wise summaries would be rather tedious since there are many, many more rows (loci) than columns (samples) in the matrix of $\beta$-values. However, a few examples are shown in __FIGURE__. These row-wise summaries are often modelled by the beta distribution \cite[e.g.,]{Hebestreit:2013ko, Lacey:2013iy}. The Beta distribution is a flexible 2-parameter distribution on $[0, 1]$. It can be unimodal, "U"-shaped or "J"-shaped, depending on the choice of parameters. The Beta distribution also includes the Uniform and arcsine distributions as special cases (__TODO: better source than wikipedia__).

The analysis of the row-wise distributions leads to the analysis of differential methylation and differential variability, which are discussed in __SECTION__.

\subsection{Correlations}

- [x] DONE

Many researchers have observed that DNA methylation is spatially correlated along the genome, \cite[e.g.,]{Eckhardt:2006gh, Cokus:2008fc, Li:2010fb, Hansen:2011gu, Hebestreit:2013ko, Wang:2011cw, Pedersen:2012vl, Lacey:2013iy, Sofer:2013bk, Liu:dy, Lyko:2010dr, Landan:2012kp, Lister:2009hy}). I call this correlation of methylation levels _co-methylation_.

Just as we can explore the distribution of DNA methylation levels within a sample or between samples, so too can we explore co-methylation within a sample or between samples. I examine this in detail in Chapter \ref{chap:co-methylation}

\subsection{Biases}

- [x] DONE

The natural interpretation of $\beta_{i, j}$ as the average level of methylation at the $i^{th}$ locus in the $j^{th}$ sample will be biased if the probability of sequencing a fragment with a methylated site is different from the probability of sequencing a fragment with an unmethylated site. In fact, it has been shown that methylated DNA is overrepresented in bisulfite-sequencing data due, with the problem exacerbated by higher rounds of PCR amplification and dependent on the bisulfite-conversion protocol \cite{Ji:2014fq}. PCR amplification can result in overreprestation of one of the DNA strands in bisulfite-sequencing data \cite{Warnecke:1997eh}.

Lab-based solutions to overcome these biases exist for targeted bisulfite-sequencing, but are technically difficult and expensive to extend to whole-genome studies. \cite{Ji:2014fq} propose potential computational corrections for these biases but as yet these have not been implemented.

\subsection{Transformations}

__UP TO HERE__

- [ ] DONE

$\beta$-values are the _de facto_ standard unit for reporting methylation levels due to their natural interpretation as the average level of methylation at the locus. However, they are not necessarily the best unit for statistical inference. This is because a $\beta$-value is an estimate of a proportion and there are a well-known challenges when working with proportion data, such as:

1. Proportions are bound between 0 and 1, inclusive.
2. The estimate of the standard error depends on the estimate of the mean (i.e., $\beta$), through $se(\beta) = \sqrt{\frac{\beta (1 - \beta)}{m + u}}$. Taking the derivative of this, we see that the maximum standard error, $\sqrt(\frac{0.25}{m + u})$, occurs at $\beta = 0.5$ and the minimum standard error, $0$, occurs at $\beta = 0, 1$.
3. We need to more than just the $\beta$-value to have a sense of how precise an estimate it is. Essentially, we need to also know the sequencing coverage of the methylation loci. Consider two CpGs, one with $m = 1, u = 3$ and the other with $m = 100$ $u = 300$. Both CpGs have $\beta = 1/4$ but the second CpG is measured with much greater precision. Assuming the binomial model, the first CpG has $se(\beta) = \sqrt{\frac{1/4 \times 3/4}{4}} = 0.22$ whereas the second CpG has $se(\beta) = \sqrt{\frac{1/4 \times 3/4}{400}} = 0.02$.

To address (2), proportion data are often transformed via a variance stabilisation transformation. The aim is to make the variance (approximately) independent of the mean. Popular variance stabilisation transformations include:

- The arcsine transformation, $\arcsin{\sqrt{\beta}}$ \citep{ANSCOMBE:1948bw}. A small value, e.g. $0.5$, is added to $m$ and $u$ to avoid $\beta = 0, 1$.
- The "averaged arcsine" transformation, $\arcsin{\sqrt{\frac{m}{m + u + 1}}} + \arcsin{\sqrt{\frac{m + 1}{m + u + 1}}}$ \citep{Freeman:1950bh}. This transformation does not have a unique inverse \citep{Nunes:2009vj}.

The use of variance stabilising transformations for proportion data seem to have fallen out of favour (e.g. [http://www.esajournals.org/doi/full/10.1890/10-0340.1](http://www.esajournals.org/doi/full/10.1890/10-0340.1)). The now favoured approach is generalised linear models (__CITE__), in particular the logistic regression model (__CITE__?).

__TODO: Cite beta-values vs. M-value paper and discuss__

\section{Downstream analyses}\label{sec:downstream}

- [ ] DONE

\subsection{Regression models}\label{sec:regression_models}

- [ ] DONE

* Logistic regression
  * M-values
  * The logit transformation, $\text{logit} (\frac{\beta}{1 - \beta})$. A small value, e.g. $0.5$, is added to $m$ and $u$ to avoid $\beta = 0, 1$.
  * The probit transformation, $\Phi^{-1}(\beta)$, where $\Phi$ is the standard Normal cumulative distribution function. A small value, e.g. $0.5$, is added to $m$ and $u$ to avoid $\beta = 0, 1$.
* Beta regression

\cite{Du:2010dc} show improved inferences when using Logit transformed $\beta$-values, which they call M-values. This transformation reduces the heteroscedasticity of $\beta$-values near 0 and 1.

\section{General TODOs}

* Autocorrelation or correlation?
* Discuss distribution of coverage? Perhaps in context of sequencing bias?