### 2014-11-25 Rmarkdown/biology_background.Rmd
Polishing biology background

---
title: "Biology background"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

\section{Chapter Overview}\label{sec:biology_background_chapter_overview}

In this chapter I summarise the basic biology necessary for
understanding my thesis. I introduce DNA methylation, describe some of its biological roles and assays for studying DNA methylation, with a particular focus on bisulfite-sequencing.

\section{From DNA to genomes}\label{ref:from_dna_to_genomes}

The genome is the genetic material of an organism. In most organisms, the genome is encoded by deoxyribonucleic acid, _DNA_. In eukaryotes[^eukaryotes], which includes plants, fungi and animals, the DNA is wound around repeating units of eight _histone_ protein cores to form _nucleosomes_, which are the fundamental unit of _eukaryotic chromatin_. Chromatin compactly packages the DNA into _chromosomes_, so that the organism's complete nuclear DNA, its nuclear genome, might fit into the nucleus of its cells[^extra_nuclear_genomes].

The material in this chapter can be found in any good introductory molecular biology text, such as \cite{Alberts:2007vj}.

[^eukaryotes]: Eukaryotes are organisms composed of one or more cells with a distinct nucleus and cytoplasm. Most of a eukaryotic cell's DNA is contained in the nucleus \cite{Alberts:2007vj}.

[^extra_nuclear_genomes]: Not all of an organism's genome is present in the nucleus of a cell. Important exceptions are mitochondrial DNA (mtDNA) found in animals, plants and fungi and chloroplast DNA (ctDNA) found in plants.

\subsection{DNA}\label{sec:DNA}

The double helix is the most common, and most famous, structure of DNA. In the double helix, the two strands of DNA run in opposite directions to each other and are therefore _anti-parallel_. One strand is called the 5' strand (pronounced "5 prime strand") - also known as the sense strand, Crick strand or top strand - and the other strand called the 3' strand (pronounced "3 prime strand") - also known as the antisense strand, Watson strand or bottom strand. Along each strand of the double helix are the four DNA nucleobases (_bases_): adenine (_A_), cytosine (_C_), guanine (_G_) and thymine (_T_). These bases form _complementary_ base pairings, _A_ with _T_ and _C_ with _G_, along the DNA double helix. This is illustrated in Figure \ref{fig:double_helix}.

![Simple diagram of double-stranded DNA showing complementary base pairings. By Forluvoft (Own work) [Public domain], via Wikimedia Commons [http://commons.wikimedia.org/wiki/File%3ADNA_simple2.svg](http://commons.wikimedia.org/wiki/File%3ADNA_simple2.svg)\label{fig:double_helix}](../figures/DNA_simple2.pdf)

A gene is a sequence of DNA that is _transcribed_ to produce a functional product in the form of ribonucleic acid, _RNA_. RNA may in turn be translated into a protein sequence or perform other roles in the regulation of gene expression. It is important to note here than not all DNA is transcribed into RNA, which is not to say that untranscribed DNA is unimportant. For instance, there are untranscribed _regulatory sequences_ of DNA that determine whether a nearby gene is transcribed. There is also _junk DNA_ that is of little consequence to the organism \cite{Alberts:2007vj}. Conversely, not all transcribed DNA is a gene. DNA transcription is _permissive_ and there are many DNA sequences that are transcribed by accident or in error.

DNA is able to self-replicate. This means that, for example, eukaryotic cells created during _mitosis_ contain the same DNA as the 'parent' cell. During DNA replication, the two strands are separated and each strand's _complementary_ DNA sequence is copied by an enzyme called _DNA polymerase_. It is because the two strands of DNA are complementary that ensures the daughter cell contains the same DNA sequence as the parent cell[^dna_replication].

[^dna_replication]: This of course ignores errors in the replication process. Such errors are very rare events but because DNA replication happens so frequently these events do occur. There are error-correcting processes that reduce the chance that such an error is retained in the daughter sequence, however, these are not perfect. Hence errors in DNA replication are one source of what are known as _mutations_ in the DNA.

\subsection{Nucleosomes and chromatin}\label{sec:nucleosome_and_chromatin}

The core of a nucleosome consists of four pairs of histones, H2A, H2B, H3 and H4, which are consequently known as the _core histones_. A fifth histone, H1/H5, is known as the _linker histone_. Each of these histones has a 'tail' consisting of a string of amino acids. These tails can undergo _post-translational modifications_, such as methylation, acetylation and phosphorylation, which can alter their interactions with DNA and nuclear proteins \cite{Alberts:2007vj} (discussed in Section \ref{sec:epigenetics} on epigenetics).

The nucleosomes are interconnected by _linker DNA_ to form the macromolecule called chromatin. The linker sequences are between $20-60$ nucleotides in length, while approximately 147 nucleotides are wrapped around each nucleosome and a further 20 nucleotides wrapped around each additional H1/H5 histone \cite{Annunziato:2008ue}.

Chromatin is frequently described as either _heterochromatin_ or _euchromatin_. Heterochromatin, or closed chromatin, is more tightly packed than euchromatin, or open chromatin. Heterochromatin is associated with transcriptionally repressed regions of the genome because the machinery required to translate DNA to RNA is unable to physically access the DNA. In contrast, euchromatin is associated with transcriptionally active regions of the genome since its more open nature allows easier access for the translational machinery.

The same region of an individual's genome may vary between heterochromatin and euchromatin states at different stages of the organism's life and in different cells of the organism. This is one mechanism which by gene expression is regulated.

Basic descriptions of nucleosomes and chromatin are ripe with analogies. In one such popular analogy, the histones are the 'spool' around which the DNA 'thread' is wrapped to form nucleosomes. The chromatin then has the appearance of "beads on a string" when viewed under an electron microscope \cite{Alberts:2007vj}. Chromatin is further coiled up into various literally-named structures, such as the 30-nanometre and 250-nanometre fibres, and, ultimately, packaged into _chromosomes_. The set of chromosomes makes up an individual's _genome_.

\subsection{Genomes}\label{sec:genomes}

As is clear from the above description, the genome is a complex three-dimensional structure. Nonetheless, in bioinformatics and computational biology, the genome is mostly considered as a single-stranded, one-dimensional string of the nucleobases _A_, _C_, _G_ and _T_.

While eukaryotic genomes share the above-described features, and indeed share many regions of common DNA sequence, eukaryotic genomes come in many shapes and sizes. I only discuss the genomes of three organisms relevant to my thesis: Human (_Homo sapiens_), Mouse (_Mus musculus_) and Arabidopsis (_Arabidopsis thaliana_).

\subsubsection{Human genome}\label{sec:human_genome}

Humans are _diploid_ organisms, meaning that we have two copies of each chromosome in a typical cell[^somatic_cell]. We inherit one chromosome of each pair from our mother and one from our father. A typical human cell has 23 pairs of nuclear chromosomes, 22 _autosomes_ and 1 pair of _sex chromosomes_, as well as hundreds or thousands of copy of the small mitochondrial chromosome[^mtDNA].

[^somatic_cell]: A sperm or egg cell is haploid and has a single (recombined) copy of each chromosome.

[^mtDNA]: The mitochondrial DNA is maternally inherited.

The length of a chromosome is typically reported as the number of DNA base pairs (bp) in a single copy of that chromosome and the _haploid_ length of a genome is the sum of these chromosome lengths. The haploid human genome is approximately 3 billion base pairs long (Golden Path Length [http://asia.ensembl.org/Homo_sapiens/Location/Genome?r=1](http://asia.ensembl.org/Homo_sapiens/Location/Genome?r=1)).

A human _reference genome_ was jointly completed by the International Human Genome Consortium and Celera Genomics in 2003 \cite{Venter:2001fq,Lander:2001hk}. This reference genome does not represent the genome of any one human since it uses DNA donated by several different people \cite{Venter:2001fq,Lander:2001hk}. Rather, a reference genome is a kind of map or scaffold that can be used to identify similarities and differences between individual genomes.

The human genome has obvious uses in medical research and biotechnology, but is also used to learn about evolution and human history, such as human migration patterns \cite{Hellenthal:2014ky}. Every person, even a _monozygotic_ ('identical') twin \cite{Hellenthal:2014ky}, has their own unique genome. However, genomes of any two randomly selected people are identical at approximately $99.9\%$ of sites. Furthermore, the vast majority, $98\%$ by some estimates \cite{Elgar:2008dm}, of the human genome is made up of non-coding DNA and upwards of $50\%$ is repetitive sequence \cite{Treangen:2012dn}.

\subsubsection{Mouse genome}\label{sec:mouse_genome}

Mice are also diploid organisms, but have 19 pairs of autosomes, one pair of sex chromosomes and a mitochondrial chromosome. The mouse genome is slightly smaller than the human genome, at 2.7 billion base pairs long (Golden Path Length [http://asia.ensembl.org/Mus_musculus/Location/Genome?r=1](http://asia.ensembl.org/Mus_musculus/Location/Genome?r=1)). Like the human genome, there is a mouse reference genome \cite{MouseGenomeSequencingConsortium:2002fn}. It is based on several female mice from the _C57BL/6J_ strain, an important strain of mouse widely used in medical research.

Mouse strains used in medical research are highly inbred due to years of concerted mating programs. This reduced genetic variability, and the control that researchers have over it, make these mice a very powerful tool in identifying the biological cause of a diverse range of phenotypes.

\subsubsection{Arabidopsis thaliana genome}\label{sec:arabidopsis_genome}

Arabidopsis is a small flowering plant. The most widely studied species is _A. thaliana_, which is diploid and has one of the smallest genomes known in the plant kingdom at 135 million base pairs long, which is packaged as five pairs of chromosomes ([http://plants.ensembl.org/Arabidopsis_thaliana/Info/Index](http://plants.ensembl.org/Arabidopsis_thaliana/Info/Index)). The small size of its genome, coupled with its small physical size, rapid development and its tolerance to both cross-pollination and self-pollination[^cross_and_self_pollination], make _A. thaliana_ an ideal organism for studying plant development.

[^cross_and_self_pollination]: The advantage in genetic studies of using a plant that is tolerant of self-pollination is that recessive mutations can be quickly made homozygous and thus expressed. Cross-pollination allows the creation of new strains with multiple mutations.

\subsection{Genetic variation}\label{sec:genetic_variation}

A key question in biology, perhaps _the_ key question, is what determines an individual's _phenotype_. An individual's phenotype is the set of its observable characteristics resulting from the interaction of its genotype with the environment. Two simple examples of phenotypes are height and weight. Both have a genetic component, e.g., the offspring of tall parents are on average taller than the offspring of short parents, but also have environmental components, such as the contribution of diet to weight. In medical research, a person's phenotype might be whether she is affected by a particular disease. It might also be some proxy, such as her blood pressure or the expression levels of particular genes.

Variation between individual's genomes, be it at single nucleotides[^snvs_snps] or across larger regions[^indels], is one important source of phenotypic variation. Importantly, genetically-driven phenotypic variation is frequently _heritable_, meaning that phenotypes can be passed on from one generation to the generation via the genome.

[^snvs_snps]: An individual nucleotide that is variable in the population is called a single nucletoide variant (_SNV_). A SNV that is frequently variable in the population, say at least $1\%$ of the population are variable at that site, is called a single nucleotide polymorphism (SNP). All SNPs are SNVs but the converse is not true.

[^indels]: An indel is a short insertion or deletion of DNA sequence in an organism, usually relative to a reference genome.

Environmental variation has clear influence on certain phenotypes. However, it often can be difficult to determine whether phenotypic variation is driven by genetic varation or environment variation, particularly in humans where genetically similar individuals typically also grow up in similar environments.

The above discussion has been about phenotypic variation in a population. But there is also phenotypic variation within the individual. If that sounds strange, consider the fact that in your body a neuron, a leukocyte (a white blood cell) and a cone cell (a photoreceptor in the retina) all have identical genomes. In fact, all cells in an organism, excluding the gametes, have an identical genome yet play very different biological roles. This is due to different genes being active in different cells.

_Epigenetics_, which I describe in the next section, plays a role in determining both the between-individual phenotypic variation as well as the within-individual phenotypic variation.

\section{Epigenetics}\label{sec:epigenetics}

_Epigenetics_ is a real Humpty Dumpty phrase; each author seems to believe that, "When I use [the] word, ... it means just what I choose it to mean - neither more nor less" \cite{carroll1897through}. I am sure to continue in this grand tradition.

Conrad Waddington coined the phrase in 1942 as a portmanteau of the words "epigenesis" and "genetics" \cite{Waddington:2012bu}. Waddington meant epigenetics as the study of how "processes involved in the mechanism by which the genes of the genotype bring about phenotypic effects" \cite{Waddington:2012bu}.

A popular contemporary definition of epigenetics is attributed to the epigeneticist Arthur Riggs - epigenetics is "the study of mitotically and/or meiotically heritable changes in gene function that cannot be explained by changes in DNA sequence[^mitotic_meiotic]" \cite[p1]{russo1996epigenetic}. The _epi_ prefix, derived from the Greek word for 'upon', 'near to', or 'in addition', emphasises the idea that epigenetics encodes information 'on top of' the DNA sequence. However, it is quite different to Waddington's original definition.

[^mitotic_meiotic]: Mitotically heritable means heritable during cell division and meitotically heritable means heritable during sexual reproduction.

More recently, the definition of epigenetics has taken on a "more biochemical flavour" \cite{Daxinger:2010bu} to include marks whose heritability is yet to be established. The heritability, or lack thereof, of histone modifications means that many epigeneticists do not consider these to be truly epigenetic \cite{Berger:2009ex}, and describing them as such is a sure-fire way to annoy a good percentage of your audience \cite{Ledford:2008ew}.

Sir Adrian Bird, an esteemed British geneticist and someone who is in favour of the broader definition, attempts to unite these definitions in the following \cite{Bird:2007em}:

> "The following could be a unifying definition of epigenetic events: the structural adaptation of chromosomal regions so as to register, signal or perpetuate altered activity states. This definition is inclusive of chromosomal marks [e.g. histone marks], because transient modifications associated with both DNA repair or cell-cycle phases and stable changes maintained across multiple cell generations qualify."

Regardless of which definition you subscribe to, an _epigenetic mark_ is the modification that causes this "epigenetic change". Actually, "causes" may be too strong a phrase, as much of current epigenetics research is in identifying associations rather than causations, and the question of whether the epigenetic mark is a cause or consequence of the ascribed function is oft-debated.

The _epigenome_ of a cell is the set of epigenetic marks present on the cell's genome. In contrast to the genome, which is identical between cells within an individual, the epigenome is highly variable between cells within an individual. Indeed, we can identify variation for a single epigenetic mark within cells of the same cell type from the same individual as well will see in later chapters. _Epigenomics_ is the study of the epigenome, analogous to genomics being the study of the genome. However, one can rarely study the epigenome in isolation from the underlying genetic sequence, as there is evidence that the epigenetic variation is associated with genetic variation \cite{McVicker:2013ih,Bell:2011fm,Zhang:2010bg,vanEijk:2012hc}

One mark that most authors agree is an epigenetic mark is DNA methylation, which I describe in the next section and whose study is the focus of my thesis.

\section{DNA methylation}\label{sec:dna_methylation}

DNA methylation is a chemical modification of DNA that can impart information on top of the DNA sequence. It is heritable during mitotic cell division, which means that it is faithfully copied across to the daughter cell during cell division[^dnam_copying]. It therefore fits into Arthur Riggs' afore-mentioned definition of an epigenetic modification \cite[p1]{russo1996epigenetic}. DNA methylation is found in bacteria, fungi, plants and animals.

[^dnam_copying]: In practice this copying is not as faithful as, say, the copying of DNA from the parent to the daughter strand. Furthermore, the faithfulness of this copying will be different in different conditions, such as in a healthy liver cell compared to a cancerous liver cell. Nevertheless, the copying of DNA methylation is faithful enough for most biologists to consider it as a mitotically heritable mark, most of the time.

A major reason that DNA methylation is studied is that it is essential for normal development in mammals \cite{Li:1992bp}. It is also involved in many key biological processes related to human development and health, such as the regulation of gene expression \cite{Razin:1991vj}, silencing of transposable elements\cite{Jones:2001eg}, X chromosome inactivation \cite{Mohandas:1981uf} and tumorigenesis \cite{Ehrlich:2002kz}.

However, DNA methylation is not found in all organisms, for example, it not detectable in yeast \cite{Capuano:2014cu}, and so is not essential for life. Furthermore, the level of DNA methylation varies widely amongst different organisms, with many having very low levels of methylation\cite{Capuano:2014cu}.

When people speak of DNA methylation they are generally referring to methylation of the cytosine nucleobase. Even more specifically, they are referring to the molecule _5-methylcytosine_. 5-methylcytosine, abbreviated as _5mC_, is by far the most common form of DNA methylation in the animal and plant kingdoms[^other_DNAm]. However, I will continue to use the term _DNA methylation_ to describe the more specific 5mC, as is standard in the literature.

[^other_DNAm]: Two additional forms of DNA methylation are N6-methyladenine (m6A) and N4-methylcytosine (m4C). N6-methyladenine is a methylated form of adenine, which is found in mRNA \cite[e.g.,]{Adams:1975ki} and DNA, although the latter only in bacterial DNA \cite{Ratel:2006dv}. N4-methylcytosine has also been detected in bacterial DNA \cite[e.g.,]{Ehrlich:1985bv, Ratel:2006dv}.

A German chemist, W.G. Ruppel, first identified a methylated nucleic acid in 1898. Ruppel was studying _tuberculinic acid_, the poison of _Mycobacterium tuberculosis_[^tuberculosis] , and discovered that it contained a methylated nucleotide \cite{Ruppel:1899hc}. In 1925, Johnson and Coghill isolated 5-methylcytosine as a product of hydrolysis of tuberculinic acid, the nucleic acid of _Mycobacterium tuberculosis_ \cite{Johnson:1925js}. However, Johnson and Coghill's results were disputed for over twenty years by other researchers who were unable to replicate the original findings \cite{VISCHER:1949ty}.

[^tuberculosis]: _Mycobacterium tuberculosis_ was then known as _Tubercle bacillus_.

In 1945, Hotchkiss ultimately proved Johnson and Coghill correct when he isolated _5-methylcytosine_ from nucleic acid prepared from cow thymus \cite{HOTCHKISS:1948va}. Using paper chromatography, Hotchkiss demonstrated that methylated cytosine existed and was distinct from conventional cytosine and uracil.

The typical site of DNA methylation is at the C5 carbon position of a cytosine nucleotide. This modified cytosine is called _5-Methylcytosine_ or _5mC_. Figure \ref{fig:5mC} shows the structure of 5mC.

\begin{figure}[h!]
\centering
\includegraphics[width=0.3\textwidth]{../figures/5-methylcytosine.png}
\caption{Chemical structure of 5-methylcytosine. "5-methylcytosine". By Yikrazuul (Own work) [Public domain], via Wikimedia Commons [http://commons.wikimedia.org/wiki/File:5-Methylcytosine.svg](http://commons.wikimedia.org/wiki/File:5-Methylcytosine.svg)}
\label{fig:5mC}
\end{figure}

It is common to describe a cytosine as being 'methylated' or 'unmethylated', however, care must be taken in defining these terms. At the lowest level, the level of single-stranded DNA, methylation is a binary event: a cytosine is either methylated or unmethylated. Double-stranded DNA, at least at _palindromic_ methylation loci[^palindromic_sequence], is generally symmetrically methylated, i.e., the loci on each strand are both methylated or both unmethylated. However, _hemimethylation_, where the methylation loci on one strand is methylated and its partner on the opposite strand is unmethylated, can and does occur.

[^palindromic_sequence]: A palindromic DNA sequence is one that is identical on when read in the 5' to 3' direction on both the original strand and the complementary strand of the double helix. For example, _CG_ is a palindromic sequence.

Within a diploid cell, a particular cytosine may be unmethylated or methylated on both homologous chromosomes or methylated on one chromosome and unmethylated on the other. While the former is more common, examples of the latter case, such as _allele-specific methylation_ \cite{Shoemaker:2010jz} and _genomic imprinting_ \cite{Li:1993ea}, are of considerable biological interest.

\subsection{DNA methylation in mammals}\label{sec:dna_methylation_in_mammals}

The importance of 5-methylcytosine in mammalian genomes is such it has been dubbed the "fifth base" of the DNA code \cite{Lister:2009boa}. In mammalian genomes, most cytosines are unmethylated except for those at _CpG dinucleotides_. A _CpG dinucleotide_, or more simply a _CpG_, is a cytosine followed by a guanine in the linear DNA sequence. The 'p' stands for the phosphate backbone of DNA and some authors omit it in favour of simply calling it CG methylation. A CpG is a palindromic sequence and is generally symmetrically methylated. Approximately 70% of CpGs are methylated in mammals \cite{Laird:2003kw}, meaning that the cytosine in the CpG dinucleotide is a 5-methylcytosine.

\subsection{CpG dinucleotides}\label{cpg_dinucleotides}

CpGs are underrepresented in the human genome. The GC-content of each the human genome, which is defined as the percentage of nucleotides that are either guanines or cytosines, is between $41\%$. If the nucleotides were uniformly distributed across the human genome we would expect about $4.1\%$ of dinucleotides to be Cpgs. Instead, only $1\%$ of dinucleotides are CpGs. See the Appendix for details of these calculations and a figure comparing the observed:expected ratio of each dinucleotide in the human reference genome (hg19).

One reason for this relative scarcity of CpGs is that methylated cytosines can spontaneously deaminate to thymines \cite{Scarano:1967ue}. Thus, over time, many methylated CpGs will become TpGs, leading to a genome-wide reduction in the proportion of CpGs and a genome-wide increase in the proportion of TpGs (see Figure \ref{fig:dinuclotide_distribution}). There are many other evolutionary pressures on the distribution of nucleotides in a genome. One effect of this is that the distribution of CpGs is far from uniform. In fact, CpGs tend to form clusters, which are termed CpG islands.

\subsection{CpG islands and other sandy metaphors}\label{sec:cpg_islands}

One way to explore the distribution of CpGs in the human genome is to look at the distribution of distances from one CpG to the next, the _intra-pair distances_ (IPDs). __FIGURE__ is the empirical cumulative distribution function of CpG IPDs for the human reference genome (hg19). We see that approximately $70\%$ of CpGs are within 100 bp of the next CpG. __FIGURE__ also shows the expected IPD distribution under a model where CpGs are uniformly distributed along the genome with probability equal to the observed frequency of CpGs on each chromosome. By comparing the observed IPDs to the expected IPDs we see that the distribution of distances between CpGs has more "close" pairs and more "distant" pairs than we would expect by chance.

__FIGURE__ plots a frequency polygon of the observed IPDs. From this we can see there is a cluster of CpGs with $IPD < 10$, which largely correspond to what are called _CpG islands_ (CGIs).

Within these CpG islands are the 20-40% of CpGs that are frequently unmethylated in mammalian genomes. CpG islands are important regulatory elements in the genome and are where most differences in DNA methylation between different cell types are found \cite{Wu:2010do}. The classical definition of a CpG island, given by \cite{GardinerGarden:1987we}, and used by, for example, the UCSC genome browser, is a region of the genome where:

1. The (moving) average of GC-content is greater than $50\%$, and
2. The observed-to-expected ratio of CpGs is greater than $0.6$, and
3. The region is longer than 200 bp.

This definition was refined by \cite{Takai:2002jb} to exclude _Alu_-repetitive elements, which otherwise are misclassified as _bona fide_ CpG islands. More recently, \cite{Wu:2010do} developed a hidden Markov model to predict CpG islands based on the CpG density and GC-content of the region.

An alternative definition, and one that pre-dates the definition of \cite{GardinerGarden:1987we}, is based on the identification of unmethylated regions of the genome, which are typically CpG-dense. Such regions were previously called HpaII tiny fragment islands, or _HTF islands_, and named after the restriction enzyme used to identify them \cite{Cooper:1983ev, Bird:1985gj}.

These "sandy/beachy" metaphors have been continued (i.e., stretched to breaking point) with various authors defining CpG island shores, CpG island shelves, CpGs in the "open sea", CpG deserts and CpG canyons. CpG island shores, shelves and the open sea are all defined with respect to CpG islands. Specifically,

- CpG island shores are regions within 2 kb of CpG islands \cite{Irizarry:2008hg}. CpG island shores have been demonstrated to have an increased variability of CpG methylation \cite{Irizarry:2008hg}. Shelves have a comparatively low density of CpGs.
- CpG island shelves are defined as regions within 2 kb of a CpG island shore \cite{Bibikova:2011fw}.
- The "open sea" contains those CpGs not classed as being in a CpG island, CpG island shore or CpG island shelf \cite{Sandoval:2011dj}.

Other metaphors, these based on methylation levels rather than CpG density, include methylation deserts \cite{Li:2012gg} and CpG canyons \cite{Jeong:2014bla}. Because these regions are defined with respect to methylation levels rather than DNA sequence, these are generally identified in a tissue-specific manner.

\subsection{Non-CpG methylation}\label{sec:non_cpg_methylation}

In humans, cytosine methylation in most cell types is found almost exclusively at CpGs \cite{Jones:2012jr}. There are, however, certain cell types with widespread non-CpG methylation. Non-CpG methylation is often classified as CHH methylation or CHG methylation, where H is the IUPAC code for any nucleotide except G ([http://www.bioinformatics.org/sms2/iupac.html](http://www.bioinformatics.org/sms2/iupac.html)). The rule-of-thumb for mammalian genomes is that non-CpG methylation is rare in somatic cells but common in pluripotent cells. Of course, there are exceptions to every rule, especially in biology.

To give a few examples, \cite{Lister:2009hy} found that in a _fibroblast_ cell line that $99.98\%$ of cytosines that displayed statistically significant evidence of methylation occured at CpGs. In contrast, in an embryonic stem cell line, they found that $24.5\%$ of cytosines that displayed statistically significant evidence of methylation occured in a non-CpG context. However, it should also be noted that these non-CpG loci that were methylated had on average a much lower level of methylation than their CpG counterparts.

A subsequent paper from the same group extended this result. \cite{Lister:2011kg} reported that, more generally, non-CpG cytosines account for $20-30\%$ of cytosines with statistically significant evidence of methylation in _pluripotent_ cell lines . Pluripotent cells includes embryonic stem cells along with induced pluripotent stem cell lines.

An exception to the rule that non-CpG methylation is largely restricted to pluripotent cells is provided by \cite{Lister:2013et}, where they found that neurons also have non-CpG methylation, albeit at a lower level ($1.3-1.5\%$ of all non-CpG cytosines were methylated).

Overall, non-CpG methylation in humans is less well studied and less well understood than CpG methylation. This is partly due to sampling bias since commonly used assays for studying DNA methylation, such as the Illumina 27k and 450k beadchips, only measure CpG methylation. However, recent technological advances mean that cytosine methylation can be routinely assayed regardless of the sequence context. This discussed further in Section sec:dna_methylation_assays.

Non-CpG methylation is very common in other organisms, such as plants. For example, \cite{Lister:2008bh} found that in the widely-studied _Arabidopsis Thaliana_ that $45\%$ of cytosines that displayed statistically significant evidence of methylation are at CHG or CHH loci. However, they also found that the level of methylation at non-CpG loci is typically lower than that observed at CpG dinucleotides.

\subsection{Modifications of a modification}\label{sec:modifications_of_a_modification}

Methylation is not the only chemical modification of cytosines, although it is by far the most common. These additional modifications may be thought of as 'modifications of a modification' since they extend 5-methylcytosine. Listed from most frequent to least frequent, these are are 5-hydroxymethylcytosine (_5hmC_), 5-formylcytosine (_5fC_) and 5-carboxylcytosine (_5caC_) \cite{Plongthongkum:2014jg}. The biological significance of these marks is still being determined, in part because the assays for studying these are still in development and because their relatively scarcity in the genome means that experiments to detect these modifications are more difficult and expensive.

One genome-wide study of 5hmC found that less than $1\%$ of all assayed cytosines in mouse fetal cortex and adult cortex cells displayed any statistically significant evidence of hydroxymethylation \cite{Lister:2013et}. These cytosines appeared to be restricted to the CpG context and had only very low levels of 5hmC.

\cite{Kriaucionis:2009bm} and \cite{Tahiliani:2009kl} discovered that the TET enzymes can convert 5mC to 5hmC, 5hmC to 5fC and 5fC to hcaC. This suggests a role for 5hmC, 5fC and 5caC in the process of removing 5mC marks, as is discussed in the next section.

\subsection{Writers, readers and erasers}\label{sec:writers_reads_and_erasers}

A common analogy used in describing epigentic marks uses 'writers', 'readers' and 'erasers' \cite{Moore:2013in}. In the case of DNA methylation, writers catalyse the methyl group onto the DNA, readers recognise methylated DNA and erasers remove the methyl group from the DNA.

In mammalian cells, the writers are the DNA _methyltransferase_ (DNMT) enzymes. The DNMTs are commonly split into two groups, namely the maintenance methyltransferase and the _de novo_ methyltransferases. DNA methylation is not preserved by the DNA replication machinery and so it is the role of the maintainence methyltransferases to restore the methylation pattern on the daughter strand of DNA following DNA replication. In mammals, _DNMT1_ is known as the maintenance DNA methyltransferase. _DNMT3a_ and _DNMT3b_ are known as the _de nove_ methyltransferases, although these are also required for the maintenance of DNA methylation \cite{Jones:2009ji}. Both _DNMT1_ and _DNMT3b_ appear to be essential for mammalian development as mouse knockouts[^KO] for either gene is embryonically lethal \cite{Li:1992bp}. In contrast, mouse knockouts for _DNMT3a_ are runted but survive for ~4 weeks after birth \cite{Li:1992bp}.

[^KO]: A knockout mouse for gene $X$ is a mouse that has been genetically engineered to remove or otherwise inactivate gene $X$. Mouse knockouts can be either heterozygous knockouts (one copy still of the gene is still present/active) or homozygous knockouts (both copies of gene absent/inactive).

_DNMT2_, now known as _TRDMT1_, was once thought to be a DNA methyltransferase but was shown to in fact methylate a small RNA and not DNA \cite{Goll:2006kb}. Another protein, DNMT3L, is homologous to DNMT3a and DNMT3b but does contain catalytic domain that is necessary for methyltransferase activity. Instead, DNMT3L is thought to stimulate the activity of DNMT3a and DNMT3b (__TODO: cite__).

The readers of DNA methylation recognise methylated DNA. These readers can recruit additional proteins to the site of the methylated cytosine to perform a variety of functions related to gene expression. For example, the methyl-CpG-binding domain (_MBD_) group of proteins bind to DNA containing a methylated CpG, which can then suppress gene expression by preventing transcription factor binding at that site \cite{Nan:1993gma}. Another group, the ubiquitin-like, containing PHD and RING finger domain (_UHRF_) proteins, help _DNMT1_ methylate hemimethylated DNA, such as the daughter strand created during DNA replication \cite{Sharif:2007ft, Bostick:2007jh}.

The removal or erasure of DNA methylation, called _demethylation_, is characterised as _passive_ loss or _active_ removal. Passive loss occurs when the maintenance methyltransferases do not efficiently perform their role of restoring DNA methylation following cell division. This leads to a gradual, stochastic and genome-wide loss of DNA methylation after multiple cell divisions. This form of passive demethylation, sometimes called replication-dependent demethylation, cannot explain observations of local tissue-specific differences in DNA methylation \cite{Irizarry:2009ky} nor the two stages of rapid global demethylation that occur during development \cite{Wu:2014gw}.

Active demethylation is currently a hot topic in epigenetics research. Multiple mechanisms have been proposed, and it is indeed likely that there are multiple ways to achieve active demethylation. These mechanisms were recently reviewed by \cite{Wu:2014gw}, which I briefly summarise:

1. The direct removal of the methyl group from _5mC_ is considered unlikely due to the strong carbon-carbon bond between the methyl group and the cytosine.
2. There is evidence that the DNA repair machinery can be co-opted to remove a methylated base or the surrounding region. The excised base or region is then repaired with unmethylated cytosines replacing 5mCs.
3. 5mC oxidation-dependent active DNA demethylation. This follows from the observation that the TET enzymes can iteratively oxidate a _5mC_ -> _5hmC_ -> _5fC_ -> _5caC_ reaction. The removal of _5hmC_, _5fC_ or _5caC_ is biochemically "easier" than the removal of _5mC_ and could occur via a more efficient form of replication-dependent demethylation, direct removal of the oxidized methyl group or through the DNA repair machinery.

One question raised by point (3) is whether _5hmC_, _5fC_ and _5caC_ are simply intermediate products in an active demethylation cycle or if they themselves are bona fide epigenetic marks. This is an active area of investigation.

\section{Assays for studying DNA methylation}\label{sec:dna_methylation_assays}

A challenge to measuring DNA methylation is that it is erased by standard molecular biology techniques, such as the _polymerase chain reaction_ (PCR) and bacterial cloning, and it is not revealed by DNA hybridization assays \cite{Laird:2010iv}. Therefore, almost all assays of DNA methylation require one of the following _pre-treatments_ of the DNA:

1. _Enzyme digestion_
2. _Affinity enrichment_
3. _Sodium bisulfite conversion_

Each of these pre-treatments can then be combined with a second step of amplification or hybridisation. For example,

1. Gel-based analysis
2. Sanger sequencing
3. Microarray hybridisation
4. Massively parallel sequencing

An exception to this classification scheme are a new class of assay that seek to directly 'read' whether a position is methylated or unmethylated without requiring a pre-treatment of the DNA. For example, both \cite{Laszlo:2013kf} and \cite{Schreiber:2013in} measure the change in current as a DNA molecule passes through a nanopore to infer whether a cytosine is methylated.

Almost all assays of DNA methylation measure a "population average" from a pool of hundreds or thousands of cells. For a diploid organism, this is an average over several different levels: the two DNA strands, the two homologous chromosomes within a diploid cell and the hundreds or thousand of cells used in the assay. Hundreds or thousands of cells are required in order to have sufficient material as input for the assay. Assays that require only a single cell as input do exist \cite[e.g.,]{Smallwood:2014kn, Guo:2013ih} but are still in development and not yet in widespread use.

The _resolution_ and _throughput_ of an assay are two key variables when choosing which to use for an experiment. The resolution of an assay is the scale on which DNA methylation can be measured[^resolution]. For example, a high resolution assay allows a researcher to quantify the level of DNA methylation at a single base whereas a low resolution assay might only allow for qualitative assessment (i.e. presence or absence) of DNA methylation at larger regions, such as CpG islands.

[^resolution]: Depending on the experiment and its aims, the resolution of an assay might instead be defined as the scale on which DNA methylation can be quantified or the scale at which allows inference to address a specific hypothesis.

The throughput of an assay can be quantified in two ways. The first is per-sample throughput, which is how many measurements of DNA methylation are made per-sample[^throughput1]. This is typically what people mean when they describe an assay as being 'high-throughput' or 'low-throughput' and is the definition I use in the title of my thesis. Depending on your definition of 'high', a high-througput assay will produce on the order of tens of thousands to billions of measurements per sample. The second definition of throughput is  related to cost, be it money or time, i.e., 'how many samples can I afford to analyse?''.

[^throughput1]: This might reasonably be argued as being a definition of resolution.

The choice of which assay to use for an experiment is a trade-off between resolution, per-sample throughput and per-cost throughput. Experiments that use an assay with high-resolution and high per-sample throughput generally have fewer samples than experiments using a lower resolution assay or an assay with lower per-sample throughput.

In this section I will describe each of the pre-treatments but will focus on the bisulfite-conversion assays. In particular, I will describe in detail the 'gold standard; assay of DNA methylation, _whole-genome bisulfite-sequencing_ (WGBS), that combines the sodium bisulfite conversion pre-treatment with massively parallel sequencing to produce whole-genome maps of DNA methylation at base pair resolution.

\subsection{Enzyme digestion assays}\label{sec:enzyme_digestion_assays}

_Restriction endonucleases_ are an important technique in molecular biology. These enzymes can preferentially 'cut', 'cleave' or 'digest' DNA at or near to particular sequence motifs. The motif at which a restriction enzyme cleaves DNA is called the _recognition motif_ or _restriction sequence_. The methylation of a position in the recognition motif can inhibit a restriction enzyme from cleaving the DNA. This can be used to design an assay to infer the methylation state of a DNA fragment.

For example, the recognition site of the restriction enzyme _HpaII_ is `CCGC`. However, _HpaII_ will only digest DNA when the second cytosine in the motif is unmethylated. The _HELP_ (_HpaII_ tiny fragment enrichment by ligation-mediated PCR) assay compares DNA digested by _HpaII_ to one digested with another restriction enzyme that has the same recognition motif but is methylation-insensitive (_MspI_) to identify _hypomethylated_ regions of a genome \cite{Khulan:2006gx}.

Assays based on restriction enzyme were some of the first developed for studying DNA methylation. They were initially developed for studying particular loci although they have been extended to genome-scale analysis approaches \cite{Laird:2010iv}. Restriction enzyme assays have a relatively low resolution.

\subsection{Affinity enrichment assays}\label{sec:affinity_enrichment_assays}

Affinity enrichment assays compare measurements between an 'enriched' version and an 'input' (control) version of the same sample to infer the presence or absence of DNA methylation. This may use antibody immunoprecipitation or methyl-binding proteins. Some examples of affinity enrichment assays for DNA methylation are the microarray-based MeDIP, mDIP and mCIP and their sequencing-based relatives, MeDIP-seq and mDIP-seq.

These are all low resolution assays since they are based on the enrichment of regional differences between the enriched and input samples. Furthermore, the bioinformatic analysis of data from these assays is complicated by the varying CpG density along the genome, which leads to different enrichment affinities for different regions of the genome. However, these assays are can provide a relatively cheap and efficient genome-wide assessment of DNA methylation \cite{Laird:2010iv}.

\subsection{Sodium bisulfite conversion assays}\label{sec:bs_assays}

In the 1980s, two research groups independently discovered that when DNA is treated with sodium bisulfite ($NaHSO_{3}$), unmethylated cytosines deaminate to uracils much faster than do methylated cytosines \cite{Shapiro:1970im, Hayatsu:1970jv}. The methylated cytosines are said to be 'protected' from conversion to uracils. This discovery led to the development of assays for studying studying cytosine methylation based on the pre-treatment of DNA with sodium bisulfite \cite{Frommer:1992wq, Clark:1994wc}, which are referred to as bisulfite-conversion assays.

When bisulfite-treated is amplified by PCR, the uracils are converted to thymines. These bisulfite-conversion assays are all based on the idea of comparing the sequence of the untreated DNA to the sequence of the bisulfite treated DNA to infer the methylation state of all cytosines in the sequence by whether or not they were converted to uracil/thymine following the bisulfite treatment (Figure \ref{fig:bisulfite-conversion}).

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/bisulfite-conversion.pdf}
\caption{The effect of bisulfite-treatment of DNA. The double-stranded DNA is denatured and each strand undergoes bisulfite-treatment. Methylated cytosines remain as cytosines while unmethylated cytosines become uraciles. These bisulfite-converted DNA strands then undergo PCR amplification which converts the uracils to thymines. Note that while there are four possible PCR products, some bisulfite-sequencing protocols do not sample PCR products from the \textit{CTOT} or \textit{CTOB} strands. \textit{OT} = original top strand; \textit{OB} = original bottom strand; \textit{CTOT} = complementary to the original top strand; \textit{CTOB} = complementary to the original bottom strand. This figure is adapted from \citet{Krueger:2012ks}.}
\label{fig:bisulfite-conversion}
\end{figure}

Initial experiments based on the sodium bisulfite pre-treatment of DNA used Sanger sequencing of cloned PCR products, a very laborious task, which restricted experiments to studying a limited number of short segments of DNA. Although subsequent enhancements in the automation of Sanger sequencing improved the throughput of these assays, it was never going to be able to deliver a cost-effective, genome-scale assay of DNA methylation. The development of hybridisation microarrays provided an alternative way of analysing bisulfite-treated DNA that could provide cheap, genome-wide measurements of DNA methylation.

Microarrays contain thousands, even millions, of short _oligonucleotide_ probes. Each probe is is designed to hybridise to a particular DNA sequence and emits a fluorescent signal that can be measured to infer the strength of the hybridisation. Therefore, an (idealised) way to analyse DNA methylation with a microarray is to hybridise bisulfite-converted DNA to a microarray that contains probes for both the methylated and unmethylated versions of all sequences of interest. The relative methylation of each sequence can be inferred from the relative intensities of the 'methylated probe' to the 'unmethylated probe'. Such an idealised experiment brushes over many complications including \cite{Laird:2010iv}:

- The reduced complexity of bisulfite-converted DNA (from a 4-base alphabet to a largely 3-base alphabet) leads to decreased hybridisation specificity
- Sequences containing multiple cytosines require multiple probe versions in order to assay all possible methylation patterns
- This approach requires the design of organism-specific microarrays

The Illumina Infinium HumanMethylation450 BeadChip (_Illumina 450k_) provides a modern implementation of this type of assay for studying human DNA. This array assays $482,422$ cytosines, $99.3\%$ of these CpGs, across a wide variety of genomic features \cite{Stirzaker:2014cu}. There are well-established methods and software for analysing Illumina 450k data, which were recently reviewed in \cite{Dedeurwaerder:2014bt}.

Genomics research was revolutionised by the development of cheap high-throughput sequencing technology, and the study of DNA methylation was no exception. In 2008, two papers were published describing methods for whole-genome shotgun sequencing of bisulfite-converted DNA using the nascent Illumina sequencing technology \cite{Cokus:2008fc, Lister:2008bh}. Whole-genome bisulfite-sequencing is the gold-standard for analysing genome-wide DNA methylation data.

\cite{Cokus:2008fc} termed their approach _BS-seq_ while \cite{Lister:2008bh} called theirs _methylC-seq_. From a bioinformatics perspective, the main difference is that the _BS-seq_ protocol produces sequencing reads from four bisulfite-converted DNA strands - the original top strand (_OT_), the complementary strand to the original top strand (_CTOT_), the original bottom strand (_OB_) and the complementary strand to the original bottom strand (_CTOB_) - which require mapping to four different _in silico_ converted reference genomes, followed by a merge of the alignment results. In contrast, the _methylC-seq_ protocol only produces sequencing reads from two bisulfite-converted strands - _OT_ and _OB_ - and so only requires mapping to two different _in silico_ converted reference genomes, followed by a merge. The methylC-seq protocol is now the standard whole-genome bisulfite-sequencing protocol, due in part to the simpler bioinformatics analysis.

Whole-genome bisulfite-sequencing remains an expensive assay, which limits its use in studies involving large numbers of samples. Furthermore, approximately $40\%$ of reads will not contain any CpGs, which, depending on the experiment, might be considered a gross waste of money. Several assays have been developed to perform targeted high-throughput bisulfite-sequencing. The targeted nature of these assays are their obvious advantage and disadvantage - only a subset of the genome needs be sequenced but you only obtain information about methylation in those regions. Depending on the experiment, this tradeoff may be worthwhile, and these assays have been successfully used in a number of studies such as the BLUEPRINT Epigenome project ([http://www.blueprint-epigenome.eu/](http://www.blueprint-epigenome.eu/)) and the NIH Roadmap Epigenomics Mapping Consortium ([http://www.roadmapepigenomics.org/](http://www.roadmapepigenomics.org))

Some examples of these targeted bisulfite-sequencing are:
- Reduced representation bisulfite-sequencing (_RRBS_, \cite{Meissner:2005eb}): RRBS uses restriction enzymes to first select regions of the genome with a high CpG density (based on the Msp-I cleavage motif), which are subsequently treated with sodium-bisulfite and sequenced.
- NimbleGen's SeqCap Epi Enrichment System and Agilent's SureSelectxt Human Methyl-Seq: Both of these commercial products use DNA hybridisation to enrich the sequencing library for pre-defined regions of interest. This enriched library is then bisulfite-converted and sequenced.

One final class of bisulfite-conversion assays does not use microarrays or high-throughput sequencing. Sequenom's EpiTYPER \cite{Ehrich:2006tt} uses mass spectrometry to analyse DNA methylation from bisulfite-converted DNA. This platform can provide quantitative measurements of CpG methylation across hundreds of loci and multiple samples and may be used to validate findings discovered using other platforms \cite{Laird:2010iv}.

\subsubsection{Pros and cons of bisulfite-conversion assays}\label{sec:pros_and_cons_of_bs_assays
}

Bisulfite-conversion assays are considered the 'gold standard' for studying DNA methylation since cytosine methylation can be detected at single-base resolution \cite{Stirzaker:2014cu}. In fact, single-molecule, base pair resolution is even possible for short DNA sequences when bisulfite-treated DNA is analysed with sequencing[^microarrays].

[^microarrays]: Microarray hybridisation can provide base pair resolution but not single molecule resolution. The signal from a microarray-based experiment a sample-wide average since, for a given locus, the signal is relative to the proportion of DNA fragments in the sample methylated at that locus.

Almost all bisulfite-based assays require a considerable amount of DNA, e.g., $1-5$ $\mu$g for whole-genome bisulfite-sequencing, which is extracted from a population of cells. However, the minimal amount of DNA is being reduced with each technological advance. Recently, single-cell bisulfite-sequencing assays have been reported \cite{Guo:2013ih, Smallwood:2014kn}, although they are not yet a 'commercial' sequencing assay. Furthermore, these are not yet proper genome-wide assays. The technique of \cite{Guo:2013ih} is adapted from RRBS and so only assays a small percentage of cytosines in the genome. And the technique of \cite{Smallwood:2014kn}, while intended as a genome-wide assay, can reportedly only "accurately measure DNA methylation at up to 48.4% of cpG sites" \cite{Smallwood:2014kn}.

A recently discovered disadvantage is that bisulfite-conversion assays are unable to distinguish 5hmC from 5mC because hydroxymethyl similarly protects a cytosine from deamination to uracil. In effect, the detection of 5mC, and all subsequent inference, is confounded with that of 5hmC. For most experiments this isn't much of a problem - most cells have very low levels of 5hmC and so there is little confounding - however, in certain experiments this needs a more careful approach. To address this issue, \cite{Booth:2012fh} developed _oxidative bisulfite-sequencing_ (oxBS-seq) and \cite{Yu:2012ej} developed _Tet-assisted bisulfite sequencing_ (TAB-seq) for separate 5mC and 5hmC detection.

oxBS-seq specifically measures 5mC. The input DNA is oxidated by potassium perruthenate ($RKRuO_{4}$), which converts 5hmC to 5fC, prior to bisulfite-treatment. Only 5mC is protected from conversion during the bisulfite-treatment, which effectively means that only 5mC remains to be detected at the sequencing stage. The level of 5hmC can be estimated by performing traditional bisulfite-sequencing and then 'subtracting' the oxBS-seq signal (5mC) from the bisulfite-sequencing signal (5mC + 5hmC).

TAB-seq takes the opposite approach to oxBS-seq by specifically measuring 5hmC. The input DNA is treated with a $\beta$-glucosyltransferase, which converts 5hmC to $\beta$-glucosyl-5-hydroxymethylcytosine (5gmC), followed by TET oxidation. Only 5gmC is protected from TET oxidation, which effectively means that only 5hmC remains to be detected at the sequencing stage. The level of 5mC can be estimated by performing traditional bisulfite-sequencing and then 'subtracting' the TAB-seq signal (5hmC) from the bisulfite-sequencing signal (5mC + 5hmC).

There is also potential confounding of 5C, 5mC and 5hmC levels with 5fC and 5caC, although this has received less attention since 5fC and 5caC are believed to exist at far lower quantities than 5mC and 5hmC. Nonetheless, sequencing assays of 5fC and 5caC exist based on the idea of treating the DNA with a chemical and performing 'signal subtraction' with traditional bisulfite-sequencing or another assay \cite{Wu:2014jj}.

Another disadvantage of bisulfite-conversion assays is that they require knowledge of the underlying DNA sequence in order to infer the methylation states of cytosines. This requires either a parallel experiment to sequence the target region(s) or reliance on a reference genome. When relying on a reference genome, the inference of the methylation state can be confounded by the DNA sequence of the sample \cite{Liu:2012ge}. Figure \ref{fig:bisulfite_CT_SNP.pdf} illustrates such an example.

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/bisulfite_CT_SNP.pdf}
\caption{Sequence variation between the reference genome and the sample's genome can result in incorrect inferrence about the methylation state of the sample's genome. The locus in orange is a cytosine in the refererence genome but a thymine in the sample's genome. Because the read is compared against the reference genome it is incorrectly inferred to be an unmethylated cytosine. The locus in purple is a thymine in the reference genome but an unmethylated cytosine in the sample's genome. Because the read is compared against the reference genome it is incorrectly inferred to be a thymine.}
\label{fig:bisulfite_CT_SNP.pdf}
\end{figure}

The bisulfite-treatment of DNA can introduce biases and other problems. Three examples are _PCR-bias_, _incomplete bisulfite conversion_ and _DNA degradation_. _PCR-bias_ is the difference in amplification efficiency of methylated and unmethylated versions of the same DNA sequence \cite{Warnecke:1997eh}. _Incomplete bisulfite-conversion_ leads to cytosines being incorrectly inferred as 5mC since they cytosines were not converted to uracils by the bisulfite-treatment. Finally, treating DNA with sodium bisulfite damages the DNA, resulting in the fragmentation of long molecules. This limits the size of the fragments that can be studied using bisulfite-conversion assays to $\approx 500$ bp \cite{Ecker:2010et}

\subsection{DNA kinetics assays}\label{sec:dna_kinetics_assays}

Ideally, an assay of DNA methylation could naturally 'read' methylated cytosines as a distinct signal from unmethylated cytosines. This idea forms the basis of assays using the kinetics of DNA to infer the presence of DNA methylation and other DNA modifications. While very exciting, these assays do not yet scale to studying genome-wide DNA methylation levels in mammalian-sized genomes.

The only commercially available assay in this class is the Pacific Biosciences SMRT technology \cite{Flusberg:2010hra}. It infers the presence of DNA modifications by comparing the time it takes to 'read' the modified form of a nucleotide, such as 5mC, to the time it takes to read its unmodified form.

\section{Summary}

DNA methylation is an epigenetic modification with important roles in many biological systems. In mammals, CpGs are frequently methylated while non-CpG sites are less frequently methylated. Most unmethylated CpGs are found in CpG islands, which are important regulatory elements in the genome.

Whole-genome bisulfite-sequencing provides single-nucleotide resolution data and is the gold-standard assay for studying DNA methylation. Methods for analysing whole-genome bisulfite-sequencing data are the focus of my thesis. The following chapter describes the bioinformatics analysis of a bisulfite-sequencing experiment, including its many statistical and computational challenges.