### 2014-12-03 Rmarkdown/wgbs_statistical_analysis.Rmd
Described BiSeq and data-driven DMRs

---
title: "Statistical analysis of whole-genome bisulfite-sequencing data"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

\chapter{Statistical analysis of whole-genome bisulfite-sequencing data}\label{chap:wgbs_statistical_analysis}

\section{Chapter overview}\label{sec:statmodel_chapter_overview}

I set out a statistical framework for analysing bisulfite-sequencing data. I begin by explaining the various levels of stochasticity and then define the mathematical notation that I use throughout my thesis. I describe key variables, common estimators of these and their statistical properties. The ideas here are simple, although when extended to their full generality there are several subtleties that must be kept in mind.

As is clear from the length of Chapter \ref{chap:wgbs_bioinformatics_analysis} alone, there is a considerable amount of bioinformatics work before we even consider analysing the data to address the biological question of interest. I call these the 'downstream' analyses, and there are many that might be considered. Using the above statistical framework, I explain the statistical question of identifying differential methylation. I also discuss other downstream analyses.

- Chapter overview
  - One sample
    - Pre-sequencing
    - Post-sequencing
    - Some complications for a single sample
      - What is $\mathcal{I}$?
      - Genetic heterozygosity at a methylation locus
  - Multiple samples
    - Some complications with multiple samples
      - What is $\mathcal{I}$?
  - Parameter estimation
    - Estimating $M$, $U$
    - Estimating $B$
      - Beta-binomial models
      - Smoothing
    - Statistical properties of $\beta$-values
      - Empirical distributions
      * Write a function(s) to do the following and apply it to the EPISCOPE, Lister, Seisenberger and Ziller datasets:
        - Genome-wide distribution of CpG $\beta$-values
        * Kernel density estimates
        * Stratify by CGI status, genic, intragenic, repeats
        * Correlation across strands
        * Effect of sequencing coverage threshold (`min_cov`)
      - What is the genome-wide methylation level?
        * Various figures get bandied about; what's a good summary?
        - A criticism of ``identifying methylcytosines''
      - Correlations
      - Biases
      - Transformations
  - Downstream analyses
    - Differential methylation
      - Experimental design considerations
      - DMCs: Identify $\bar{B_{i}^{(1)}} \neq \bar{B_{i}^{(2)}}$
        - One-group experiments
        - Two-group experiments
        - Multi-group experiments
      - DMRs: Identify runs of DMCs
        - A priori vs. data-driven regions
        - One-group experiments
        - Two-group experiments
        - Multi-group experiments
    - Differential variability
      - DVCs: Identify $\sigma(B_{i}^{(1)}) \neq \sigma(B_{i}^{(2)})$
      - DVRs: Identify runs of DVCs
    - Allele-specific methylation
    - Methylation entropy
    - Epipolymorphism
    - Epiallele detection
    - Integrating DNA methylation data with other genomes data

\section{One sample}\label{sec:one_sample}

Even with only a single sample, there are several levels of variation to consider in a bisulfite-sequencing experiment. I find it convenient to separate these into:

1. Pre-sequencing sources of variation
2. Sequencing and post-sequencing sources of variation.

I describe these sources of variation, which are illustrated in Figure \ref{fig:bs-seq_protocol}.

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/bs-seq_protocol.pdf}
\caption{A schematic illustrating several sources of variation in a bisulfite-sequencing experiment. Each sample starts as a population of cells, with potentially different methylation patterns. Each DNA fragment is coloured by its originating cell, although this is unknown in practice. Illustrated are PCR amplication bias (unequal representation of DNA from each cell following PCR amplification), sampling variation (not all DNA fragments are sequenced), alignment errors (not all sequenced fragments are mapped or may be mapped to the wrong location, as is the case for the green fragment), and filtering during methylation calling (not reads are used, the blue read, and reads may have read-positions removed, the black reads).}
\label{fig:bs-seq_protocol}
\end{figure}

\subsection{Pre-sequencing}\label{sec:one_sample_pre_sequencing}

A methylation locus is a single cytosine, that is, a CpG, CHG or CHH. The set of these loci is labelled $\mathcal{I} = \{pos_{i}: i = 1, \ldots, N_{loci} \}$, where $pos_{i}$ is the genomic co-ordinates of the $i^{th}$ locus, e.g., chr1:+:666. Generally, the number of methylation loci, $N_{loci}$, is not known, although estimates may be made based from the sample or a reference genome. Not knowing the exact value of $N_{loci}$ is no great concern.

It is sometimes convenient to refer to loci by the subscript $i$ rather than by $pos_{i}$. It is important to remember that the number of base pairs between the $i^{th}$ and $(i + 1)^{th}$ methylation loci varies along the genome. Also, depending on whether the data are stranded, a pair of loci may be on opposite strands. In a small number of instances the $i^{th}$ and $(i + 1)^{th}$ methylation loci are on separate chromosomes, e.g., the last CpG on chromosome 1 and the first CpG on chromosome 2.

The methylation state of a locus can vary within a sample due to cell-to-cell variability of the methylation state. As discussed, a sample in a bisulfite-sequencing experiment contains DNA that is extracted from hundreds or thousands of cells and each cell can have a slightly different methylation profile. Furthermore, within a diploid cell there are two copies of each chromosome, and therefore two copies of each methylation locus, and these two copies can have different methylation states. Therefore, it is also necessary to consider not just the genomic co-ordinates of the locus but from which DNA fragment the methylation state originated.

Suppose that in the pool of DNA fragments for the sample that there are $H_{i}$ fragments containing the $i^{th}$ methylation locus. In general, $H_{i}$ is unknown and will vary from locus to locus within a sample[^H_i]. Note that the value of $H_{i}$ is determined following the library preparation, including PCR amplification of the DNA; therefore, it can give a grossly distorted picture of the true representation of the cells. I denote by $\mathcal{H}_{i}$ the set of all fragments containing the $i^{th}$ locus.

[^H_i]: Knowing $H_{i}$ would require knowing: (1) the number of cells used as input (which might only be known to within an order of magnitude), (2) the ploidy of each cell (generally known) and (3) the number of PCR cycles (generally known). But the real problem is that none of the steps in creating the pool of DNA fragments is perfect. In particular, PCR introduces biases -- some molecules are preferentially amplified while others 'drop out'. So even if we knew (1), (2) and (3) we cannot simply multiply these together to compute $H_{i}$, although this might at least give us a rough estimate.

Although we do not know the number of fragments in the pool, we can define (and measure) the methylation state of a locus on a single DNA fragment. I denote by the indicator random variable, $Z_{h, i}$, the methylation state of $i^{th}$ methylation locus on the $h^{th}$ DNA fragment:

\begin{equation*}
Z_{h, i} = \left\{
  \begin{array}{l l}
    1 & \quad \text{if methylated on the } h^{th} \text{ fragment}\\
    0 & \quad \text{if unmethylated on the } h^{th} \text{ fragment}
  \end{array} \right.
\end{equation*}

By summing over the number of fragments containing the $i^{th}$ locus, we obtain the number of fragments that are methylated at the $i^{th}$ locus ($M_{i}$) and unmethylated at the $i^{th}$ locus ($U_{i}$):

\begin{align*}
  M_{i} &= \sum_{h = 1}^{H = H_{i}} Z_{h, i} = |\{Z: Z \in \mathcal{H}_{i}, Z = 1 \}| \\
  U_{i} &= \sum_{h = 1}^{H = H_{i}} (1 - Z_{h, i}) = |\{Z: Z \in \mathcal{H}_{i}, Z = 0 \}|
\end{align*}

From these we can compute the proportion of fragments that methylated at the $i^{th}$ locus:

\begin{equation*}
  B_{i} = \frac{M_{i}}{M_{i} + U_{i}}
\end{equation*}

The above definitions can be extended from individual methylation loci, 1-tuples, to m-tuples. Mathematically, an m-tuple is denoted by a sequence of methylation loci, $(i, i + 1, \ldots, i + m - 1)$.

I denote by the vector of indicator random variables, $Z_{h, (i, i + 1, \ldots, i + \text{m} - 1)}$, the methylation pattern of the m-tuple $(i, i + 1, \ldots, i + m - 1)$ on the $h^{th}$ DNA fragment containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$:

\begin{equation*}
Z_{h, i} = \left\{
  \begin{array}{l l}
  (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus on the } h^{th} \text{ fragment}\\
  (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th}) \text{ locus and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus on the } h^{th} \text{ fragment} \\
  \vdots \\
  (1, 1, \ldots, 1) & \quad \text{if methylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus on the } h^{th} \text{ fragment}
  \end{array} \right.
  \end{equation*}

$\mathcal{H}_{(i, i + 1, i + m - 1)}$ denotes the set of all fragments containing the m-tuple $(i, i + 1, i + m - 1)$ and $\mathcal{R}_{(i, i + 1, i + m - 1)}$ denotes the set of all reads containing the m-tuple $(i, i + 1, i + m - 1)$.

There are $2^{\text{m}}$ possible methylation patterns at an m-tuple. Rather than describe a methylation pattern by a m-vector of zeros and ones, I also write these using $U$ and $M$; for example, the possible methylation patterns at a 2-tuple are $MM, MU, UM$ and $UU$.

Analogous to the definition of $M_{i}$ and $U_{i}$ for 1-tuples ($\text{m} = 1$), we have when $\text{m} = 2$:

\begin{align*}
  MM_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (1, 1)\}| \\
  MU_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (1, 0)\}| \\
  UM_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (0, 1)\}| \\
  UU_{(i, i + 1)} &= |\{Z: Z \in \mathcal{H}_{(i, i + 1)}, Z = (0, 0)\}| \\
  \end{align*}

We could extend the $B_{i}$ values to m-tuples, although the intuitive interpretation of these as the average methylation level is lost. Instead, it reflects the relative frequencies of each methylation pattern. Here are the definitions for $\text{m} = 2$:

  \begin{align*}
  B_{(i, i + 1)}^{MM} &= \frac{MM_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}} \\
  B_{(i, i + 1)}^{MU} &= \frac{MU_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}}  \\
  B_{(i, i + 1)}^{UM} &= \frac{UM_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}}  \\
  B_{(i, i + 1)}^{UU} &= \frac{UU_{(i, i + 1)}}{MM_{(i, i + 1)} + MU_{(i, i + 1)} + UM_{(i, i + 1)} + UU_{(i, i + 1)}}
  \end{align*}

The definitions for $\text{m} > 2$ follow in the obvious manner.

Again, I emphasise that $H_{(i, i + 1, i + \text{m} - 1), j}, Z_{h, (i, i + 1, i + \text{m} - 1), j}, B_{(i, i + 1, i + \text{m} - 1), j}$ and the set of methylation patterns are unobservable. However, by sequencing the pools of DNA fragments we aim to estimate these variables.

\subsection{Post-sequencing}\label{sec:one_sample_post_sequencing}

In a whole-genome bisulfite-sequencing experiment we do not sequence every DNA fragment in the pool. Rather, sequencing can be thought of as sampling without replacement from the pool of DNA fragments. We have a large number(~$10^{10}$) of fragments in the pool and each methylation locus is only present on a small number of those fragments. Therefore, we can approximate this sampling by Poisson sampling, where the rate parameter for locus $i$ is proportional to the number of fragments in the pool and inversely proportional to $H_{i}$.

At this point I ignore reads that do not contain any methylation loci as these are not relevant to this discussion and I make three simplifying assumptions:

1. Sequencing is performed without error.
2. Read mapping is perfect.
3. We perform single-end sequencing.

The effect of the first two assumptions are discussed in Section \ref{sec:mapping_and_post-processing}. The effect of the third assumption is minor. When using single-end sequencing, the methylation loci from a single read will always form a positively ordered sequence without gaps, i.e., $(i, i + 1, i + 2)$ and not $(i, i - 1, i - 2)$ nor $(i, i + 1, i + 3)$. However, when using paired-end sequencing the methylation loci from a paired-end read will still be an ordered sequence but one of the following may occur (__FIGURE__):

1. There may be gaps due to the insert size being longer than the sum of the read lengths, e.g. $(i, i + 1, i + 3, i + 4))$. In effect, we have missing data for any intervening methylation loci; the $(i + 2)^{th}$ loci in this example.
2. Loci may be measured twice if the insert size is less than the sum of the read lengths, e.g. read_1 gives us $(i, i + 1)$ and read_2 gives us $(i + 1, i + 2, i + 3)$. In this example we must use only one of read_1 or read_2 as the measurement of the $(i + 1)^{th}$ locus because otherwise we are 'double-counting'.

Each read measures the methylation state of one or more loci from a single DNA fragment. I denote by $\mathcal{R}_i$ the set of all reads containing the $i^{th}$ locus. Therefore, the number of reads containing the $i^{th}$ locus is $|\mathcal{R}_{i}|$, where $|\mathcal{R}_{i}| \leq H_{i}$ with strict inequality for almost all $i$.

A single read containing the $i^{th}$ locus is denoted $z: z \in \mathcal{R}_{i}$ and the observed methylation state is indicated by:

\begin{equation*}
z: z \in \mathcal{R}_{i} = \left\{
  \begin{array}{l l}
    1 & \quad \text{if methylated at the } i^{th} \text{ locus}\\
    0 & \quad \text{if unmethylated at the } i^{th} \text{ locus}
  \end{array} \right.
\end{equation*}

By summing over the number of reads containing the $i^{th}$ locus we obtain the number of reads that are methylated at the $i^{th}$ locus ($m_{i}$) and unmethylated at the $i^{th}$ locus ($u_{i}$):

\begin{align*}
  m_{i} &= \sum_{z: z \in \mathcal{R}_{i}} z \\
        &= |\{z: z \in \mathcal{R}_{i}, z = 1 \}| \\
  u_{i} &= \sum_{z: z \in \mathcal{R}_{i}} (1 - z) \\
      &= |\{z: z \in \mathcal{R}_{i}, z = 0 \}
\end{align*}

From these we can compute the proportion of reads that are methylated at the $i^{th}$ locus as:

\begin{equation*}
  \beta_{i} = \frac{m_{i}}{m_{i} + u_{i}}
\end{equation*}

These are the so-called $\beta$-values, which are commonly interpreted as an estimate of the proportion of cells in the sample that are methylated at the $i^{th}$ locus. I discuss this interpretation, and other estimators of the  methylation level at a locus, in Section \ref{sec:estimating_B}.

Again, these definitions can be extended from 1-tuples to m-tuples. The set of all reads containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$ is denoted by $\mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)}$. A single read containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$ is denoted by $z: z \in \mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)}$, and the observed methylation state is given by:

\begin{equation*}
z: z \in \mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)} = \left\{
  \begin{array}{l l}
    (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus}\\
    (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th}) \text{ locus and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus} \\
    \vdots \\
    (1, 1, \ldots, 1) & \quad \text{if methylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus}
      \end{array} \right.
\end{equation*}

Note that we do not know from which haplotype ($h$) each read came from, only that all methylation loci in the read came from the same DNA fragment.

By counting over the number of reads containing the m-tuple $(i, i + 1, \ldots, i + \text{m} - 1)$ we obtain the number of reads contain each methylation pattern at that m-tuple. Here are the definitions for $\text{m} = 2$:

\begin{align*}
  mm_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (1, 1)\}| \\
  mu_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (1, 0)\}| \\
  um_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (0, 1)\}| \\
  uu_{(i, i + 1)} &= |\{z: z \in \mathcal{R}_{(i, i + 1)}, z = (0, 0)\}|
\end{align*}

We can extend the $\beta_{i}$ values to m-tuples. Here are the definitions for $\text{m} = 2$:

\begin{align*}
  \beta_{(i, i + 1)}^{mm} &= \frac{mm_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}} \\
  \beta_{(i, i + 1)}^{mu} &= \frac{MU_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}} \\
  \beta_{(i, i + 1)}^{um} &= \frac{UM_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}} \\
  \beta_{(i, i + 1)}^{uu} &= \frac{UU_{(i, i + 1)}}{mm_{(i, i + 1)} + mu_{(i, i + 1)} + um_{(i, i + 1)} + uu_{(i, i + 1)}}
\end{align*}

The definitions for $\text{m} > 2$ follow in the obvious manner.

One final definition is the average methylation of each read, which is used by some authors, including \cite{Landan:2012kp}. For each read, $z \in \mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)}$, the average methylation of the read, $\zeta_{z}$, is defined as the proportion of methylation loci in the read that are methylated. Thus, $\zeta_{z} = 0, \frac{1}{m}, \frac{2}{m}, \ldots, 1$.

\subsection{Some complications for a single sample}\label{sec:one_sample_complications}

I now discuss some complications and how this framework might accommodate these issues in practice.

\subsubsection{What is $\mathcal{I}$?}\label{sec:what_is_I}

As mentioned in Chapter \ref{chap:wgbs_bioinformatics_analysis}, studies using bisulfite-conversion assays rely on either a reference genome or, less commonly, separate DNA sequencing of the sample that is assayed. Different analysis strategies lead to different definitions of $\mathcal{I}$, which are approximations to the 'true' set of methylation loci in the sample, $\mathcal{I}^{truth}$. Listed here are definitions of $\mathcal{I}$ from least closely matching to most closely matching $\mathcal{I}^{truth}$:

1. $\mathcal{I}^{ref}$: Defined by the set of methylation loci in the reference genome. This ignore all genetic variation between the sample and the reference.
2. $\mathcal{I}^{refFilter}$: Defined by filtering out problematic loci from $\mathcal{I}^{ref}$. A conservative approach that removes many sites of genetic variation between the sample and the reference as well as sites that do not display genetic variation between the sample and the reference. This approach cannot identify sample-specific methylation loci.
3. $\mathcal{I}^{Bis-SNP}$: Defined by calling genetic variants from the bisulfite-sequencing data using `Bis-SNP` \cite{Liu:2012ge}. Identifies sample-specific methylation loci and removes reference-specific methylation loci. This is the best approach if only bisulfite-sequencing data are available.
4. $\mathcal{I}^{WGS}$: Defined by identifying all methylation loci from _whole-genome sequencing_ (WGS) of the sample's genome. The gold standard. All methylation loci are defined with respect to the sample's genome. The only differences between $\mathcal{I}^{WGS}$ and $\mathcal{I}^{truth}$ are due to sequencing errors, incomplete sequencing coverage of the sample's genome and variant calling errors from the WGS data.

\subsubsection{Genetic heterozygosity at a methylation locus}\label{sec:heterozygosity}

The genome of a diploid organism has sites that are heterozygous due to differences between the maternal and paternal chromosomes. Such heterozygous loci are sometimes also methylation loci. For example, a locus where the maternal chromosome is a CpG and the paternal chromosome is an ApG. In effect, the maternal and paternal chromosomes within the sample have different $\mathcal{I}^{truth}$.

The number of these genetically heterozygous methylation loci is often small enough not to worry about. However, in some studies, such as those of allele-specific methylation, these loci can be very important and might be identified by calling heterozygous genetic variants using `Bis-SNP` or from WGS of the sample. In practice, the existence of such loci is often ignored.

\section{Multiple samples}\label{sec:multiple_samples}

From a purely notational perspective, the move from a single sample to multiple samples simply requires an additional subscript, $j = 1, \ldots, n$, where $n$ is the number of samples. This defines the three levels in the hierarchy of a typical experiment -- DNA fragments ($h$), methylation loci ($i$) and samples ($j$). For example, $\mathcal{I}_{j}$ is the set of methylation loci in the $j^{th}$ sample and $\beta_{i, j}$ is the beta value for the $i^{th}$ locus in the $j^{th}$ sample.

A fourth level is how the sample's relate to one another, such as the treatment group of each sample. This fourth level might be defined up-front, such as in a designed experiment looking for differences in methylation between tumour and normal tissue. Conversely, the aim of the experiment might be to _discover_ this grouping, such as in a clustering analysis.

We can define this fourth level using a design matrix $X = [X_{j}]$. For example, in a two-group experiment $X_{j} = 1$ if the sample is from group $1$ and $X_{j} = 0$ if the sample is from group $2$. We can also include covariates in the standard way by allowing $X_{j}$ to be a row vector, $X_{j} = (x_{1, j}, \ldots, x_{P, j})$, where $x_{p, j}$ encodes the information on the $p^{th}$ covariate for the $j^{th}$ sample.

\subsection{Some complications with multiple samples}\label{sec{multi_sample_complications}}

In addition to the complications of the Section \ref{sec:one_sample_complications}, we now have sample-to-sample variation that must be addressed by this framework.

\subsubsection{What is $\mathcal{I}$?}

Each sample has their own set of methylation loci, that is, $\mathcal{I}_{j}$ differs between $j$. Furthermore, sequencing coverage varies from sample-to-sample. This means that even if the samples have exactly the same $\mathcal{I}_{j}$, i.e., the samples are genetically identical, each sample will have a different set of loci with 'sufficient' sequencing coverage. Loci without sufficient sequencing coverage are effectively missing data.

In practice, we might choose to study $\mathcal{I}^{common} = \cap_{j} \mathcal{I}_{j}$ or some other intersection of the $\mathcal{I}_{j}$, such as all methylation loci present in at least some fraction of the $n$ samples.

A conservative analysis might only analyse those loci where at least some fraction of the $n$ samples have sufficient sequencing coverage. A less conservative analysis might try to impute the missing values based on methylation levels at neighbouring loci.

\section{Parameter estimation}\label{sec:parameter_estimation}

The main parameter of interest for each sample is the vector of methylation levels for each locus, $\bm{B_{j}}$. To do so it is necessary to estimate $M$ and $U$. In this section I review various methods for estimating these key parameters. When necessary, I have 'translated' the original work into my notation to make these methods more readily comparable. I have suppressed $j$ subscript when referring to a single sample.

\subsection{Estimating $M$, $U$}\label{sec:estimating_m_and_u}

It is rare to need direct estimates of $M_{i}$ or $U_{i}$. In order to estimate $M_{i}$ and $U_{i}$, the absolute number of methylated and unmethylated DNA fragments at the $i^{th}$ locus, we would also require an estimate of $|\mathcal{H}_{i}|$, the number of DNA fragments containing the $i^{th}$ locus. Rather, we are generally interested in estimating the proportion of reads that are methylated, $B_{i} = \frac{M_{i}}{M_{i} + U_{i}}$.

\subsection{Estimating $B$}\label{sec:estimating_B}

The simplest estimator of $B_{i}$ is $\beta_{i} = \frac{m_{i}}{m_{i} + u_{i}}$, which has been widely used \citep[e.g.,]{Cokus:2008fc, Lister:2008bh, Lister:2009hy, Lister:2011kg}). The values of $m_{i}$ and $u_{i}$ are obtained by methylation calling (see Section \ref{sec:methylation_calling}).

Under a Binomial model for the number of methylated reads at the $i^{th}$ locus in the $j^{th}$ sample, $M_{i, j} | n_{i, j} \eqd Binomial(n_{i, j}, B_{i, j})$, $\beta_{i, j}$ is the maximum likelihood estimator of $B_{i, j}$.

Recently, more sophisticated methods have been proposed to estimate or model the average methylation level. These methods, which are based on $m_{i, j}$ and $u_{i, j}$, include beta-binomial models \citet{Feng:2014iq, Sun:2014fk, Dolzhenko:2014bo}, and smoothing-methods \cite{Hansen:2011gu, Hansen:2012gr, Hebestreit:2013ko}.

\subsubsection{Beta-binomial models}\label{sec:beta-binomial_models}

Several papers have proposed the beta-binomial distribution since, as noted by \cite{Dolzhenko:2014bo}, it is "a natural model for describing methylation levels of an individual site across replicates" . The 'beta' component of the distribution models the underlying methylation level, $B_{i, j}$, while the 'binomial' component models the sampling of reads by sequencing. Another way to think of this is that the 'beta' component models the biological variability of the data, while the 'binomial' component models the sampling variability of sequencing. This separation of biological and technical variability has proven successful in detecting differential gene expression from RNA-seq data. For example, the `edgeR` software \cite{Robinson:2010cw} uses the negative binomial distribution, which can be thought of as a gamma-poisson mixture distribution, to account for both the biological and sampling variability.

An attractive feature of the beta-binomial distribution is that it can be motivated by, and analysed with, Bayesian methods, including empirical Bayes methods, or frequentist techniques such as maximum likelihood. For example, the software `DSS` \cite{Feng:2014iq} and `MOABS` \citet{Sun:2014fk} both use the beta-binomial distribution in an empirical Bayes analysis of differential methylation from bisulfite-sequencing data. In contrast, `RADmeth` \cite{Dolzhenko:2014bo} uses the beta-binomial model in a maximum likelihood framework to address the same problem.

\subsubsection{Smoothing $\beta$-values}\label{sec:smoothing_beta-values}

`BSmooth`, published in \citet{Hansen:2011gu, Hansen:2012gr} and available in the R/Bioconductor package `bsseq`, and `BiSeq`, published in \citet{Hebestreit:2013ko} and available in the R/Bioconductor package `BiSeq`, take a different approach to getting improved estimates of the $B_{i, j}$. Both `bsseq` and `BiSeq` use statistical smoothing of the 'raw' $\beta_{i, j} = \frac{m_{i, j}}{m_{i, j} + u_{i, j}}$. Smoothing is motivated and justified by the fact that the $B_{i, j}$ are spatially correlated within a sample, i.e., because of the presence of co-methylation (discussed in Chapters \ref{chap:co-methylation} and \ref{chap:co-methylation_review}).

Smoothing is particularly powerful for loci with low sequencing coverage, where the denominator $m_{i, j} + u_{i, j}$ is small and the corresponding standard error of $\beta_{i, j}$ is large. The smoothed $\beta$-values, rather than the raw $\beta$-values, are then generally used in all downstream analyses.

Both `bsseq` and `BiSeq` use a binomial local likelihood smoother. In each case this smoother was chosen because `BSmooth` and `BiSeq` model the number of methylated reads at the $i^{th}$ locus in the $j^{th}$ sample by ($M_{i, j} | n_{i, j} \eqd Binom(n_{i, j}, B_{i, j})$. The smoothing is 'local' to exploit co-methylation, which is a local phenomenon.

In both `bsseq` and `BiSeq` the raw $\beta$-values are weighted according to the binomial likelihood and a kernel function. The binomial likelihood weights points inversely to their standard error, $se(\beta_{i, j})$, and the kernel gives greater weight to those $\beta_{i, j}$ near the centre of the window. \citet{Lacey:2013iy} note that loci with very high sequencing coverage will strongly influence the smoother, potentially biasing estimates at neighbouring loci with lower coverage.

`bsseq` assumes that for each sample that the underlying methylation level, $B_{i, j}$, is a smoothly varying function of the position in the genome, $i$. In contrast, `BiSeq` first creates clusters of CpGs and only assumes that the underlying methylation level is smooth at positions within each cluster.

Whenever smoothing is used, a key parameter is the bandwidth, which is the size of the window in which observations are included at each iteration of the smoother. `bsseq` uses a much larger window size than `BiSeq`; the default window size in `bsseq` is one that contains at least 70 CpGs and is at least 2000kb wide, whereas the default window size in `BiSeq` is 80bp, regardless of CpG-density. This is due to `BiSeq` being developed for RRBS data, which has a high CpG-density per window, whereas `bsseq` was developed for whole-genome data, which has a more variable, and lower on average, CpG density per window.

Another 'parameter' choice when smoothing is the choice of kernel, although this is generally less important than the choice of bandwidth. `bsseq` uses a tricube kernel and `BiSeq` uses a triangular kernel.

\citet{Hebestreit:2013ko} and \citet{Lacey:2013iy} compare the smoothing results of `BiSeq` to `bsseq`. Both \citet{Hebestreit:2013ko} and \citet{Lacey:2013iy} provide instances where they claim `BiSeq` gives more 'reasonable' smoothed values than `bsseq`, but the comparison is a little unfair since both use RRBS data, for which `bsseq` is not designed[^rrbs_sim].

[^rrbs_sim]: Both \citet{Hebestreit:2013ko} and \citet{Lacey:2013iy} altered the default `bsseq` parameters to try to make them comparable to `BiSeq`. \citet{Hebestreit:2013ko} changed the default minimum window size to 80bp but still required at least 20 CpGs per window. \citet{Lacey:2013iy} kept the default minimum window size of 2000 bp but reduced the minimum number of CpGs per window to 50 from the default of 70.

\section{Statistical properties of $\beta$-values}\label{sec:beta}

Under the binomial model, ($M_{i, j} | n_{i, j}) \eqd Bin(n_{i, j}, B_{i, j})$, $\beta_{i, j} = \frac{m_{i, j}}{m_{i, j} + u_{i, j}}$ is an unbiased estimator of $B_{i, j}$ with standard error $se(\beta_{i, j}) = \sqrt(\frac{\beta_{i, j}(1 - \beta_{i, j})}{n_{i, j}})$ \citep{Hansen:2012gr}. The natural interpretation of $\beta_{i, j}$ is then as an estimator of the average level of methylation at the $i^{th}$ locus in the $j^{th}$ sample. In this section I discuss this interpretation and statistical properties of this estimator.

\subsection{Empirical distributions}

- TODO: Lots of figures and discussion. See chapter plan.

In a study involving multiple samples, the set of $\beta$-values can be summarised as a matrix where each row is a locus and each column is a sample. Some values will be missing, either because there was insufficient sequencing coverage to estimate a $\beta$-value or because that locus is not a cytosine for that sample. This matrix might be visualised to learn about the distribution of methylation levels, either row-wise, to learn about the variability across samples, or column-wise, to learn about the variability within samples.

\subsubsection{Genome-wide distribution of $\beta$-values}

Restricting our attention to CpGs, Figure \ref{fig:EPISCOPE_beta_distribution} shows the genome-wide distributions of $\beta$-values, that is, the column-wise summaries, for each EPISCOPE sample. What is immediately clear is that these distributions are bimodal. Figure \ref{fig:EPISCOPE_beta_distribution_CGI_stratified} shows that this bimodality is largely driven by the fact that CpGs in CpG islands are mostly unmethylated whereas those outside of CpG islands are generally methylated. Intermediate $\beta$-values, those between $0.2-0.8$, are evidence of heterogeneity.

Both Figures \ref{fig:EPISCOPE_beta_distribution} and  \ref{fig:EPISCOPE_beta_distribution_CGI_stratified} also reveal that the genome-wide distributions of $\beta$-values for a given tissue are very similar between the three individuals. The three adipose-derived tissues, _SA_, _VA_ and _VAT_, have similar genome-wide profiles while the blood-derived tissue, _BUF_, has more intermediate $\beta$-values, potentially indicating more cellular heterogeneity.

```{r EPISCOPE_genome_wide_beta-value_distributions, eval = FALSE, echo = FALSE}
# Run on one of the unices
setwd("/wehisan/home/allstaff/h/hickey/methylation_m-tuples/analyses/processed_data/EPISCOPE")
library(MethylationTuples)
library(ggplot2)
library(dplyr)
library(tidyr)
thesis_theme <- theme_classic(base_size = 20)
thesis_theme <- theme_bw(base_size = 20)

x <- readRDS("EPISCOPE_1_tuples_strand_collapsed.rds")

cgi <- read.table("~/methylation_m-tuples/CGI/model-based-cpg-islands-hg19.txt",
                  header = TRUE, stringsAsFactors = FALSE)
cgi <- GRanges(seqnames = cgi$chr, IRanges(cgi$start, cgi$end))
cgi_status <- overlapsAny(x, cgi)

beta <- methLevel(x, min_cov = 10L)

cn <- colnames(beta)
beta <- data.frame(Individual = rep(substr(cn, 1, 3), each = nrow(beta)),
                   Tissue = rep(substr(cn, 4, 6), each = nrow(beta)),
                   beta = as.vector(beta),
                   CGI = rep(cgi_status, times = ncol(beta)),
                   stringsAsFactors = FALSE)
beta <- tbl_df(beta)

#a <- beta[sample(nrow(beta), 100000), ]

g <- ggplot(aes(x = beta), data = beta) +
  geom_density(size = 1.5) +
  facet_grid(Tissue ~ Individual, labeller = label_both) +
  thesis_theme + scale_colour_brewer(palette = "Dark2") +
  ggtitle(expression(paste("EPISCOPE: Distribution of ", beta, "-values", sep = ""))) +
  xlab(expression(paste(beta, "-values", sep = "")))
ggsave("EPISCOPE_beta_distribution.pdf", g, height = 9, width = 16)

g <- ggplot(aes(x = beta), data = beta) +
  geom_density(aes(colour = CGI), size = 2) +
  facet_grid(Tissue ~ Individual, labeller = label_both) +
  thesis_theme + scale_colour_brewer(palette = "Dark2") +
  ggtitle(expression(paste("EPISCOPE: Distribution of ", beta, "-values", sep = ""))) +
  xlab(expression(paste(beta, "-values", sep = "")))
ggsave("EPISCOPE_beta_distribution_CGI_stratified.pdf", g, height = 9,
       width = 16)
```

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_beta_distribution.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. Most CpGs are highly unmethylated or highly methylated, with the buffy coat samples being the most extreme example of this bimodality. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values. Abbreviations: \em{BUF} = buffy coat; \em{SA} = subcutaneous adipose; \em{VA} = visceral adipose; \em{VAT} = adipocytes derived from visceral adipose.}
\label{fig:EPISCOPE_beta_distribution}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_beta_distribution_CGI_stratified.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Most CpGs outside of CpG islands are methylated whereas most CpGs within CpG islands are unmethylated. Spikes' in the density estimate are due to the discreteness of $\beta$-values. Abbreviations: \em{BUF} = buffy coat; \em{SA} = subcutaneous adipose; \em{VA} = visceral adipose; \em{VAT} = adipocytes derived from visceral adipose; \em{CGI} = CpG island.}
\label{fig:EPISCOPE_beta_distribution_CGI_stratified}
\end{figure}

\subsubsection{Strand-specific summaries}

CpG $\beta$-values are often computed by aggregating the $m$ and $u$ counts across the forward and reverse strands. This assumes that the two strands are indeed symmetrically methylated. To investigate this, for each CpG we can compute the $\beta$-values using only the forward-strand reads, $\beta^{+}$, and only the reverse-strand reads, $\beta^{-}$.

```{r EPISCOPE_strand_specific_beta-value_distributions, eval = FALSE, echo = FALSE}
# Run on one of the unices
setwd("/wehisan/home/allstaff/h/hickey/methylation_m-tuples/analyses/processed_data/EPISCOPE")
library(MethylationTuples)
library(ggplot2)
library(dplyr)
library(tidyr)
library(RColorBrewer)

thesis_theme <- theme_classic(base_size = 20)
thesis_theme <- theme_bw(base_size = 20)

y <- readRDS("EPISCOPE_1_tuples.rds")
y_fwd <- y[strand(y) == "+"]
y_rev <- y[strand(y) == "-"]

ol <- findOverlaps(unstrand(y_fwd), shift(unstrand(y_rev), -1L), type = "equal")
beta_fwd <- methLevel(y_fwd, min_cov = 5L)[queryHits(ol), ]
beta_rev <- methLevel(y_rev, min_cov = 5L)[subjectHits(ol), ]

cors <- matrix(unlist(mclapply(1:10, function(min_cov, y_fwd, y_rev, ol) {
  beta_fwd <- methLevel(y_fwd, min_cov = min_cov)[queryHits(ol), ]
  beta_rev <- methLevel(y_rev, min_cov = min_cov)[subjectHits(ol), ]
  sapply(seq_len(ncol(y)), function(i, x, y) {
    cor(x[, i], y[, i], use = "complete.obs")
    }, x = beta_fwd, y = beta_rev)
  }, y_fwd = y_fwd, y_rev = y_rev, ol = ol, mc.cores = 10)), ncol = 10)

beta_pairs <- data.frame(Individual = rep(substr(cn, 1, 3),
                                          each = nrow(beta_fwd)),
                         Tissue = rep(substr(cn, 4, 6), each = nrow(beta_fwd)),
                         beta_fwd = as.vector(beta_fwd),
                         beta_rev = as.vector(beta_rev),
                         stringsAsFactors = FALSE)
beta_pairs <- tbl_df(beta_pairs)

beta_pairs %>% group_by(Individual, Tissue) %>%
  summarise(cor(beta_fwd, beta_rev, use = "complete.obs"))

mypar <- function(a = 1, b = 1, brewer.n = 8, brewer.name = "Dark2",
                  cex.lab = 2, cex.main = 2, cex.axis = 1.5,
                  mar = c(5.1, 5.1, 3.5, 2.1), mgp = c(3, 1, 0), ...) {
    par(mar = mar, mgp = mgp, cex.lab = cex.lab, cex.main = cex.main,
        cex.axis = cex.axis)
    par(mfrow = c(a, b), ...)
    palette(brewer.pal(brewer.n, brewer.name))
}
pdf("EPISCOPE_beta_correlations_across_strand.pdf", width = 16, height = 9)
mypar(4, 3)
sapply(seq_len(ncol(beta_fwd)), function(i, x, y, ns) {
  smoothScatter(x[, i], y[, i],
                xlab = expression(beta["+"]),
                ylab = expression(beta["-"]),
                main = paste0(ns[i], ": cor = ", round(cor(x[, i], y[, i],
                                                        use = "complete.obs"),
                                                    2)))
  }, x = beta_fwd, y = beta_rev, ns = colnames(beta_fwd))
dev.off()
```

\subsection{Interpretation}

In papers about DNA methylation there is often a sentence like the following (given by \cite{Laird:2003kw}): "About 70% of the CpG dinucleotides in the mammalian genome are methylated". In the context of a whole-genome bisulfite-sequencing experiment, this can be interpreted as the expected $\beta$-value of a randomly selected CpG. However, as can be seen from __FIGURE__, the extreme bimodality of the $\beta$-value distribution means that the expected value is not a particularly useful estimate of the methylation level of a particular CpG. To make a useful statement about the methylation level of a particular CpG really requires that we know whether it is within a CpG island.

The "$70\%$" statement can also be interpreted as an estimate of the probability that a cytosine randomly sampled from a haploid copy of a mammalian genome is a methylcytosine. Because in a whole-genome bisulfite-sequencing experiment we measure methylation from a pool of cells, not a single haploid copy of the genome, I do not think it makes sense to describe a particular cytosine in the genome as being a 'methylcytosine'. That is, it only makes sense to describe an observation of a cytosine, $z_{i}$, as being methylated or unmethylated and not to describe a cytosine, $i$, as being a methylated or unmethylated. However, several important whole-genome bisulfite-sequencing papers, \citet{Lister:2008bh, Lister:2009hy, Lister:2011kg}[^Lister_2013], have sought to use the latter definition, which I believe to be an unnecessary source of confusion.

[^Lister_2013]: It is worth noting that this concept was not used in more recent paper from the same group \citep{Lister:2013et}.

\citet{Lister:2008bh} introduced a method "to identify the presence of a methylated cytosine" (detailed in the supplementary material). In the language of these papers, a 'methylcytosine' is a cytosine in the reference genome where "at least s [sic; I believe this should be "a"] subset of the genomes within the sample were methylated" \citep[supplementary material]{Lister:2009hy}. This amounts to testing the hypothesis $H_{0}: \beta_{i, j} = 0$ against the one-sided alternative $H_{1}: \beta_{i, j} > 0$. This can be thought of as testing the null hypothesis that the observed number of methylated reads at the $i^{th}$ cytosine were simply due to 'error', where the 'error' is a combination of the estimated sequencing error and the estimated bisulfite-converstion error.

Although the exact procedure is not particularly well described in any of \citet{Lister:2008bh, Lister:2009hy, Lister:2011kg}[^mathematical_description], nor is any code made available, I believe the method is as follows. For each cytosine compute a P-value, $P_{i} = \sum_{k = m_{i} + 1}^{k = m_{i} + u_{i}} Prob(X = k)$, where $X = Binom(u_{i} + m_{i}, \epsilon))$ and $\epsilon$ is the estimated 'error'. Any site with an FDR-adjusted P-value below a threshold was declared a 'methylcytosine'[^fdr]

[^mathematical_description]: The earliest of these papers, \citet{Lister:2008bh}, includes a short non-mathematical description, while the most detailed description is given in the supplementary material of \citet{Lister:2009hy}. \citet{Lister:2011kg} simply refers to \citet{Lister:2009hy}

[^fdr]: \citet{Lister:2008bh} used an FDR-adjusted P-value cutoff of $0.05$; \citet{Lister:2009hy} used an FDR-adjusted P-value cutoff of $0.01$. I presume the FDR-adjustment to be based on the Benjamini-Hochberg procedure \citep{Benjamini:1995ws}, although this is not explicitly stated. This procedure was performed separately for each methylation context in \citet{Lister:2009hy}, but it is not clear if this is the case for \citet{Lister:2008bh} (a study of _A. thaliani_, which has large amounts of non-CG methylation) or \citet{Lister:2011kg} (a study that includes pluripotent human cell lines that have non-negligible levels of non-CG methylation). This  affects the false discovery rate correction since the number of tests is far greater if all cytosines are simultaneously corrected compared to a separate correction for each context.

The $\epsilon$ are estimated on a per-sample basis, with bisulfite-conversion error rates estimated from the unmethylated chloroplast genome \citep{Lister:2008bh} or from the genome of the lambda phage spike-in control \citep{Lister:2009hy, Lister:2011kg}. It is not clear how the sequencing error rate was estimated, particularly given that the base qualities are not included in the data available from the website.

The proportion of cytosines that are identified by this procedure as 'methylcytosines' is a poor estimator of the probability that a cytosine randomly sampled from a haploid genome is methylated, unless the sample is incredibly homogeneous. For example, suppose we had a sample where the true methylation level of every CpG, $B_{i}$, was $0.2$. Given sufficient sequencing coverage, every CpG in the genome would be declared a 'methylcytosine' when in fact for any haploid copy of the genome only $20\%$ of CpGs would be expected to be methylcytosines.

Furthermore, referring to CpGs as 'methylcytosines' results in a loss of information since two 'methylcytosines' may have very different $\beta$-values. For example, in \citet[supplementary figure 2a]{Lister:2009hy} the authors use a Venn diagram to compare the number of 'methylcytosines' called in two biological replicates to summarise the concordance between the two biological replicates. A far better summary of the biological replicability is to plot the $\beta$-values from each replicate against one another as a scatter plot, as this includes the magnitude of the $\beta$-values and not just whether they are statistically different from zero.

In summary, for bisulfite-sequencing experiments where the DNA for each sample comes from multiple cells I do not think it makes sense, nor is it useful, to refer to individual cytosines, $i$, as being methylated or unmethylated. Instead, it is better to summarise the methylation level at a CpG by a $\beta$-value since this has a natural interpretation as the estimated proportion of haploid genomes in the sample that are methylated at that CpG. Unfortunately, $\beta$-values are not without their own issues, as discussed below.

\subsection{Correlations}

Many researchers have observed that DNA methylation is spatially correlated along the genome, \cite[e.g.,]{Eckhardt:2006gh, Cokus:2008fc, Li:2010fb, Hansen:2011gu, Hebestreit:2013ko, Wang:2011cw, Pedersen:2012vl, Lacey:2013iy, Sofer:2013bk, Liu:dy, Lyko:2010dr, Landan:2012kp, Lister:2009hy}). I call this correlation of methylation levels _co-methylation_.

Just as we can explore the distribution of DNA methylation levels within a sample or between samples, so too can we explore co-methylation within a sample or between samples. I examine this in detail in Chapters \ref{chap:co-methylation_review} and \ref{chap:co-methylation}.

\subsection{Biases}

The natural interpretation of $\beta_{i, j}$ as the average level of methylation at the $i^{th}$ locus in the $j^{th}$ sample will be biased if the probability of sequencing a fragment with a methylated site is different from the probability of sequencing a fragment with an unmethylated site. In fact, it has been shown that methylated DNA is overrepresented in bisulfite-sequencing data due, with the problem exacerbated by higher rounds of PCR amplification and dependent on the bisulfite-conversion protocol \cite{Ji:2014fq}. PCR amplification can result in overreprestation of one of the DNA strands in bisulfite-sequencing data \cite{Warnecke:1997eh}.

Lab-based solutions to overcome these biases exist for targeted bisulfite-sequencing, but are technically difficult and expensive to extend to whole-genome studies. \cite{Ji:2014fq} propose potential computational corrections for these biases but as yet these have not been implemented.

\subsection{Transformations}\label{sec:transformations}

$\beta$-values are the _de facto_ standard unit for reporting methylation levels due to their natural interpretation as an estimate of the average level of methylation at the locus. However, they are not necessarily the best unit for statistical inference. This is because a $\beta$-value is an estimate of a proportion and there are a well-known challenges when working with proportion data, such as:

1. The estimate of the standard error depends on the estimate of the mean (i.e., $\beta$), through $se(\beta) = \sqrt{\frac{\beta (1 - \beta)}{m + u}}$. Taking the derivative of this, we see that the maximum standard error, $\sqrt(\frac{0.25}{m + u})$, occurs at $\beta = 0.5$ and the minimum standard error, $0$, occurs at $\beta = 0, 1$.
2. We need to more than just the $\beta$-value to have a sense of how precise an estimate it is. Essentially, we need to also know the sequencing coverage of the methylation loci. Consider two CpGs, one with $m = 1, u = 3$ and the other with $m = 100$ $u = 300$. Both CpGs have $\beta = 1/4$ but the second CpG is measured with much greater precision. Assuming the binomial model, the first CpG has $se(\beta) = \sqrt{\frac{1/4 \times 3/4}{4}} = 0.22$ whereas the second CpG has $se(\beta) = \sqrt{\frac{1/4 \times 3/4}{400}} = 0.02$.
3. Proportions are bound between 0 and 1, inclusive.

To address (1), proportion data are often transformed via a variance stabilisation transformation. The aim is to make the variance independent of the mean, at least approximately. Popular variance stabilisation transformations include:

- The arcsine transformation, $\arcsin({\sqrt{\frac{m + 1}{m + u + 1}}})$ \citep{ANSCOMBE:1948bw}. A small value, in this case 1, is added to both $m$ and $u$ to avoid $\beta = 0, 1$.
- The "averaged arcsine" transformation, $\arcsin{\sqrt{\frac{m}{m + u + 1}}} + \arcsin{\sqrt{\frac{m + 1}{m + u + 1}}}$ \citep{Freeman:1950bh}. One problem with this transformation is that it does not have a unique inverse \citep{Nunes:2009vj}.

However, the general use of variance stabilising transformations for proportion data has fallen out of favour with the widespread availability of generalised linear model software, in particular for the logistic regression model \cite[e.g.,]{Warton:2011gm}.

One transformation that remains popular, at least in the analysis of DNA methylation microarray data, is the logit-transformation, also known as $\mathcal{M}$-values. An $\mathcal{M}$-value is defined as $logit_{2}(\beta) = \log_{2}(\frac{\beta}{1 - \beta} = \log_{2}\Big ( \frac{m + \alpha}{u + \alpha} \Big )$, where here $m$ and $u$ are the intensities from the methylated and unmethylated probes, respectively, and $\alpha$ is an offset to avoid a numerator or denominator that is zero. Similarly defined $\mathcal{M}$-values, also known as log-ratios, are widely and successfully used in analyses of RNA expression two-colour microarrays \cite[e.g.]{Smyth:2005ta}.

\cite{Du:2010dc} advocate for the use of $\mathcal{M}$-values in statistical analyses of methylation levels from microarray data. The main reason for this is that $\mathcal{M}$-values are approximately _homoscedastic_, i.e., their variances are approximately constant across the full range of $\mathcal{M}$-values. As already noted, the logit-transformation is not the only possible variance-stabilising transformation, but the familiarity of log-ratios to bioinformatics and genomics researchers makes it a favourable choice. Nonetheless, \cite{Du:2010dc} also advocate that the results of analyses should be reported as $\beta$-values, owing to their "more intuitive biological interpretation".

As with $\beta$-values, $\mathcal{M}$-values derived from bisulfite-sequencing data generally cannot be directly analysed due to the variable sequencing coverage across loci. However, the variable sequencing coverage can be accounted for using the afore-mentioned beta-binomial models and related regression models, which are discussed in the next section.

\section{Differential methylation}

A common downstream analysis is the identification of differential methylation, for which many methods have been proposed \citep[e.g.,]{Akalin:2012cmChen:2013eh, Chen:2014jb, Dolzhenko:2014bo, Gokhman:2014cp, Jaffe:2012gx, Lacey:2013iy, Lister:2009hy, Rijlaarsdam:2014hp, Sofer:2013bk, Xie:2014ez, Feng:2014iq, Hebestreit:2013ko, Sun:2014fk, Park:2014ho, Hansen:2012gr}. This includes the identification of differentially methylated cytosines (DMCs) and differentially methylated regions (DMRs).

\subsection{Experimental design}\label{sec:experimental_design}

In any analysis of differential methylation we want the DMCs and DMRs to be both _biologically_ and _statistically_ significant; it's no good if all the differences are simply due to technical artefacts or random fluctuations. Key to ensuring biological relevance is good experimental design, such as the use of _replicates_ in each experimental group. A distinction is often made between _technical replicates_ and _biological replicates_. Briefly, biological replicates are experimental units that all undergo the same treatment and are used to estimate the within-group variability of the treatment. Technical replicates are repeated measurements of the same experimental unit, perhaps with slight variations in the sample preparation, and are used to estimate the variability of the sample preparation and measurement process.

As is clear from this definition, the boundary between biological and technical replication is often a fuzzy one. For example, \cite{Lister:2009hy} state that, "_For each cell type, two __biological__ replicates were performed with cells of different passage number_" (emphasis mine). I contend that these are better defined as technical replicates since each replicate came from the same cell line, underwent passaging under near-identical conditions and differ only by the number of cell passages in each media[^cell_passaging].

[^cell_passaging]: In the case of IMR90 cell line, the first replicate, IMR90_r1, underwent 4 cell passages and the second replicate, IMR90_r2, underwent 5 cell passages. In the case of the H1 cell line, the first replicate, H1_r1, underwent 25 passages in the first media and 9 passages in the second media, and the second replicate, H1_r2, underwent 27 passages in the first media and 5 passages in the second media.

Unfortunately, early experiments with whole-genome bisulfite-sequencing rarely had replicates of any kind, or pooled the replicates prior to analysis, thus ignoring all within-group variability, as is the case of the H1 and IMR90 cell lines in \cite{Lister:2009hy}.

Ideally, technical variability is orthogonal to the biological variability, but this rarely occurs. Indeed, high-throughput sequencing experiments are particularly susceptible to batch effects, and other sources of unwanted variation, that can swamp the biological variation of interest \citep{Leek:2010jq}. Again, this emphasises the importance of good experimental design such as randomisation, replication and the use of control, both positive and negative.

\subsection{DMCs}\label{sec:dmcs}

There have been reports of differential methylation at individual CpGs resulting in a phenotypic difference \citep[e.g.,]{Furst:2012gh}. However, with approximately 25 million CpGs in the human genome, not to mention the many, many more non-CpG cytosines, it is an optimist who aims for the reliable detection of DMCs from whole-genome experiments with sample sizes you can count on one or two hands and an average sequencing depth of $10-30\times$. Nonetheless, identifying DMCs, even if not an end in itself, is an important step in identifying DMRs.

Identifying DMCs boils down to identifying differences in the mean level of methylation, for which there is an enormous body of statistical literature. The problem can be viewed as a 'stand-alone' test, such as a t-test, or framed as a regression problem to allow for the inclusion of additional covariates.

Regardless of the test used, all attempts to identify DMCs from whole-genome bisulfite-sequencing data must pay a large multiple-hypothesis testing penalty. Correcting for multiple hypothesis testing is standard practice in the analysis of genomics data. but the sheer number of tests, in this case approximately 25 million, is at least an order of magnitude greater than commonly tested in other genomics experiments (e.g., approximately 1 million tests in genome-wide association studies and 20,000 tests in studies of differential gene expression). Various methods have been used to correct for this multiple testing, as illustrated in Table \ref{tab:multiple_testing}.

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
Method                                       & Used by                                                                            \\ \midrule
\cite{Benjamini:1995ws}                      & \cite{Akalin:2012cm, Jaffe:2012gx, Lacey:2013iy, Rijlaarsdam:2014hp, Sofer:2013bk} \\
\cite{Wang:2011cw}                           & \cite{Akalin:2012cm}                                                               \\
"False discovery rate correction/adjustment" & \cite{Dolzhenko:2014bo, Gokhman:2014cp, Lister:2009hy, Xie:2014ez}                 \\
\cite{Storey:2007cp}                         & \cite{Jaffe:2012gx}                                                                \\
\cite{Benjamini:2001ho}                      & \cite{Sofer:2013bk, Hebestreit:2013ko}                                             \\
"Bonferonni adjustment"                      & \cite{Feng:2014iq}                                                                 \\ \bottomrule
\end{tabular}
\caption{Methods proposed for adjusting for the multiple hypothesis testing performed in an analysis of DMCs. Several papers use describe their analysing as performing a "false discovery rate adjustment" or "false discovery rate correction" without explicitly stating what they are doing or citing a reference. One paper uses an ultra-conservative Bonferonni correction.}
\label{tab:multiple_testing}
\end{table}

One thing to note, however, is that the _effective_ number of tests is fewer than the actual number of tests. This is because the methylation levels at neighbouring loci are correlated (see Chapters \ref{chap:co-methylation_review} and \ref{chap:co-methylation}), which means that tests of differential methylation are generally positively correlated, thus reducing the effective number of independent tests. The classical Benjamini-Hochberg procedure also controls the false discovery rate under certain forms of positive dependence \citep{Benjamini:2001ho}. (__TODO: Do these techniques work in the high-dependence setting of DMC testing?__)

\paragraph{One-group experiments}\label{sec:one_group_dmcs}

While an uncommon design, I did work on a one-group whole-genome bisulfite-sequencing experiment during my PhD. The data, courtesy of Harry Oey and Emma Whitelaw (La Trobe University), came from whole-genome bisulfite-sequencing of five _isogenic_ mice, meaning that these mice were genetically identical (or close to it).

In a one-group experiment a DMC is a cytosine where at least one of the samples has a different methylation level to the group average. The data for the $i^{th}$ methylation locus can be summarised by a $2 \times n$ contingency table, as shown in Table \ref{tab:one-group_dmc_testing}.

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
  \begin{tabular}{@{}lll@{}}
  \toprule
  & m          & u          \\ \midrule
  $Sample_{1}$ & $m_{i, 1}$ & $u_{i, 1}$ \\
  $\vdots$     & $\vdots$   & $\vdots$   \\
  $Sample_{n}$ & $m_{i, n}$ &            \\ \bottomrule
  \end{tabular}
}
\caption{$2 \times n$ contingency table summarising the methylation evidence for the $i^{th}$ methylation locus in a one-group experiment.}
\label{tab:one-group_dmc_testing}
\end{table}

Such a contingency table can be analysed in many different ways. Mathematically, to test whether this cytosine is a DMC we wish to test the null hypothesis  $H_{0}: \beta_{i, j} = \beta_{i, 0}$ for $j = 1, \ldots, n$ against the alternative that $H_{A}: \beta_{i, j} \neq \beta_{i, 0}$ for some $j. $\beta_{i, 0}$ may be pre-specified or estimated from the data by, for example, the group average $\frac{1}{n} \sum_{j = 1}^{n} \beta_{i, j}$).

To the best of my knowledge, none of the software specifically written for analysing differential methylation analyses is able to test this hypothesis. This is because these software are almost all designed for two-group experiments. However, this hypothesis is readily testing using routine statistical methods. As an example, consider the following data:

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
& m    & u    \\ \midrule
$Sample_{1}$ & $38$ & $1$  \\
$Sample_{2}$ & $79$ & $2$  \\
$Sample_{3}$ & $59$ & $1$  \\
$Sample_{4}$ & $69$ & $2$  \\
$Sample_{5}$ & $44$ & $46$ \\ \bottomrule
\end{tabular}
\caption{An example of a differentially methylated cytosine in a one-group experiment.}
\label{tab:one-group_example}
\end{table}

We could analyse this table using a simple stand-alone test, such as Pearson's $\chi^{2}$ test. Alternatively, we could frame this as a generalised linear model, either a log-linear model or a logistic regression. The equivalence of these methods is shown in the Appendix.

A key advantage of using either of the generalised linear models is the capability to include additional covariates in the model, such as the sex of each sample or terms to account for batch effects. More complex regression analyses are also possible, such as empirical Bayes or fully Bayesian models.

\paragraph{Multi-group experiments}\label{sec:two_group_dmcs}

Due to the cost of whole-genome bisulfite-sequencing, early experiments using this technology to study differential methylation compared one sample against another \citep[e.g.,]{Lister:2009hy, Berman:2012ga}. While this might be considered a two-group experiment, one without replicates, I believe this is better thought of as a one-group experiment with two samples. Indeed, the analysis of such an experiment uses techniques from Section \label{sec:one_group_dmcs}, e.g., Pearson's $\chi^{2}$ test or Fisher's exact test, which is better suited to the small sample sizes routinely seen in each $2 \times 2$ table.

The more interesting experiment is one that includes replicates within each group, since we can then estimate the within-group variation. Unfortunately, some authors have still used Fisher's exact test to analyse such experiments. For example, \cite{Lister:2009hy} test for differential methylation between the _H1_ and _IMR90_ samples using Fisher's exact test by first pooling data the two replicates they have for each sample. This is incorrect, because such an analysis effectievly ignores all within-group variability \citep{Hansen:2012gr}. Instead, a test that incorporates the within-group variability should be used.

In a two-group experiment this might be a t-test comparing the the average methylation level in the first group, $\bar{\beta_{i}^{(1)}}$, to that in the second group, $\bar{\beta_{i}^{(2)}}$. More generally, we can use a regression model to allow for multiple groups (i.e., ANOVA) and to incorporate additional covariates.

Several software packages are now available for identifying DMCs, although some are limited to the two-group experiment. Rather than directly analysing the raw $\beta$-values, these software generally make additional modelling assumptions, such as the beta-binomial model (Section \ref{sec:beta-binomial_model}), and/or perform some transformation, such as smoothing of the $\beta$-values (Section \ref{transformation}).

`DSS` \citep{Feng:2014iq}, `BiSeq` \citep{Hebestreit:2013ko}, `MOABS` \citep{Sun:2014fk} `methylSig` \citep{Park:2014ho} and `RADmeth` \citep{Dolzhenko:2014bo} are software packages that all use a Beta-Binomial hierarchical model of DNA methylation, although the exact details differ considerably between packages. `DSS` and `MOABS` use empirical Bayes methods to estimate parameters whereas `methylSig`, `BiSeq` and `RADmeth` use maximum likelihood estimation. `BiSeq` and `methlySig` also perform spatial smoothing of the data.

The test used to identify DMCs in these regression models is variously a Wald test (`BiSeq`, `DSS`), a likelihood ratio test (`methlySig`, `RADmeth`) or based on the credible interval (Bayesian confidence interval) of the difference in methylation between the two groups (`MOABS`).

Not all methods for identifying differential methylation are designed for identifying DMCs. Both `bsseq` \citep{Hansen:2012gr} and `Aclust` \citep{Sofer:2013bk} are methods explicitly designed for identifying DMRs rather than DMCs and, in the case of `bsseq`.

\subsection{DMRs}

A differentially methylated region is a region of the genome where there are multiple cytosines with evidence of differential methylation. Importantly, not all cytosines in the region must be genome-wide significant DMCs. Rather, the idea is that a DMR might capture a weaker but broader difference in methylation. Such a signal may well have a clearer biological interpretation and, indeed, might be expected such as in differences in methylation at a gene promoter.

There are two very different strategies for identifying differentially methylated regions:

1. Using regions that are defined _a priori_, which are then tested for differential methylation. Such regions might be CpG islands \citep[e.g.,]{Huang:1999fs,Doi:2009ky}[^array], MspI restriction fragments \citep[e.g.,]{Stockwell:2014fq}, any other predefined genomic feature, such as a gene promoters or transcription factor binding sites, or general predefined bins \citep[e.g., 100 bp bins]{Park:2014ho}.
2. Using data-driven regions, e.g., defined from an analysis of DMCs, which are then tested for differential methylation.

[^array]: Both these examples are from microarray studies, but the same idea can be applied to sequencing studies.

The former is much simpler to analyse but is limited in its ability to discover new differentially methylated regions. It is also hampered because the correct unit or scale for differential methylation may not be known for the experiment.

The latter offers the opportunity to identify previously unknown regions that are subject to differential methylation. Included in this is the opportunity to discover the scale over which differential methylation acts. However, valid statistical inference of these regions is far more challenging.

\subsection{Using _a priori_ regions}

The idea of testing for differential methylation is relatively straightforward. For each sample, the region-wise methylation level is some summary of the methylation level at each loci in that region, for example, the weighted average of $\beta$-values for all loci in the region, where the weights are proportional to the sequencing coverage. These averages are then compared between samples. I describe basic implementations of this for both one-group and multi-group experiments, and then detail a more sophisticated analysis implemented in `BiSeq` \citep{Hebestreit:2013ko}.

\subsubsection{One-group experiments}\label{sec:one_group_dmrs}

For a region containing $c$ methylation loci, construct the $n \times 2 \times c$ contingency table, i.e., each 'slice' of the table is a $n \times 2$ table containing the data for a single locus. Simply aggregating the counts across loci and testing the resulting $n \times 2$ table would be incorrect, since such an aggregating could introduce spurious differences in the proportions (__CITE: http://www.biostathandbook.com/cmh.html__). Instead, we could fit the log-linear model with `sample`, `meth_type` and `locus` as the set of factors. The saturated model is

\begin{equation*}
\log(F_{a, b, c}) = \lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} + \lambda_{a, b}^{meth_state, sample} + \lambda_{a, c}^{meth_state, locus} + \lambda_{b, c}^{sample, locus} + \lambda_{a, b, c}^{meth_state, sample, locus}
\end{equation*}

We would first need to test whether there is a significant three-way interaction between `meth_type`, `sample` and `locus` by computing the deviance of the model without the three-way interaction

\begin{equation*}
\log(F_{a, b, c}) = \lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} + \lambda_{a, b}^{meth_state, sample} + \lambda_{a, c}^{meth_state, locus} + \lambda_{b, c}^{sample, locus}
\end{equation*}

On the one hand, if there is no three-way interaction, then we can test for an interaction between `meth_type` and `sample`, i.e., test for a difference in methylation level across the loci, by comparing the model

\begin{equation*}
\log(F_{a, b, c}) = \lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} + \lambda_{a, b}^{meth_state, sample} + \lambda_{a, c}^{meth_state, locus} + \lambda_{b, c}^{sample, locus}
\end{equation*}

to the model

\begin{equation*}
\log(F_{a, b, c}) = \lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} + \lambda_{a, c}^{meth_state, locus} + \lambda_{b, c}^{sample, locus}
\end{equation*}

This test is analogous to the generalised Cochran-Mantel-Haenszel test \citep{Landis:1978vk,MANTEL:1978iv,Birch:1965uj} of the $n \times 2 \times c$ table \citep[p231]{agresti1990categorical}. The advantage of the Cochran-Mantel-Haenszel is its applicability for sparse datasets for which the asymptotic theory does not hold for the test based on the log-linear regression  \citep[p232]{agresti1990categorical}.

On the other hand, if there is a significant three-way interaction between `meth_type`, `sample` and `locus` then this requires more careful modelling to determine the nature of this interaction and whether the region is a consistent DMR.

\subsubsection{Multi-group experiments}\label{sec:multi_group_dmrs}

__TODO__

\subsubsection{A more complex example}\label{sec:a_more_complex_example}

`BiSeq` is designed for RRBS data and tests for differential methylation at "CpG clusters". These CpG clusters are based on an initical scan of the genome looking for CpGs that are within a given distance of one another and that have sufficient sequencing coverage across the set of samples. While there is clearly a data-dependent component to these cluster definitions, I reserve the use of 'data-driven regions' for those which are based on the methylation levels of loci rather than their genomic co-ordinates. \citep{Hebestreit:2013ko} alternatively recommend using the target regions of the assay, such as MspI fragments in the case of RRBS, instead of creating these CpG clusters.

One the clusters are defined, the $\beta$-values in each cluster are smoothed using a local binomial likelihood smoother. For each CpG `BiSeq` fits a beta regression model[^not_beta-binomial], which is tested for evidence of differential methylation at that cytosine. Based on these P-values, `BiSeq` then use a hierarchical testing procedure to control the false discovery rate at both the cluster-level and locus-level. This method is based on several papers by Yoav Benjamini and colleagues \citet{Benjamini:1997fh, Benjamini:2001ho, Benjamini:2006gd, Benjamini:2007cm}. It aims to first control the false discovery rate at the cluster-level and then refine the signal by trimming non-DMCs from those clusters that have been declared as differentially methylated. Finally, these differentially methylated clusters are _post hoc_ filtered to ensure they are _consistent_ (see Section \ref{sec:consistency_of_dmrs}).

[^not_beta-binomial]: This is different to the beta-binomial regression framework described in Section \ref{sec:beta-binomial_models}.

Based on my understanding of this procedure, only cytosines within differentially methylated clusters can be declared DMCs since it is only those cytosines that are tested in the second-step of the hierarchical testing algorithm. Therefore, isolated DMCs will not be detected by this procedure.

\subsection{Using data-driven regions}

The most common type of algorithm for identifying data-driven DMRs, and arguably also the most _ad hoc_, is based on detecting clusters of DMCs \citep[e.g.,]{Lister:2009hy, Lister:2011kg, Hansen:2011gu, Feng:2014iq}. The initial identification of DMCs will typically use a relaxed statistical significance threshold. For example, \cite{Hansen:2011gu} use all CpGs with a P-value in the lowest $5\%$ and declare putative DMRs to be contiguous runs of such CpGs that are within a given distance of one another and with "all differences in the same direction". This latter requirement is what I call the _consistency_ of the DMR (see Section \ref{sec:consistency_of_dmrs}).

These putative DMRs may be subject to further filtering, such as requiring that they contain a minimum number of CpGs and span a minimum number of bases, and the merging of nearby putative DMRs into a single putative DMR \citep{Hansen:2011gu}. Notably, many of these methods do not include a formal statistical test of differential methylation at the region-level \citep[e.g.,]{Lister:2009hy, Lister:2011kg, Hansen:2011gu, Feng:2014iq}.

The challenges of valid statistical inference at such data-driven regions are not unique to the problem of testing for differentially methylated regions. Similar problems arise in the analysis of chromatin immunoprecipitation sequencing (ChIP-seq) experiments \citep{Schwartzman:2011to, Lun:2014gn} and in the field of signal processing \citep{Schwartzman:2011vg}. It boils down to the problem of how to assign statistical significance to regions or peaks using the same data that were used to define these regions or peaks. These regions are already enriched for differences, which biases the interpretation the results of tests for differences at these regions. One way, and I would argue the best way, to avoid this issue is to test these regions using a separate dataset, which completely avoids the issue of statistical 'double-dipping'. If the sample size is large enough, then permutation testing may also be appropriate. For example, \cite{Hansen:2013eo} permute the group labels of their samples and re-ran their analysis to determine for each of the observed regions "how often we see another block of similar length and effect size anywhere in the genome and in any of the permutations". The chief limitation of the permutation strategy is the restricted number of permutations that are possible from small sample sizes and the time and computational resources it takes to perform analyse the data for a single permutation, which are often substantial. For example, \cite{Hansen:2013eo} only performed nine permutations.

A somewhat less _ad hoc_ strategy for identify differentially methylated regions is called "bump-hunting" \citep{Jaffe:2012gx}. Initially developed for methylation microarrays, and now available for broader use in the R/Bioconductor package `bumphunter` ([http://www.bioconductor.org/packages/release/bioc/html/bumphunter.html](http://www.bioconductor.org/packages/release/bioc/html/bumphunter.html)), bump-hunting is quite a general statistical problem. For the case of identifying DMRs, the idea is to plot the value of a statistic used to test for DMCs as a function of the position in the genome[^smoothing] and then to identify bumps in this plot. Bumps are identified as contiguous regions of the genome where the signal is above some threshold and may include an error term to account for the spatial correlation of the signal. The significance of these peaks may be assessed with a permutation strategy.

Another alternative for combining individual loci into data-driven DMRs is based on the loci-specific P-values. These methods can be thought of extensions to Fisher's method for combining P-values \citep{Fisher:1936vv}, by attempting to account for the correlation of tests at nearby methylation loci. `comb-p` \citep{Pedersen:2012vl} uses the Stouffer-Liptak-Kechris \citep{Stouffer:1949ua, Kechris:2010fw, Zaykin:2011jq,} correction for spatially correlated P-values. `SLIM` \citep{Wang:2011cw} is used by `methylKit` \citep{Akalin:2012cm} to do a similar correction.

Finally, there are a class of methods that turn the problem of identifying data-driven DMRs on its head by constructing the regions without first testing the individual loci for differential methylation. Then, only once these regions are defined, these methods test for differential methylation. This is different to using _a priori_ regions, since the regions are still data-defined, but not with respect to differential methylation. The only example I know of such a method is `Aclust` \citep{Sofer:2013bk}. `Aclust` first clusters CpGs into candidate regions by performing agglomerative nested clustering of the between-sample co-methylation. Briefly, this is the correlation of methylation levels at two loci across the samples[^between_sample_co-methylation]. These clusters are then tested for differential methylation using generalised estimating equations.

[^between_sample_co-methylation]: See Chapter \ref{co-methylation_review} for further details of between-sample co-methylation.

\subsubsection{Consistency of DMRs}\label{sec:consistency_of_dmrs}

While not often explicitly stated, it is desirable that a DMR is _consistent_. The definition of consistency is best illustrated by a pathological example describing a region that is inconsistent. Suppose we have a one-group experiment with five samples[^one_by_five] and are studying a region with five CpGs, all of which are DMCs. However, the first CpG is differentially methylated in the first sample, the second CpG is differentially methylated in the second sample, etc. While the entire region is 'differentially methylated', the difference is not observed consistently in any one sample. A consistent region is therefore one where the differences are 'in the same direction'.

[^one_by_five]: Equivalently, a five group experiment without replicates.

Consistent DMRs are most easily identified in two-group experiments, where we can use the sign of the test statistic to determine the 'direction' of the signal. This is exactly what is done by, for example, \cite{Hansen:2011gu}.

In one-group experiments, the idea of consistency is similar to testing the homogeneity of odds ratios for all $n \times 2$ contigency tables in the region. However, __TODO. What about a region that is consistent but the odds ratios increases across the region__

\section{Other downstream analyses}

\subsection{PMDs}

- Lister
- Berman

\subsection{Variably methylated regions and differential variability}\label{sec:vmrs}

- Belinda
- Jaffe

\subsection{Allele-specific methylation}\label{sec:asm}

\subsection{Methylation entropy}\label{sec:methylation_entropy}

\subsection{Epipolymorphism}\label{sec:epipolymorphism}

\subsection{Epiallele detection}\label{sec:epialleles}

\subsection{Integrating DNA methylation data with other genomes data}\label{sec:data_integration}

\section{General TODOs}

* Discuss distribution of coverage? Perhaps in context of sequencing bias?