### 2014-12-06 data/observable_m-tuples/histogram_of_observable_m-tuples.Rmd
CpGs per read data

---
title: "Observable m-tuples"
author: "Peter Hickey"
date: "12/6/2014"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(scales)
thesis_theme <- theme_classic(base_size = 20)
thesis_theme <- theme_bw(base_size = 20)
```

# EPISCOPE

```{r}
dataset <- "EPISCOPE"
hist <- lapply(paste0(dataset, "/",
                      list.files(dataset, "\\.CG\\_per\\_read\\.hist")), 
               read.table, header = TRUE)
sn <- gsub("\\.CG\\_per\\_read\\.hist", "", 
           list.files(dataset, "\\.CG\\_per\\_read\\.hist"))
hist_df <- bind_rows(hist) %>% 
  mutate(Sample = rep(sn, sapply(hist, nrow))) %>%
  group_by(Sample) %>%
  mutate(Proportion = count / sum(as.numeric(count)))
g <- hist_df %>%
  ggplot(aes(x = n, y = Proportion * 100)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ Sample, ncol = 4, nrow = 3) +
  coord_cartesian(xlim = c(-1, 8.5)) + 
  scale_y_continuous(label = comma) + 
  xlab("Number of CpGs per read\n(Axis truncated at 8)") + 
  ylab("Percentage of total reads") + 
  ggtitle(paste0(dataset, ": Number of CpGs per read")) +
  thesis_theme
ggsave(paste0(dataset, "_CpGs_per_read.pdf"), g, width = 16, height = 9)
```

# Lister

```{r}
dataset <- "Lister"
hist <- lapply(paste0(dataset, "/",
                      list.files(dataset, "\\.CG\\_per\\_read\\.hist")), 
               read.table, header = TRUE)
sn <- gsub("\\.CG\\_per\\_read\\.hist", "", 
           list.files(dataset, "\\.CG\\_per\\_read\\.hist"))
sn <- gsub("DM\\_", "", sn)
hist_df <- bind_rows(hist) %>% 
  mutate(Sample = rep(sn, sapply(hist, nrow))) %>%
  group_by(Sample) %>%
  mutate(Proportion = count / sum(as.numeric(count)))
g <- hist_df %>%
  ggplot(aes(x = n, y = Proportion * 100)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ Sample, ncol = 5, nrow = 4) +
  coord_cartesian(xlim = c(-1, 8.5)) + 
  scale_y_continuous(label = comma) + 
  xlab("Number of CpGs per read\n(Axis truncated at 8)") + 
  ylab("Percentage of total reads") + 
  ggtitle(paste0(dataset, ": Number of CpGs per read")) +
  thesis_theme
ggsave(paste0(dataset, "_CpGs_per_read.pdf"), g, width = 16, height = 9)
```

# Seisenberger

```{r}
dataset <- "Seisenberger"
hist <- lapply(paste0(dataset, "/",
                      list.files(dataset, "\\.CG\\_per\\_read\\.hist")), 
               read.table, header = TRUE)
sn <- gsub("\\.CG\\_per\\_read\\.hist", "", 
           list.files(dataset, "\\.CG\\_per\\_read\\.hist"))
sn <- gsub("DM\\_", "", sn)
hist_df <- bind_rows(hist) %>% 
  mutate(Sample = rep(sn, sapply(hist, nrow))) %>%
  group_by(Sample) %>%
  mutate(Proportion = count / sum(as.numeric(count)))
g <- hist_df %>%
  ggplot(aes(x = n, y = Proportion * 100)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ Sample, ncol = 3, nrow = 1) +
  coord_cartesian(xlim = c(-1, 8.5)) + 
  scale_y_continuous(label = comma) + 
  xlab("Number of CpGs per read\n(Axis truncated at 8)") + 
  ylab("Percentage of total reads") + 
  ggtitle(paste0(dataset, ": Number of CpGs per read")) +
  thesis_theme
ggsave(paste0(dataset, "_CpGs_per_read.pdf"), g, width = 16, height = 9)
```

# Ziller

```{r}
dataset <- "Ziller"
hist <- lapply(paste0(dataset, "/",
                      list.files(dataset, "\\.CG\\_per\\_read\\.hist")), 
               read.table, header = TRUE)
sn <- gsub("\\.CG\\_per\\_read\\.hist", "", 
           list.files(dataset, "\\.CG\\_per\\_read\\.hist"))
sn <- gsub("\\.rmdup", "", sn)
hist_df <- bind_rows(hist) %>% 
  mutate(Sample = rep(sn, sapply(hist, nrow))) %>%
  group_by(Sample) %>%
  mutate(Proportion = count / sum(as.numeric(count)))
g <- hist_df %>%
  ggplot(aes(x = n, y = Proportion * 100)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ Sample, ncol = 5, nrow = 4) +
  coord_cartesian(xlim = c(-1, 8.5)) + 
  scale_y_continuous(label = comma) + 
  xlab("Number of CpGs per read\n(Axis truncated at 8)") + 
  ylab("Percentage of total reads") + 
  ggtitle(paste0(dataset, ": Number of CpGs per read")) +
  thesis_theme
ggsave(paste0(dataset, "_CpGs_per_read.pdf"), g, width = 16, height = 9)
```