### 2014-12-09 Rmarkdown/appendix.Rmd
Re-jigged appendix

---
title: "Appendix"
author: "Peter Hickey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
    keep_tex: no
  html_document:
    keep_md: yes
bibliography: ../latex/phd_thesis.bib
---

\section{Equivalence of tests to identify DMCs in a one-group experiment}\label{app:one_group_equivalence}

\subsection{Mathematical equivalence}

The saturated log-linear model can be written as $\log(F_{a, b}) = \lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{a, b}^{meth_state,sample}$, where $F_{a, b}$ is the value of the $(a, b)$ element in the $n \times 2$ contingency table. The null hypothesis that the cytosine is not a DMC is then tested by testing for an association between `sample` and `meth_state`, i.e., whether the parameter $\lambda_{a, b}^{meth_state, sample}$ is significantly different from zero.

This log-linear model can be reformulated as a logistic regression by noting that the log-odds (logits) for each row are $\log(\frac{F_{a, 1}}{F_{a, 2}}) = \log(F_{a, 1}) - \log(F_{a, 2})$. Therefore,

\begin{align*}
\log(\frac{F_{a, 1}}{F_{a, 2}}) &= \lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{1}^{sample} + \lambda_{a, 1} \\
&- (\lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{2}^{sample} + \lambda_{a, 2}) \\
&= \underbrace{(\lambda_{1}^{sample} - \lambda_{2}^{sample})}_{\alpha} + \underbrace{(\lambda_{a, 1} - \lambda_{a, 2})}_{\gamma_{i}}
\end{align*}

The expression $\log(\frac{F_{a, 1}}{F_{a, 2}}) = \alpha + \gamma_{i}$ is a logistic regression model. We can test the null hypothesis that the cytosine is not a DMC by testing for a `sample`-effect on the log-odds, i.e., whether the parameter $\gamma_{i}$ is significantly different from zero.

Regardless of whether we use the log-linear model or the logistic model, to test the null hypothesis that the cytosine is not a DMC we can compute the deviance[^deviance] of the reduced model and compare it to a $\chi^{2}$ random variable with $(2 - 1) \times (n - 1)$ degrees of freedom. For the log-linear regression the reduced model is $\log(F_{a, b}) = \lambda_{0} + \lambda_{a}^{meth_state} + \lambda_{b}^{sample}$ and for the logistric regression the reduced model is $\log(\frac{F_{a, 1}}{F_{a, 2}}) = \alpha$.

[^deviance]: The deviance is the log-likelihood ratio statistic comparing the current model within a saturated model.

\subsection{Computational equiavelence}

The computational equivalence of using Pearson's $\chi^{2}$ test, the deviance of a log-linear model and the deviance of a logistic regression are shown in the following R code:

```{r}
# The data
x <- cbind(M = c(38, 79, 59, 69, 44), U = c(1, 2, 1, 2, 46))
names(dimnames(x)) <- c("sample", "meth_state")
x

# Using Pearson's chi-square test
chisq.test(x)

# Fitting a log-linear model to perform iterative proportional fitting to the
# contingency table.
loglin(x, margin = list("sample", "meth_state"))
# An alternative formulation using loglm(), a convenient wrapper of loglin(),
# from the MASS package.
library(MASS)
loglm(formula =  ~ sample + meth_state, data = x)

# Fitting a generalised linear model with a binomial link using two different
# formulations
# (1) Using the array of frequencies
glm(x ~ 1, family = binomial)
# (2) Using relative frequencies (i.e., the beta-values) and weights (i.e.,
  # the sequencing coverage).
  glm(x[,1] / (x[, 1] + x[, 2]) ~ 1, family = binomial, weights = x[, 1] + x[, 2])
```

\section{The probability that two dependent Bernoulli random variables are identical}\label{app:bernoulli_equal}

\cite{Lindqvist:1978fm} wrote a brief note on Bernoulli trials with dependence.
Building on earlier work by \cite{Klotz:1973if}, Lindqvist parameterises the
Bernoulli process $X_{1}, X_{2}, \ldots$ on $\{0, 1\}$ by the parameters
$p = Pr(X_{i} = 1)$ and $c = cor(X_{i - 1}, X_{i})$ and shows that the
transition matrix is given by

\begin{equation*}
\label{eq:transition_matrix_id}
\Pi =
 \begin{pmatrix}
  (1 - p) + cp & p (1 - c) \\
  (1 - p) (1 - c) & p + c (1 - p)
 \end{pmatrix}
\end{equation*}

provided that $max(1 - \frac{1}{p}, 1 - \frac{1}{1 - p}) \leq c \leq 1$.

From this we can compute the joint distribution,

\begin{equation*}
Pr(X_{1} = x_{1}, \ldots, X_{n} = x_{n}) = Pr(X_{1} = x_{1}) Pr(X_{2} = x_{2} | X_{2} = x_{2}) \cdots Pr(X_{n} = x_{n} | X_{n - 1} = x_{n - 1})
\end{equation*}

In particular, in the case $n = 2$ we can compute the probability that two
dependent but identically distributed Bernoulli random variables are equal.

To extend the above result to the probability that two dependent and
__non-identically__ distributed Bernoulli random variables are equal, simply
requires that we derive the appropriate transition matrix. Switching notation
to that used in Chapter \ref{chap:co-methylation_review}, let
$Z_{h, i} \eqd Bernoulli(p_{h})$ and $Z_{h', i'} \eqd Bernoulli(p_{h'})$. The
transition matrix,
$\Pi = \bigl(\begin{smallmatrix}
Pr(Z_{h', i'} = z_{h', i'} | Z_{h, i} = z_{h, i})
\end{smallmatrix} \bigr)$, is given by

\begin{equation*}
\label{eq:transition_matrix_nid}
\Pi =
 \begin{pmatrix}
  (1 - p_{h'}) + c p_{h} & p_{h'} - c p_{h} \\
  \big [ (1 - p_{h'}) + c p_{h} \big ] (1 - p_{h}) & \big [ p_{h'} + c (1 - p_{h}) \big ] p_{h}
 \end{pmatrix}
\end{equation*}

We can then compute the desired probability

\begin{align*}
Pr(Z_{h, i} = Z_{h', i'}) &= Pr(Z_{h, i} = 0, Z_{h', i'} = 0) + Pr(Z_{h, i} = 1, Z_{h', i'} = 1) \\
                  &= Pr(Z_{h', i'} = 0 | Z_{h, i} = 0) Pr(X_{0} = 0) +
                     Pr(Z_{h', i'} = 1 | Z_{h, i} = 1) Pr(X_{0} = 1) \\
                  &= \big [ (1 - p_{h'}) + c p_{h} \big ] (1 - p_{h}) +
                     \big [ p_{h'} + c ( 1 - p_{h}) \big ] p_{h} \\
                  &= (1 - p_{h}) (1 - p_{h'}) + c p_{h} \big [1 - p_{h}] +
                     p_{h} p_{h'} + c p_{h} (1 - p_{h}) \\
                  &= (1 - p_{h}) (1 - p_{h'}) + p_{h} p_{h'} + 2 c p_{h} (1 - p_{h})
\end{align*}

__TODO: What is the general restriction on $c$ when $p_{h, i} \neq p_{h', i'}$?__

\section{Estimate of number of CpGs used by \citet{Lacey:2013iy}}\label{app:lacey}

I want to estimate the number of CpGs used by \citet{Lacey:2013iy} to estimate the correlation of methylation levels. They used the chromosome 11 data from an RRBS experiment on a normal myotube cell line, _MTCTL2_. Unfortunately, there is no URL provided to these data. \citet{Lacey:2013iy} cite their data as having come from \citet{Tsumagari:2013gn}, which includes a sample "Normal Human Skeletal Muscle Myoblasts", _HSMM_ ([http://genome-mirror.duhs.duke.edu/ENCODE/cellTypes.html](http://genome-mirror.duhs.duke.edu/ENCODE/cellTypes.html), [http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/](http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/)). I believe this to be the same cell line. In any case, for my purposes I simply needed some comparable RRBS data, not necessarily the exact same sample as  _MTCTL2_.

There are approximately $60,000$ CpGs on chromosome 11 in these comparable RRBS data. The following R code shows how I computed this using three RRBS replicates of the _HSMM_ sample.

Firstly, I download the data and unzip it:
```{r, cache = FALSE, message = FALSE}
library(R.utils)
library(data.table)
dl_dir <- "../data/appendix/"
url <- 'http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/'
files <- paste0('wgEncodeHaibMethylRrbsHsmmtubefshdDukeSitesRep',
                1:3, '.bed.gz')
dests <- paste0(dl_dir, files)
x <- mapply(function(f, d) {
  if (!file.exists(gsub(".gz", "", paste0(dl_dir, f)))) {
    download.file(url = paste0(url, f), d)
    gunzip(filename = d, overwrite = TRUE)
    }
  fread(gsub(pattern = '.gz', replacement = '', x = d), sep = '\t',
        verbose = FALSE)
  }, files, dests, SIMPLIFY = FALSE)
```

Then I count the number of CpGs on chromosome 11 in each replicate:
```{r}
y <- rbindlist(x)
y[V1 == "chr11", .N, by = list(V4)]
```

\section{Computing details}\label{app:computing_details}

All computational work was performed on one of the Bioinformatics Division's HP Blade servers. These are shared-use, shared-memory machines. The basic specifications are shown in Table \ref{tab:unices}.

\begin{table}[h]
\centering
\begin{tabular}{@{}lllll@{}}
\toprule
Machine name & Processors                           & Cores per processor & Total number of cores & RAM    \\ \midrule
unix88       & $4\times$ Intel Xeon X7350 @ 2.93GHz & 4                   & 16                    & 128 GB \\
unix301      & $4\times$ AMD Opteron 8435 @ 2.6GHz  & 6                   & 24                    & 256 GB \\
unix302      & $4\times$ AMD Opteron 6174 @ 2.2GHz  & 12                  & 48                    & 512 GB \\
unix303      & $4\times$ AMD Opteron 6176 @ 2.3GHz  & 12                  & 48                    & 512 GB \\
unix305      & $4\times$ AMD Opteron 6276 @ 2.3GHz  & 16                  & 64                    &        \\ \bottomrule
\end{tabular}
\caption{Bioinformatics Division server specifications.}
\label{tab:unices}
\end{table}

## Software details

```{r}
devtools::session_info()
```

### TODOs

* Cite data source and publication