### 2014-12-11 Rmarkdown/wgbs_downstream_analyses.Rmd

---
title: "Downstream analyses of whole-genome bisulfite-sequencing data"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

\chapter{Downstream analyses of whole-genome bisulfite-sequencing data}\label{chap:wgbs_downstream_analyses}

This chapter discusses various statistical analyses of bisulfite-sequencing data, which I call _downstream analyses_.

The first half of the chapter deals with analyses that are based on aggregate methylation levels, such as $\beta$-values. I explain the statistical questions behind the common downstream analyses of identifying differential methylation, variable methylation and partially methylated domains.

The second half of the chapter reviews downstream analyses that are based on methylation patterns at m-tuples. I introduce `MethylationTuples`, an R package I am developing for managing, analysing and visualising methylation patterns at m-tuples, which is complementary to `methtuples` --- `methtuples` extracts the methylation patterns at m-tuples and `MethylationTuples` provides a framework for analysing these data, which is currently lacking for this type of data.

\section{Differential methylation}\label{sec:differential_methylation}

A common downstream analysis is the identification of differential methylation. This includes the identification of differentially methylated cytosines (DMCs) and differentially methylated regions (DMRs). Many methods have been proposed \citep[e.g.,]{Akalin:2012cm, Chen:2013eh, Chen:2014jb, Dolzhenko:2014bo, Gokhman:2014cp, Jaffe:2012gx, Lacey:2013iy, Lister:2009hy, Rijlaarsdam:2014hp, Sofer:2013bk, Xie:2014ez, Feng:2014iq, Hebestreit:2013ko, Sun:2014fk, Park:2014ho, Hansen:2012gr}.

During my PhD I needed to analyse data from one-group experiment for differential methylation, admittedly an unusual design. Almost all these existing methods are designed for two-group experiments, with some being general enough to extend to multiple-group experiments[^multi-group_extension]. Since I spent time working on the one-group cases, and also because a review article for the two-group case was recently published \citep{Robinson:2014jz}, I pay somewhat disproportianate attention to the one-group scenario in what follows.

[^multi-group_extension]: The extension to multiple-group experiments isn't always implemented in the software, even if discussed in the paper.

\subsection{Experimental design}\label{sec:experimental_design}

In any analysis of differential methylation we want the DMCs and DMRs to be both _biologically_ and _statistically_ significant; it's no good if all the differences are simply due to technical artefacts or random fluctuations. Key to ensuring biological relevance is good experimental design, such as the use of _replicates_ in each experimental group. A distinction is often made between _technical replicates_ and _biological replicates_. Briefly, biological replicates are experimental units that all undergo the same treatment and are used to estimate the within-group variability of the treatment. Technical replicates are repeated measurements of the same experimental unit, perhaps with slight variations in the sample preparation, and are used to estimate the variability of the sample preparation and measurement process.

As is clear from this definition, the boundary between biological and technical replication is often a fuzzy one. For example, \citet{Lister:2009hy} state that, "For each cell type, two __biological__ replicates were performed with cells of different passage number" (emphasis mine). I contend that these are better defined as technical replicates since each replicate came from the same cell line, underwent passaging under near-identical conditions and differ only by the number of cell passages in each media[^cell_passaging].

[^cell_passaging]: In the case of IMR90 cell line, the first replicate, IMR90_r1, underwent 4 cell passages and the second replicate, IMR90_r2, underwent 5 cell passages. In the case of the H1 cell line, the first replicate, H1_r1, underwent 25 passages in the first media and 9 passages in the second media, and the second replicate, H1_r2, underwent 27 passages in the first media and 5 passages in the second media.

Unfortunately, early experiments with whole-genome bisulfite-sequencing rarely had replicates of any kind, or pooled the replicates prior to analysis, thus ignoring all within-group variability, as is the case of the H1 and IMR90 cell lines in \citet{Lister:2009hy}.

Ideally, technical variability is orthogonal to the biological variability, but this rarely occurs. Indeed, high-throughput sequencing experiments are particularly susceptible to batch effects, and other sources of unwanted variation, that can swamp the biological variation of interest \citep{Leek:2010jq}. Again, this emphasises the importance of good experimental design such as randomisation, replication and the use of control, both positive and negative.

\subsection{DMCs}\label{sec:dmcs}

There have been reports of differential methylation at individual CpGs resulting in a phenotypic difference \citep[e.g.,]{Furst:2012gh}. However, with approximately 25 million CpGs in the human genome, not to mention the many, many more non-CpG cytosines, it is an optimist who aims for the reliable detection of DMCs from whole-genome experiments with sample sizes you can count on one or two hands and an average sequencing depth of $10-30\times$. Nonetheless, identifying DMCs, even if not an end in itself, is an important step in identifying DMRs.

Identifying DMCs boils down to identifying differences in the mean level of methylation, for which there is an enormous body of statistical literature. The problem can be viewed as a 'stand-alone' test, such as a t-test, or framed as a regression problem to allow for the inclusion of additional covariates.

Regardless of the test used, all attempts to identify DMCs from whole-genome bisulfite-sequencing data must pay a large multiple-hypothesis testing penalty. Correcting for multiple hypothesis testing is standard practice in the analysis of genomics data, but the sheer number of tests, in this case approximately 25 million, is at least an order of magnitude greater than commonly tested in other genomics experiments[^multiple_testing]. Various methods have been used to correct for this multiple testing, as illustrated in Table \ref{tab:multiple_testing}.

[^multiple_testing]: For example, there are approximately 20,000 tests in studies of differential gene expression and one million tests in genome-wide association studies.

\begin{table}[h]
\centering
\begin{tabulary}{\textwidth}{LL}
\toprule
Method                                       & Used by                                                                            \\ \midrule
\citet{Benjamini:1995ws}                      & \citet{Akalin:2012cm, Jaffe:2012gx, Lacey:2013iy, Rijlaarsdam:2014hp, Sofer:2013bk} \\
\citet{Wang:2011cw}                           & \citet{Akalin:2012cm}                                                               \\
``False discovery rate correction/adjustment'' & \citet{Dolzhenko:2014bo, Gokhman:2014cp, Lister:2009hy, Xie:2014ez}                 \\
\citet{Storey:2007cp}                         & \citet{Jaffe:2012gx}                                                                \\
\citet{Benjamini:2001ho}                      & \citet{Sofer:2013bk, Hebestreit:2013ko}                                             \\
"Bonferonni adjustment"                      & \citet{Feng:2014iq}                                                                 \\ \bottomrule
\end{tabulary}
\caption{Methods proposed for adjusting for the multiple hypothesis testing performed in an analysis of DMCs. Several papers use describe their analysing as performing a "false discovery rate adjustment" or "false discovery rate correction" without explicitly stating what they are doing or citing a reference. One paper uses an ultra-conservative Bonferonni correction.}
\label{tab:multiple_testing}
\end{table}

One thing to note, however, is that the _effective_ number of tests is fewer than the actual number of tests. This is because the methylation levels at neighbouring loci are correlated (see Chapters \ref{chap:co-methylation_review} and \ref{chap:co-methylation}), which means that tests of differential methylation are generally positively correlated, thus reducing the effective number of independent tests. The classical Benjamini-Hochberg procedure also controls the false discovery rate under certain forms of positive dependence \citep{Benjamini:2001ho}. (__TODO: Do these techniques work in the high-dependence setting of DMC testing?__)

\subsubsection{One-group experiments}\label{sec:one_group_dmcs}

While an uncommon design, I did work on a one-group whole-genome bisulfite-sequencing experiment during my PhD. The data, courtesy of Harry Oey and Emma Whitelaw (La Trobe University), came from whole-genome bisulfite-sequencing of five _isogenic_ mice, meaning that these mice were genetically identical (or close to it).

In a one-group experiment a DMC is a cytosine where at least one of the samples has a different methylation level to the group average. The data for the $i^{th}$ methylation locus can be summarised by a $2 \times n$ contingency table, as shown in Table \ref{tab:one-group_dmc_testing}.

\begin{table}[h]
\centering
  \begin{tabular}{@{}lll@{}}
  \toprule
  & m          & u          \\ \midrule
  $Sample_{1}$ & $m_{i, 1}$ & $u_{i, 1}$ \\
  $\vdots$     & $\vdots$   & $\vdots$   \\
  $Sample_{n}$ & $m_{i, n}$ &            \\ \bottomrule
  \end{tabular}
\caption{$2 \times n$ contingency table summarising the methylation evidence for the $i^{th}$ methylation locus in a one-group experiment.}
\label{tab:one-group_dmc_testing}
\end{table}

Such a contingency table can be analysed in many different ways. Mathematically, to test whether this cytosine is a DMC we wish to test the null hypothesis  $H_{0}: \beta_{i, j} = \beta_{i, 0}$ for $j = 1, \ldots, n$ against the alternative that $H_{A}: \beta_{i, j} \neq \beta_{i, 0}$ for some $j$. $\beta_{i, 0}$ may be pre-specified or estimated from the data by, for example, the group average $\frac{1}{n} \sum_{j = 1}^{n} \beta_{i, j}$.

To the best of my knowledge, none of the software specifically written for analysing differential methylation analyses is able to test this hypothesis. This is because these software are almost all designed for two-group experiments. However, this hypothesis is readily testing using routine statistical methods. As an example, consider the following data:

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
& m    & u    \\ \midrule
$Sample_{1}$ & $38$ & $1$  \\
$Sample_{2}$ & $79$ & $2$  \\
$Sample_{3}$ & $59$ & $1$  \\
$Sample_{4}$ & $69$ & $2$  \\
$Sample_{5}$ & $44$ & $46$ \\ \bottomrule
\end{tabular}
\caption{An example of a differentially methylated cytosine in a one-group experiment.}
\label{tab:one-group_example}
\end{table}

We could analyse this table using a simple stand-alone test, such as Pearson's $\chi^{2}$ test. Alternatively, we could frame this as a generalised linear model, either a log-linear model or a logistic regression. The equivalence of these methods is shown in Appendex \ref{app:one_group_equivalence}.

A key advantage of using either of the generalised linear models is the capability to include additional covariates in the model, such as the sex of each sample or terms to account for batch effects. More complex regression analyses are also possible, such as empirical Bayes or fully Bayesian models.

\subsubsection{Multi-group experiments}\label{sec:multi_group_dmcs}

Due to the cost of whole-genome bisulfite-sequencing, early experiments using this technology to study differential methylation compared one sample against another \citep[e.g.,]{Lister:2009hy, Berman:2012ga}. While this might be considered a two-group experiment, one without replicates, I believe this is better thought of as a one-group experiment with two samples. Indeed, the analysis of such an experiment uses techniques from Section \ref{sec:one_group_dmcs}, e.g., Pearson's $\chi^{2}$ test or Fisher's exact test \citep{Fisher:1922bx}, which is better suited to the small sample sizes routinely seen in each $2 \times 2$ table.

The more interesting experiment is one that includes replicates within each group, since we can then estimate the within-group variation. Unfortunately, some authors have still used Fisher's exact test to analyse such experiments. For example, \citet{Lister:2009hy} test for differential methylation between the _H1_ and _IMR90_ samples using Fisher's exact test by first pooling data from the two replicates they have for each sample. This is incorrect, because such an analysis effectievly ignores all within-group variability \citep{Hansen:2012gr}. Instead, a test that incorporates the within-group variability should be used.

In a two-group experiment this might be a t-test comparing the the average methylation level in the first group, $\widebar{\beta_{i}^{(1)}}$, to that in the second group, $\widebar{\beta_{i}^{(2)}}$. More generally, we can use a regression model to allow for multiple groups (i.e., ANOVA) and to incorporate additional covariates.

Several software packages are now available for identifying DMCs, although some are limited to the two-group experiment. Rather than directly analysing the raw $\beta$-values, these software generally make additional modelling assumptions, such as the beta-binomial model (see Section \ref{sec:beta-binomial_models}), and/or perform some transformation of the data, such as smoothing of the $\beta$-values (see Section \ref{sec:transformations}).

`DSS` \citep{Feng:2014iq}, `BiSeq` \citep{Hebestreit:2013ko}, `MOABS` \citep{Sun:2014fk} `methylSig` \citep{Park:2014ho} and `RADmeth` \citep{Dolzhenko:2014bo} are software packages that all use a Beta-Binomial hierarchical model of DNA methylation, although the exact details differ considerably between packages. `DSS` and `MOABS` use empirical Bayes methods to estimate parameters whereas `methylSig`, `BiSeq` and `RADmeth` use maximum likelihood estimation. `BiSeq` and `methlySig` also perform spatial smoothing of the data.

The test used to identify DMCs in these regression models is variously a Wald test (`BiSeq`, `DSS`), a likelihood ratio test (`methlySig`, `RADmeth`) or based on the credible interval (Bayesian confidence interval) of the difference in methylation between the two groups (`MOABS`).

Not all methods for identifying differential methylation are designed for identifying DMCs. Both `bsseq` \citep{Hansen:2012gr} and `Aclust` \citep{Sofer:2013bk} are methods explicitly designed for identifying DMRs rather than DMCs.

\subsection{DMRs}

A differentially methylated region is a region of the genome where there are multiple cytosines with evidence of differential methylation. Importantly, not all cytosines in the region must be genome-wide significant DMCs. Rather, the idea is that a DMR might capture a weaker but broader difference in methylation. Such a signal may well have a clearer biological interpretation, such as in differences in methylation at a gene promoter.

There are two very different strategies for identifying differentially methylated regions:

1. Using regions that are defined _a priori_, which are then tested for differential methylation. Such regions might be CpG islands \citep[e.g.,][]{Huang:1999fs,Doi:2009ky}[^array], MspI restriction fragments \citep[e.g.,][]{Stockwell:2014fq}, any other predefined genomic feature, such as a gene promoters or transcription factor binding sites, or general predefined bins \citep[e.g., 100 bp bins used by][]{Park:2014ho}.
2. Using data-driven regions, e.g., defined from an analysis of DMCs, which are then tested for differential methylation.

[^array]: Both these examples are from microarray studies, but the same idea can be applied to sequencing studies.

The former is much simpler to analyse but is limited in its ability to discover new differentially methylated regions. It is also hampered because the correct unit or scale for differential methylation may not be known for the experiment.

The latter offers the opportunity to identify previously unknown regions that are subject to differential methylation. Included in this is the opportunity to discover the scale over which differential methylation acts. However, valid statistical inference of these regions is far more challenging.

\subsection{Using \em{a priori} regions}

The idea of testing for differential methylation is relatively straightforward. For each sample, the region-wise methylation level is some summary of the methylation level at each loci in that region, for example, the weighted average of $\beta$-values for all loci in the region, where the weights are proportional to the sequencing coverage. These averages are then compared between samples. I describe a basic implementation of this for a one-group experiment and then detail a more sophisticated method that is implemented in `BiSeq` \citep{Hebestreit:2013ko}.

\subsubsection{One-group experiments}\label{sec:one_group_dmrs}

For a region containing $c$ methylation loci, construct the $n \times 2 \times c$ contingency table, i.e., each 'slice' of the table is a $n \times 2$ table containing the data for a single locus. Simply aggregating the counts across loci and testing the resulting $n \times 2$ table would be incorrect, since such an aggregating could introduce spurious differences in the proportions \citep{Good:1987ce}. Instead, we could fit the log-linear model with `sample`, `meth_state` and `locus` as the set of factors. The saturated model is

\begin{align*}
\log(F_{a, b, c}) &= \lambda_{0} \\
  &+ \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} \\
  &+ \lambda_{a, b}^{meth\_state, sample} + \lambda_{a, c}^{meth\_state, locus} + \lambda_{b, c}^{sample, locus} \\
  &+ \lambda_{a, b, c}^{meth\_state, sample, locus}
\end{align*}

We would first need to test whether there is a significant three-way interaction between `meth_state`, `sample` and `locus` by computing the deviance of the model without the three-way interaction

\begin{align*}
\log(F_{a, b, c}) &= \lambda_{0} \\
&+ \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} \\
&+ \lambda_{a, b}^{meth\_state, sample} + \lambda_{a, c}^{meth\_state, locus} + \lambda_{b, c}^{sample, locus}
\end{align*}

On the one hand, if there is no three-way interaction, then we can test for an interaction between `meth_state` and `sample`, i.e., test for a difference in methylation level across the loci, by comparing the model

\begin{align*}
\log(F_{a, b, c}) &= \lambda_{0} \\
&+ \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} \\
&+ \lambda_{a, b}^{meth\_state, sample} + \lambda_{a, c}^{meth\_state, locus} + \lambda_{b, c}^{sample, locus}
\end{align*}

to the model

\begin{align*}
\log(F_{a, b, c}) &= \lambda_{0} \\
&+ \lambda_{a}^{meth_state} + \lambda_{b}^{sample} + \lambda_{c}^{locus} \\
&+ \lambda_{a, c}^{meth\_state, locus} + \lambda_{b, c}^{sample, locus}
\end{align*}

This test is analogous to the generalised Cochran-Mantel-Haenszel test \citep{Landis:1978vk,MANTEL:1978iv,Birch:1965uj} of the $n \times 2 \times c$ table \citep[p231]{agresti1990categorical}. The advantage of the Cochran-Mantel-Haenszel is its applicability for sparse datasets for which the asymptotic theory does not hold for the test based on the log-linear regression  \citep[p232]{agresti1990categorical}.

On the other hand, if there is a significant three-way interaction between `meth_state`, `sample` and `locus` then this requires more careful modelling to determine the nature of this interaction and whether the region is a consistent DMR.

\subsubsection{A more sophisticated multi-group analysis}\label{sec:a_more_complex_example}

`BiSeq` is designed for RRBS data and tests for differential methylation at "CpG clusters". These CpG clusters are based on an initical scan of the genome looking for CpGs that are within a given distance of one another and that have sufficient sequencing coverage across the set of samples. While there is clearly a data-dependent component to these cluster definitions, I reserve the use of 'data-driven regions' for those which are based on the methylation levels of loci rather than their genomic co-ordinates. \citep{Hebestreit:2013ko} alternatively recommend using the target regions of the assay, such as MspI fragments in the case of RRBS, instead of creating these CpG clusters.

One the clusters are defined, the $\beta$-values in each cluster are smoothed using a local binomial likelihood smoother. For each CpG `BiSeq` fits a beta regression model[^not_beta-binomial], which is tested for evidence of differential methylation at that cytosine. Based on these P-values, `BiSeq` then use a hierarchical testing procedure to control the false discovery rate at both the cluster-level and locus-level. This method is based on several papers by Yoav Benjamini and colleagues \citet{Benjamini:1997fh, Benjamini:2001ho, Benjamini:2006gd, Benjamini:2007cm}. It aims to first control the false discovery rate at the cluster-level and then refine the signal by trimming non-DMCs from those clusters that have been declared as differentially methylated. Finally, these differentially methylated clusters are _post hoc_ filtered to ensure they are _consistent_ (see Section \ref{sec:consistency_of_dmrs}).

[^not_beta-binomial]: This is different to the beta-binomial regression framework described in Section \ref{sec:beta-binomial_models}.

Based on my understanding of this procedure, only cytosines within differentially methylated clusters can be declared DMCs since it is only those cytosines that are tested in the second-step of the hierarchical testing algorithm. Therefore, isolated DMCs will not be detected by this procedure.

\subsection{Using data-driven regions}

The most common type of algorithm for identifying data-driven DMRs, also arguably the most _ad hoc_, is based on detecting clusters of DMCs \citep[e.g.,][]{Lister:2009hy, Lister:2011kg, Hansen:2011gu, Feng:2014iq}. The initial identification of DMCs will typically use a relaxed statistical significance threshold. For example, \citet{Hansen:2011gu} use all CpGs with a P-value in the lowest $5\%$ and declare putative DMRs to be contiguous runs of such CpGs that are within a given distance of one another and with "all differences in the same direction". This latter requirement is what I call the _consistency_ of the DMR (see Section \ref{sec:consistency_of_dmrs}).

These putative DMRs may be subject to further filtering, such as requiring that they contain a minimum number of CpGs and span a minimum number of bases, and the merging of nearby putative DMRs into a single putative DMR \citep{Hansen:2011gu}. Notably, many of these methods do not include a formal statistical test of differential methylation at the region-level \citep[e.g.,][]{Lister:2009hy, Lister:2011kg, Hansen:2011gu, Feng:2014iq}.

The challenges of valid statistical inference at such data-driven regions are not unique to the problem of testing for differentially methylated regions. Similar problems arise in the analysis of chromatin immunoprecipitation sequencing (ChIP-seq) experiments \citep{Schwartzman:2011to, Lun:2014gn} and in the field of signal processing \citep{Schwartzman:2011vg}. It boils down to the problem of how to assign statistical significance to regions or peaks using the same data that were used to define these regions or peaks. These regions are already enriched for differences, which biases the interpretation the results of tests for differences at these regions. One way, and I would argue the best way, to avoid this issue is to test these regions using a separate dataset, which completely avoids the issue of statistical 'double-dipping'. If the sample size is large enough, then permutation testing may also be appropriate. For example, \citet{Hansen:2013eo} permute the group labels of their samples and re-ran their analysis to determine for each of the observed regions "how often we see another block of similar length and effect size anywhere in the genome and in any of the permutations". The chief limitation of the permutation strategy is the restricted number of permutations that are possible from small sample sizes and the time and computational resources it takes to perform analyse the data for a single permutation, which are often substantial. For example, \citet{Hansen:2013eo} only performed nine permutations.

A somewhat less _ad hoc_ strategy for identify differentially methylated regions is called "bump-hunting" \citep{Jaffe:2012gx}. Initially developed for methylation microarrays, and now available for broader use in the R/Bioconductor package `bumphunter` ([http://www.bioconductor.org/packages/release/bioc/html/bumphunter.html](http://www.bioconductor.org/packages/release/bioc/html/bumphunter.html)), bump-hunting is quite a general statistical problem. For the case of identifying DMRs, the idea is to plot the value of a statistic used to test for DMCs as a function of the position in the genome[^smoothing] and then to identify bumps in this plot. Bumps are identified as contiguous regions of the genome where the signal is above some threshold and may include an error term to account for the spatial correlation of the signal. The significance of these peaks may be assessed with a permutation strategy.

Another alternative for combining individual loci into data-driven DMRs is based on the loci-specific P-values. These methods can be thought of extensions to Fisher's method for combining P-values \citep{Fisher:1936vv}, by attempting to account for the correlation of tests at nearby methylation loci. `comb-p` \citep{Pedersen:2012vl} uses the Stouffer-Liptak-Kechris \citep{Stouffer:1949ua, Kechris:2010fw, Zaykin:2011jq} correction for spatially correlated P-values. `SLIM` \citep{Wang:2011cw} is used by `methylKit` \citep{Akalin:2012cm} to do a similar correction.

Finally, there are a class of methods that turn the problem of identifying data-driven DMRs on its head by constructing the regions without first testing the individual loci for differential methylation. Then, only once these regions are defined, these methods test for differential methylation. This is different to using _a priori_ regions, since the regions are still data-defined, but not with respect to differential methylation. The only example I know of such a method is `Aclust` \citep{Sofer:2013bk}. `Aclust` first clusters CpGs into candidate regions by performing agglomerative nested clustering of the between-sample co-methylation. Briefly, this is the correlation of methylation levels at two loci across the samples[^between_sample_co-methylation]. These clusters are then tested for differential methylation using generalised estimating equations.

[^between_sample_co-methylation]: See Chapter \ref{chap:co-methylation_review} for further details of between-sample co-methylation.

\subsubsection{Consistency of DMRs}\label{sec:consistency_of_dmrs}

While not often explicitly stated, it is desirable that a DMR is _consistent_. The definition of consistency is best illustrated by a pathological example describing a region that is inconsistent. Suppose we have a one-group experiment with five samples[^one_by_five] and are studying a region with five CpGs, all of which are DMCs. However, the first CpG is differentially methylated in the first sample, the second CpG is differentially methylated in the second sample, etc. While the entire region is 'differentially methylated', the difference is not observed consistently in any one sample. A consistent region is therefore one where the differences are 'in the same direction'.

[^one_by_five]: Equivalently, a five group experiment without replicates.

Consistent DMRs are most easily identified in two-group experiments, where we can use the sign of the test statistic to determine the 'direction' of the signal. This is exactly what is done by, for example, \citet{Hansen:2011gu}.

In one-group experiments, the idea of consistency is similar, but not quite identical, to testing the homogeneity of odds ratios for all $n \times 2$ contingency tables in the region. For example, a region may be consistent but the strength of the effect may vary across the region, thus the odds ratios may vary.

\section{Other downstream analyses using $\beta$-values}

Two other common downstream analyses based on $\beta$-values are the identification of variable methylation and partially methylated domains.

\subsection{Variable methyation}\label{sec:variable_methylation}

Differential methylation is an analysis of the difference in means and seeks to identify differentially methylated regions (DMRs). Analogously, we can analyse differences in the variability of methylation profiles, which is known as differential variability, and those regions known as variably methylated regions (VMRs). It has been hypothesised that increased variability in DNA methylation indicates an epigentic _plasticity_ that may be highly relevant to common diseases, in particular, cancer \citep{Feinberg:2010dy}.

In a one-group experiment, a variably methylated region is one that has increased variability compared to 'similar' regions elsewhere in the genome. In a two-group experiment, a differentially variable locus is one where the variation in one group is significantly larger than that in the other group. \citet{Jaffe:2011hy} first developed formal statistical tests for both the one-group and two-group experiments. These methods were developed for us with data from the CHARM array (see \ref{sec:enzyme_digestion_assays}).

Statistical tests of variations are well known to be more difficult than tests of means and require larger sample sizes. A more subtle difficulty is in dealing with outliers. An outlier in one group will greatly increase the variation in that same group, but this does not necessarily mean that the locus is differentially methylated. It is not hard to envisage an example where the two groups in fact have very similar variability once the outlier is excluded. Tests of differential variability that are based on the F-test \citep[e.g.,][]{Hansen:2011gu} or Bartlett's test \citep[e.g.,][]{Teschendorff:2012hea} will be susceptible to calling such loci as differentially variable. By contrast, \citet{Phipson:2014ji} designed `DiffVar` to test for differential variability that is deliberately robust to such outliers by using Levene's test \citep{Olkin:1960uu}. Which approach is preferable will depend on the user's perspective as to whether such loci with outliers should be called as variably methylated.

\subsection{Partially methylated domains}

Partially methylated domains, PMDs, are long stretches of the genome where the average methylation level, $B_{i}$, is away from the extremes of $0$ and $1$, typically in the range $0.20 - 0.70$. PMDs were first identified in the IMR90 genome by \citet{Lister:2009hy}. Using a simple sliding window algorithm, \citeauthor{Lister:2009hy} unexpectedly found that approximately $40\%$ of every autosome was a PMD and that the average length of these PMDs was a large 153 kb. They also showed that these PMDs were not simply due to a methylated population and an unmethylated subpopulation of cells in the sample. This did this by showing that inidividual reads mapped to these PMDs contained both methylated and unmethylated bases.

A subsequent study found that PMDs are a common feature of somatic cell lines and that they comprise $> 30\%$ of the genome \citep{Lister:2011kg}. Perhaps even more intriguingly, across four somatic cell lines profiled with whole-genome bisulfite-sequencing, \citeauthor{Lister:2011kg} found a large amount of these genomes ($664$ Mb) comprised shared PMDs.

PMDs have also been identified within tumour methylomes \citep{Berman:2012ga, Hansen:2011gu} and \citet{Hansen:2011gu} found that these overlap with other important genomic features called large organized chromatin lysine modifications (_LOCKs_) and lamina associated domains (_LADs_). However, PMDs are conspicuous by their absence in pluripotent cell lines, including both embryonic stem cells and induced pluripotent stem cells \citep{Lister:2011kg}. Their absence in the induced pluripotent cell lines may reflect the fact that the methylome is 'reset' upon induction of pluripotency \citep{Lister:2011kg}.

More sophisticated methods have also been proposed to identify these PMDs. `methylSeekR` \citep{Burger:2013kq} is one such algorithm. It uses a hidden Markov model of the $\beta$-values to segment the genome into unmethylated, lowly methylated and partially methylayed regions. While `methylSeekR` is based on $\beta$-values and does not make use of individual reads to identify PMDs, it does require the high spatial resolution of $\beta$-values that whole-genome bisulfite-sequencing provides

\section{Downstream analyses of methylation patterns at m-tuples}\label{sec:downstream_analyses_of_m-tuples}

None of the aforementioned analyses specifically exploit the fact that bisulfite-sequencing reads measure methylation patterns at multiple loci, i.e., at m-tuples. Yet much can be learnt my studying exactly that.

In this section I review the different types of questions that can be addressed when we make use of this extra information that is only available from bisulfite-sequencing data.

\subsection{Methylation entropy}\label{sec:methylation_entropy}

There are several definitions and uses of _methylation entropy_ in the literature. \citet{Xie:2011cy} introduced the concept for bisulfite-sequencing data as a way of quantifying the variability of methylation patterns at an m-tuple. For a given 4-tuple, $(i, \ldots, i + 3)$, \citet{Xie:2011cy} analyse the set of reads aligned to it, $\{z: z \in \mathcal{R}_{(i, \ldots, i + 3} \}$, to compute the methylatione entropy. On the one hand, if we only observe a single unique methylation pattern then the m-tuple has the minimum methylation entropy of zero. On the other hand, if we observe all possible $2^{4} = 16$ methylation patterns then the m-tuple has the maximum methylation entropy of 1. Depending on the frequency of the patterns, we obtain intermediate methylation values ($0 < \text{methylation entropy} < 1)$. Of course, methylation entropy need not be limited to 4-tuples. The choice of the size of the tuples depends on the regions of interest, the read lengths and the sequencing coverage.

The natural interpretation of methylation entropy as a measure of 'disorder' allows us to quantify how heterogeneous DNA methylation is at a locus \citep{Xie:2011cy, He:2013cj}[^other_definitions]. These methylation entropies can be analysed to identify heterogeneous regions of the genome or perhaps tested for an association with a phenotype.

[^other_definitions]: This is closely related to the idea of identifying epialleles, for which methylation entropy has also played a role \citep{Li:2014ei} and which I discuss in Section \ref{sec:epialleles_and_epipolymorphism}. Methylation entropy has also been used to identify differential methylation, however, I do not discuss this further since this uses a different definition that is not based on analysing methylation patterns at m-tuples \citep{Zhang:2011dp, Su:2012hl}.

While \citet{Xie:2011cy} do not provide software implementing the method, a subsequent paper from some of the same researchers introduces the `DMEAS` software to do so \citep{He:2013cj} . Unfortunately, `DMEAS` is only available as a Windows binary or as a Perl script that itself is only available as PDF file, greatly limiting its usefulness.

\subsection{Allele-specific methylation}\label{sec:asm}

In a diploid cell, allele-specific methylation occurs when only one of the parental chromosomes is methylated at a particular locus, where the locus may be an individual cytosine or a broader region. One particular form of allele-specific methylation occurs at imprinted genes where the silenced copy of the gene is heavily methylated and the active copy is lightly methylated \citep{Shoemaker:2010jz}. However, it is now apparent that allele-specific methylation is far more prevalent than at just these imprinted regions \citep{Tycko:2010hc, Shoemaker:2010jz}.

The obvious method to detect allele-specific methylation from bisulfite-sequencing relies on the read containing a heterozygous genetic variant, such as a single nucleotide polymorphism, along with at least one methylation locus. The heterozygous variant allows the reads to be separated by the observed allele[^phase], which can then be used to test for allele-specific methylation. For example, \citet{Shoemaker:2010jz} construct a $2 \times 2$ contingency table, like that shown in Table \ref{tab:asm}, and test for an association between the allele and the methylation state using Fisher's exact test \citep{Fisher:1922bx}.

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
& Allele 1 & Allele 2 \\ \midrule
m &  m_{1}        &  m_{2}       \\
u &  u_{1}        &  u_{2}       \\ \bottomrule
\end{tabular}
\caption{$2 \times 2$ table used to test for allele-specific methylation.}
\label{tab:asm}
\end{table}

Such an approach, while straightforward, is also obviously limited to a subset of the available methylation loci. Both \citet{Fang:2012ci} and \citet{Peng:2012dh} published methods to detect allele-specfic methylation that do not require nearby heterozygous genetic variants nearby to the methylation locus of interest. These methods rely on the probabilistic assignment of reads to alleles, which are treated as missing data. Unfortunately, there is no publicly available software implementing the method proposed by \citet{Peng:2012dh} and so I do not discuss it further.

\citet{Fang:2012ci} uses reads containing multiple methylation loci and looks for regions of the genome where there are two distinct methylation patterns at the read-level that occur at roughly equal proportions, indicating one pattern comes from one allele and the other pattern from the other allele. The likelihood of allele-specific methylation is computed using an expectation-maximisation algorithm, which assigns reads to one of the two possible alleles. Neighbouring regions displaying allele-specific methylation are then joined together. While not mentioned in the paper, the proposed method is now available in the `MethPipe` software ([http://smithlabresearch.org/software/methpipe/](http://smithlabresearch.org/software/methpipe/)).

[^phase]: This does not give parent-specificity unless the phase of the genotype is also known.

\subsection{Epialleles and epipolymorphism}\label{sec:epialleles_and_epipolymorphism}

A DNA sequence may have multiple epigenetic states. For example, the cytosine in the sequence `TCGA` may be methylated or unmethylated; each of the methylated and unmethylated versions of that sequence is an _epiallele_. Restricting our attention to CpG methylation, a sequence with $m$ CpGs has $2^{m}$ potential epialleles. Therefore, an epiallele is just a methylation pattern at an m-tuple, with the additional constraint that the underlying DNA sequence also be identical[^epiallele_definition]. An epiallele may also be described as an _epimutation_, if it sufficiently rare.

[^epiallele_definition]: \citet{Rakyan:2002wz} impose a very strict requirement that in order for the methylation pattern to be called an epiallele it must "[result] in different phenotypes". While most would agree that is those methylation patterns that confer a difference in phenotype that are most interesting, limiting the definition of 'epiallele' to such methylation patterns seems unduly restrictive. After all, we refer to alternate forms of a genetic sequence as alleles regardless of whether we know of a phenotypic consequence of the variant.

Epipolymorphism, analogous to genetic polymorphism, is the presence of multiple epialleles in a population. However, unlike genetic polymorphism, where the population is typically a set of chromosomes from multiple individuals, the population here is often within an individual, even within a tissue within an individual. The 'epipolymorphism' of a locus has also been defined as the probability that two epialleles randomly sampled from the locus are different from one another \citep{Landan:2012kp}. \citeauthor{Landan:2012kp} define the epipolymorphism of an m-tuple as $1 - \sum_{p = 1}^{p = 2^{m}} f_{p}^{2}$, where $f_{p}$ is the estimated frequency of the $p^{th}$ methylation pattern, i.e., the number of times the $p^{th}$ pattern is observed divided by the total number of reads mapped to that m-tuple[^sampling_with_replacement].

[^sampling_with_replacement]: Strictly speaking, this is in fact an estimate of the epipolymorphism under a model that assumes sampling with replacement or, equivalently, an infinite population size. While neither assumption is true, the correction for sampling without replacement from a finite population will not substantially affect the results.

\subsubsection{Examples}

Most of the examples of epialleles with a phenotypic consequence come from the plant kingdom. Even then, the number of such epialleles is small --- a review from 2012 put the number at "about a dozen" \citep{Weigel:2012fj}.

In mammals, the study of epialleles has focused on identifying _metastable_ epialleles, which are epialleles that are mitotically heritable \citep{Rakyan:2002wz}. The poster child for the potential important of epialleles in mammals is the Agouti viable yellow ($A^{vy}$) allele \citep{Morgan:1999ds}.  Mice with an unmethylated version of the allele have a yellow coat, are obese, diabetic and have an increased suscpetibility to tumours compared to those mice with a methylated version of the allele, which have a _pseudoagouti[^pseudoagouti_vs_agouti]_ (brown) coat and none of the associated health defects.

[^pseudoagouti_vs_agouti]: These mice are properly described as _pseudoagouti_ rather than agouti. They are heterozygous for the wildtype agouti allele ($A^{vy}/a$) but are phenotypically indistinguishable from true agouti mice, which are homozygous for the wildtype gene ($a/a$).

In humans, there have been several interesting studies using epialleles to infer the clonality and evolution of cancer \citep{Siegmund:2009kk, Li:2014ei}, as well as the evolution of methylation dynamics and the epipolymorphism of various loci in an immortalised cell line \citep{Landan:2012kp}.

Finally, although there is much interest, if not outright hype, surrounding transgenerational epialleles, these are difficult to prove outside of very controlled experimental conditions \citep{Daxinger:2012kl}.

Regardless of where you draw the line as to what constituted an epiallele, it has become clear that the occurrence of multiple methylation patterns at an m-tuple is the norm rather than the exception when performing bisulfite-sequencing. For example, __FIGURE__ shows the number of methylation patterns observed at 4-tuples with a frequency greater than __WHAT__ from the __WHICH__ sample. There has been some progress in developing software to extract and analyse these epialleles, which I now describe. (Alternatively, cite Landan et al.)

\subsubsection{Methods}

`methclone` \citep{Li:2014ei} is software to estimate the frequency of epialleles at m-tuples and to identify "shifts" in these distributions between two samples. `methclone` is limited to pairwise comparisons. The input is a pair of `BAM` files created with the `Bismark` aligner, one for each sample. For each m-tuple (4-tuples in the example in the paper) with sufficient sequencing coverage ($60 \times$ in the example in the paper) a form of methylation entropy, called combinatorial entropy, is computed based on the set of mapped reads for the pair of samples[^methclone_output]. The "foreground" combinatorial entropy, $S$, is based on the observed frequency of epialleles in the two samples while the "background" combinatorial entropy, $\tilde{S}$, is the expected frequency of epialleles in the two samples if "all patterns of epialleles are uniformly mixed between the two stages [samples]". The difference in these combinatorial entropies, $\Delta S = S - \tilde{S}$, a kind of observed-to-expected log-ratio, is used to identify shifts in the epiallele distribution between a pair of samples. A $\Delta S = 0$ corresponds to no change and a $\Delta S = -144$ corresponds to maximal difference in entropy. It isn't clear whether the range of $\Delta S$ depends on the size of the m-tuples nor is it clear how to choose the threshold at which to declare a significant shift in the distribution of epialleles. The example in the paper shows results for a threshold of $-60, -70$ and $-80$, but these choices appear somewhat arbitrary.

[^methclone_output]: It is not clear from the paper or documentaiton whether `methclone` returns a file containing the number of methylation patterns at each m-tuple, _a ala_ `methtuple`. Unfortunately, I haven't been able to install `methclone`, so have not been able to test this out.

The obvious challenge in estimating the frequency of an epiallele is in distinguishing a _real_ epiallele from a spurious one, perhaps caused by incomplete bisulfite-conversion, sequencing error or mapping error. Another difficulty, perhaps unavoidable with the present technology, is the effect of PCR amplification bias, which will bias estimates of the relative abundance of each epiallele. `methclone` uses the observed frequencies of methylation patterns as being unbiased estimates of the true epiallele frequencies, which does not attempt to account for these potential sources of bias. In contrast, `MPFE` ([http://www.bioconductor.org/packages/release/bioc/html/MPFE.html](http://www.bioconductor.org/packages/release/bioc/html/MPFE.html), [http://f1000.com/posters/browse/summary/1097258](http://f1000.com/posters/browse/summary/1097258)), an R/Bioconductor package for "[estimating] the distribution of methylation patterns [epialleles]" at m-tuples, uses a probabilistic model to account for some of these biases.

`MPFE` is designed to estimate the frequency of epialleles by maximising a multinomial likelihood that includes error terms for both incomplete bisulfite-conversion and sequencing error. The maximisation of this likelihood is computatationally demanding, as evidenced by the need for a "fast" algorithm that approximates the likelihood. `MPFE` is designed for amplicon bisulfite-sequencing and may not scale to whole-genome data. The input is a file containing the number of times each methylation pattern was observed at that m-tuple. Unfortunately, `MPFE` does not provide a way to create this table.

Methods designed to detect allele-specific methylation, specifically those that are based on the observed methylation patterns \citep[e.g.,][]{Fang:2012ci, Peng:2012dh}, might also adapted to identify epialleles and their associated frequencies. It is worth emphasising that since all of the methods described in this section are based entirely on the observed methylation patterns, none of these actually check that the underlying DNA sequencing is identical, which strictly speaking is a requirement for the m-tuple to be an epipolymorphic locus.

\section{\texttt{MethylationTuples}}\label{sec:MethylationTuples}

A common thread in Section \ref{sec:downstream_analyses_of_m-tuples} is that while interesting methods have been proposed to address these questions, the software implementations are often lacking or, at best, very much tailored toward a particular analysis. What has been lacking is:

1. A general method for extracting methylation patterns at m-tuples,
2. Software for manipulating, analysing and visualising these methylatoin patterns.

I have made significant progress towards the first goal with `methtuple` (see Section \ref{sec:methtuple}) and now introduce `MethylationTuples` ([https://github.com/PeteHaitch/MethylationTuples](https://github.com/PeteHaitch/MethylationTuples)) to address the second missing link.

\subsection{Design}

`MethylationTuples` is an R package for managing, analysing and visualising methylation patterns at m-tuples released under a Artistic-2.0 licence. R is a very popular language for analysing data, particularly in bioinformatics, and is my computational mother tongue, hence my choice of R as the language. To improve the performance of the package, parts are written in `C++`, making use of the wonderful `Rcpp` package \citep{Eddelbuettel:2011to,Eddelbuettel:2013if}.

While initially developed to support my research into co-methylation (Chapter \ref{chap:co-methylation}), the data structures developed in `MethylationTuples` are also well-suited to analysing methylation entropy, allele-specific methylation, epialleles and epipolymorphism.

`MethylationTuples` is written to work with the Bioconductor project \citep{Gentleman:2004ju}[^bioc]. Bioconductor makes extensive use of R's S4 object system and encourages developers to re-use existing Bioconductor infrastructure. From a developer's perspective, this avoids the need to re-invent the wheel. And from the user's perspective, it helps avoid multiple versions of the 'wheel', each that acts slightly differently and that may not be as well-tested.

[^bioc]: `MethylationTuples` has not yet been submitted to Bioconductor but its development is being published to [https://github.com/PeteHaitch/MethylationTuples](https://github.com/PeteHaitch/MethylationTuples).

Bioconductor has excellent support for working with data defined on genomic ranges via the `IRanges` and `GenomicRanges` packages \citep{Lawrence:2013hi, Lawrence:2014gy}. However, genomic tuples, such as the co-ordinates of an m-tuples, do not naturally fit into this framework[^tuples_vs_ranges]. Therefore, I wrote a Bioconductor package for working with genomic tuples, rather unimaginatively called `GenomicTuples`, first released as part of Bioconductor version 3.0 ([http://www.bioconductor.org/packages/3.0/bioc/html/GenomicTuples.html](http://www.bioconductor.org/packages/3.0/bioc/html/GenomicTuples.html)). In fact, `GenomicTuples` is heavily based on the `GenomicRanges` package, with modifications for tuple-specific operations. For example, there is a tuple-specific method for the `findOverlaps` generic function to identify genomic tuples with equal co-ordinates (i.e., `type = \`equal'`). Since the classes in `GenomicTuples` extend those defined in `GenomicRanges`, these have excellent interoperability with existing Bioconductor infrastructure.

[^tuples_vs_ranges]: The difference between a genomic range and a genomic tuple  can be thought of as the difference between an interval and a set. Namely, an interval includes the co-ordinates in between the start and end whereas a set only includes those co-ordinates listed in the set. For example, the genomic interval `chr3:+:[10, 12]` includes the co-ordinates `chr3:10`, `chr3:11` and `chr3:12` on the forward strand, whereas the genomic 2-tuple `chr3:+:{10, 12}` only includes the co-ordinates `chr3:10` and `chr3:12` on the forward strand.

In the `MethylationTuples` package, I define the `MethPat` class to store the genomic co-ordinates of m-tuples and the associated frequencies of the methylation patterns. The `MethPat` class extends the `SummarizedExperiment` class, also defined in the `GenomicRanges` package, but makes use of classes defined in the `GenomicTuples` package to store the genomic co-ordinates of the m-tuples. A `MethPat` object is as a matrix-like object, where rows represent m-tuples of interest and columns represent samples. Currently, it is a requirement that all m-tuples in a `MethPat` object have the same size ($\text{m}$). Figure \ref{fig:MethPat} is a schematic of a `MethPat` object storing data on $n$ samples for 3-tuples.

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/MethPat.pdf}
\caption{Schematic of the \texttt{MethPat} class, shown here for 3-tuples. Each row represents a 3-tuple to which the genomic co-ordinates of the tuples (green box) and the frequencies of the methylation patterns (grey box) are aligned. The frequencies of each methylation pattern --- $MMM, MMU, \ldots, UUU$ --- are stored as separate matrices where the columns represent samples --- $S_{1}, \ldots, S_{n}$. Some samples may have now have any sequencing coverage for a particular m-tuple, in which case the corresponding frequencies are recorded as \texttt{NA}.}
\label{fig:MethPat}
\end{figure}

\subsection{Available methods}

The similarities to the output format of `methtuple` are clear[^methtuple_output], with the added advantage that a single `MethPat` object can contain data from multiple samples. A `MethPat` object can be constructed using the `MethPat` constructor function or from the output files of `methtuple` via the `readMethtuple` function.

[^methtuple_output]: See Section \ref{sec:methtuple_methods} for example of the `methtuple` output format.

The `MethPat` object provides fast subsetting by rows (m-tuples), including subsetting based on overlaps of m-tuples with genomic features via the `findOverlaps` methods. Several other useful utility functions are:

- `collapseStrand`: Collapse strand-specific data by aggregating the counts. Only applicable to CpG methylation loci.
- `combine`: Combine multiple `MethPat` objects into one.
- `filterOutVariants`: Remove m-tuples that contain a known variant. Variants must be provided as a VCF.
- `getCoverage`: Compute the sequencing coverage of the m-tuples in each sample.
- `IPD`: Compute the IPD vector for each m-tuple.
- `methLevel`: Compute $\beta$-values or $\mathcal{M}$-values. Only applicable to 1-tuples.
- `MBias` and associated methods and functions: Helper functions to perform a more systematic analysis of M-bias (see Section \ref{sec:identifying_m-bias}).
- `plotBetaDensity`: Plot the kernel density of the $\beta$-values for each sample, optionally stratified by a genomic feature, e.g., CpG islands.
- `tuples`: Extract the $pos_{1}, \ldots, pos_{m}$ of the m-tuples.

These are in addition to the many useful utility functions inherited from the `SummarizedExperiment` class.

In the `MethPat` class, its associated methods and other utility functions, the `MethylationTuples` package provides a toolbox for manipulating methylation patterns at m-tuples. Aside from providing the necessary infrastructure to analyse methylation patterns at m-tuples, `MethylationTuples` currently includes specific methods to analyse and visualise co-methylation (Chapter \ref{chap:co-methylation}) with the `cometh` and `methLevelCor` functions. I also plan to add methods for estimating epialleles and epipolymorphism. And it is my hope that `MethylationTuples` will provide a useful foundation on which others can implement their own methods for analysing methylation patterns at m-tuples.

\subsection{Compatability with other Bioconductor packages}

Being based on core Bioconductor functionality, `MethylationTuples` is highly compatible with existing packages. In particular, `MethPat` objects containing 1-tuples are readily coerced for use in differential methylation calling packages, e.g., `bsseq` and `BiSeq`, or to identify partially methylated domains with `MethylSeekR`. I also make extensive use of `MethylationTuples` in my `methsim` software (Chapter \ref{chap:methsim})

\subsection{Computational challenges and future directions}

The challenges of working with large datasets in R are well-known. These are in large part due to R being designed as an 'in-memory' application and it's implementation of 'copy-on-modify' semantics \citep{wickham2014advanced}. More generally with large datasets, there is often a trade-off to be made between storage efficiency and algorithm simplicity --- a more efficient way of storing the data may be less convenient to work with and _vice versa_.

I have used the `MethylationTuples` package to analyses datasets containing up to 17 whole-genome bisulfite-sequencing samples (the Lister data), albeit using a server with $512$ Gb of (shared) memory. I have found the `MethPat` class to be a very convenient representation of the data, but one that does not coincide with the most efficient storage. This is more of a problem as the size of the m-tuples increases and the data become sparser. However, for values of $m < 5$, and particularly for 'dense' data, such as RRBS, this is far less of an issue. Shown in Table \ref{tab:methpat_sparsity} is the size of the `MethPat` objects in memory for the EPISCOPE whole-genome bisulfite-sequencing data with various sized m-tuples. It also highlights the sparsity of the data as the size of the tuples grows --- most entries are $0$ (meaning that particular methylation pattern was not observed in that particular sample) or `NA` (meaning that that m-tuple was not observed in that particular sample).

\begin{table}[h]
\centering
\begin{tabulary}{\textwidth}{LLLLL}
\toprule
& \texttt{pryr::object\_size(x)} & \texttt{nrow} & Number of assays & Percentage of \texttt{NA} and $0$ values \\ \midrule
1-tuples                                 & $5.9$ GB                        & $56,348,522$    & $2$              & $28\%$                                     \\
2-tuples                                 & $20.1$ GB                        & $100,586,237$   & $4$              & $80\%$                                     \\
2-tuples (\texttt{--all-combinations}) & $60.0$ GB                          & $299,814,999$   & $4$              & $78\%$                                     \\
3-tuples                                 & $43.3$ GB                        & $109,376,348$   & $8$              & $93\%$                                     \\
4-tuples                                 & $80.5$ GB                        & $102,625,758$   & $16$             & $97\%$                                       \\ \bottomrule
\end{tabulary}
\caption{Size of \texttt{MethPat} objects for the \em{EPISCOPE} data. All m-tuples are stranded. The size of the \texttt{MethPat} object is computed using the \texttt{object\_size} function in the \texttt{pryr} package (\url{http://cran.r-project.org/web/packages/pryr/index.html}). The number of rows (\texttt{nrow}) corresponds to the number of m-tuples in the object. The number of assays is $2^{\text{m}}$, where $m$ is the size of the m-tuples. The final column is a measure of how sparse the data are --- a $0$ values means that particular methylation pattern was not observed in that particular sample and an \texttt{NA} value means that that m-tuple was not observed in that particular sample.}
\label{tab:methpat_sparsity}
\end{table}

The main inefficiency with the `MethPat` class is that the matrices storing the frequencies of each methylation pattern grow increasingly sparse as the size of the tuples increases. It may be possible to improve this by using a different approach to the internal storage of the data. For example, it may be possible to create an index so that the matrices storing the frequencies of methylation patterns can be re-ordered and stored as run-length encodings, a very space-efficient storage scheme.

In the short term, my aim for `MethylationTuples` is to release a version to Bioconductor with the core infrastructure for manipulating data of methylation patterns at m-tuples. In the longer term, I would like to extend the set of downstream analyses of m-tuples that are available in the package and for other users to do the same with their own methods.