### 2015-02-18 Rmarkdown/80-word_summary.Rmd
Removed markdown files or converted them to Markdown

---
title: "80-word summary"
output:
  pdf_document:
    fig_caption:
    - yes
    includes:
      in_header:
      - ../latex/header.tex
  html_document:
    keep_md:
    - yes
bibliography: ../latex/phd_thesis.bib
---

# The statistical analysis of data from high-throughput assays for studying DNA methylation

I develop a statistical framework for modelling and analysing DNA methylation data. In particular, I focus on the complex dependence structure of these data. I describe how this can be incorporated into statistical analyses of bisulfite-sequencing (BS-seq) experiments and develop software to simulate realistic BS-seq data by incorporating this dependence.

I also report results from my analysis of more than fifty BS-seq samples, both publicly available and those collected by collaborators, which have informed the methods developed in my thesis.