### 2015-02-18 Rmarkdown/TODO.Rmd

---
title: "TODOs"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

Those of highest priority are tagged, __URGENT__.

## PhD handbook

TODOs according to the [Unimelb PhD handbook](http://gradresearch.unimelb.edu.au/handbooks/phd/) (as of 07/10/2014).

- [ ] Read [Preparation of Graduate Research Theses Procedure (MPF1263)](https://policy.unimelb.edu.au/MPF1263).
- [ ] Read [Examination of Graduate Research Students Procedure (MPF1262)](https://policy.unimelb.edu.au/MPF1262).
- [ ] Read [Examination of Graduate Research Students Policy (MPF1207)](https://policy.unimelb.edu.au/MPF1207).

### [Copyright](http://gradresearch.unimelb.edu.au/handbooks/phd/research.html#copyright)

> Before you contemplate publication of your thesis either online or in print, it is important that you understand both your rights and obligations under copyright. Some publishers require that you sign an author or publishing agreement. Before signing any agreement, you should make sure that you fully understand the terms and your rights under the agreement. __If your thesis includes copyright material created by other people, you must ensure that you have the right to publish this material. While the fair dealing provisions of the Copyright Act allow you to reproduce a limited amount of third party copyright works in your thesis for the purpose of study and research, you cannot rely upon these provisions when you publish the work to the world. If you intend to publish your work, you will generally need to seek permission from the copyright owner as there are only limited provisions in the Copyright Act that allow material to be published without permission.__ Visit the University's Copyright website for more information.

- [ ] What does this mean for the use of figures published in articles, textbooks or on websites?
  - [ ] Further details on copyright [http://unimelb.edu.au/copyright/research.html](http://unimelb.edu.au/copyright/research.html)

### [Ethics approval](http://gradresearch.unimelb.edu.au/handbooks/phd/research.html#ethics)

> __All projects requiring the use of human participants, animals, genetically modified organisms or hazardous biological agents are subject to review and approval by the appropriate University committee prior to commencement of the work. Approval will not be granted retrospectively.__ Further information about research ethics and the requirements of the Human Research Ethics Committee, Animal Welfare Committee or Gene Technology and Biosafety Committee is available in departments or from the Office of Research Ethics and Integrity.

- [ ] Clarify what, if any, ethics statement I need in my thesis.
  - I am not directly performing any experiments as part of my PhD. However, I am using data from experiments that would require such ethics approval. In particular, the collaborations with Sue Clark (human samples) and Emma Whitelaw (mouse samples).

### [Editing of theses](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#editing)

> The use of third party editorial assistance, either paid or voluntary, must be acknowledged in the Preface and is limited to the guidelines adopted by the University.

- [ ] Acknowledge all proofreaders in Preface.

### [Completion seminar](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#completion)

> Completion of the seminar requirement will be signed off by the supervisors and head of department on the Completion Seminar section of the [Submission of Thesis: Statement by Candidate, Supervisor and Chairperson of Examiners](http://gradresearch.unimelb.edu.au/docs/exams/GS6B_Submission_Form.pdf) form and will include the date and location of the completion seminar.

- [ ] Complete [Submission of Thesis: Statement by Candidate, Supervisor and Chairperson of Examiners](http://gradresearch.unimelb.edu.au/docs/exams/GS6B_Submission_Form.pdf)

### [Word limit](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#wordlimit)

> Candidates should aim to write a thesis of 80,000 words.

> The word limit is exclusive of words in tables, maps, bibliographies and appendices. Footnotes are included as part of the word limit. Appendices must be limited to supporting material genuinely subsidiary to the main argument of the thesis.

> Candidates may write up to 100,000 words without seeking special permission from the RHD Committee.

- [ ] Find a `wordcount` function for LaTeX or markdown files.

### [Preparation of a thesis that contains published work](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#published)

> Thesis chapters may be comprised of work that is written specifically for the thesis, reprints or other reproductions of published works or an adaptation of published work..

- I don't expect this be an issue for me as I am not re-using any published work.

> Publications, such as reprints of journal articles, published creative writing and catalogues and documentation of public performances or exhibited work, may also be included in an appropriate form in the appendix, __including for example as DVD, CD or URLs, if not included as a chapter__.
- [ ] Include all URLs of software in appendix.

### [Format of the thesis](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#format)

> A thesis should include the following formatting:
* International Standard Paper Size A4 (297 x 210mm)
* 1.5 spacing and presented in a clear and legible font and would normally be expected to be double-sided
* Left and right margins of no less than 30mm and page numbers that appear inside the margins
* Pages that are numbered consecutively and clearly
* Folding diagrams or charts arranged so as to open to the top and right.

- [ ] Check the spacing and margins of my LaTeX pdf.

#### Order of contents

> A thesis follows the following order:
* Title page
* Abstract
* Declaration
* Preface (if applicable)
* Acknowledgments
* Table of contents
* List of tables, figures and illustrations (if list items are fewer than 10 in number, this is not necessary)
* Main text
* Bibliography or List of References
* Appendices.

- [ ] Check order of my thesis.

#### Title page

> A thesis must be preceded by a title page. The title page of the thesis should show the:
* Title of the thesis
* Full name of the author (as it appears in the Student Portal)
* Degree for which submitted (see below)
* Month and year
* Name of the department or faculty in which the research was carried out.

> Candidates who have pursued a course of study by research alone, including combined Masters/PhD degrees, shall state on the title page: "Submitted in total fulfilment of the requirements of the degree of Doctor of Philosophy".

- [ ] Check title page fits criteria.

#### Abstract

> The title page must be followed by an abstract of 300–500 words in English (in the case of creative arts the Abstract must include a description of the form and presentation of the creative work).

#### Declaration

> The following declaration page, signed by the candidate:

> This is to certify that:
* the thesis comprises only my original work towards the PhD except where indicated in the Preface,
* due acknowledgement has been made in the text to all other material used,
* the thesis is fewer than 100 000 words in length, exclusive of tables, maps, bibliographies and appendices OR the thesis is [number of words] as approved by the Research Higher Degrees Committee.

#### Preface

> If applicable, a Preface page includes a statement of:
* Work carried out in collaboration indicating the nature and proportion of the contribution of others and in general terms the portions of the work which the candidate claims as original
* Work submitted for other qualifications
* Work carried out prior to PhD candidature enrolment
* any third party editorial assistance, either paid or voluntary (as limited to the Editing of Research Theses by Professional Editors guidelines) and/or
* Where a substantially unchanged multi-author paper is included in the thesis a statement prepared by the candidate explaining the contributions of all involved. A signed copy by all authors must be included with the submission form.

#### Thesis preparation and binding

> Theses should normally be bound using thermal binding.

> Thermally bound theses must be used with covers strong enough to resist damage by bending or knocking. Twin-ring, spring-back and spiral binders are not acceptable, as theses bound in these ways frequently do not survive travel through the post. Thermal binding should not exceed 300 pages (absolute maximum thickness of 35mm per copy).

> If there are photographs or charts which need to be included in the thesis, facilities are available for colour laser printing and photocopying in the Graduate Student Association (GSA)and a scanner is available in the Graduate Presentation and Publishing Centre. The GSA also provides a thermal binding and photocopying service (refer to [The Graduate Centre](http://gradresearch.unimelb.edu.au/handbooks/phd/intro.html#centre)).

### [Notice of intention to submit thesis](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#intention)

> Three months prior to the submission date of the thesis, the candidate is required to submit an 80-word summary of the thesis argument.

> The 80-word summary should be submitted via the on-line [form](http://gradresearch.unimelb.edu.au/exams/candidates.html#summary).

> In addition to the 80-word summary the following information is required:
* Candidate's name
* Student number
* Department
* Supervisors' names (if required an Application for Change of Supervisor(s) form is available from my.unimelb)
* Thesis title
* Current mailing address
* Estimated thesis submission date (this is the date that you intend to submit your thesis for examination and it may not necessarily be the Expected Thesis Submission Date that appears in my.unimelb).

> Note: If more than three months lapses between submission of the 80-word summary and the thesis for examination, it is recommended that the candidate notifies the Examinations Office two weeks before the thesis is to be submitted to avoid delays with the examination. The Examinations Office can be notified by email (msgr-exams@unimelb.edu.au).

- [x] Submit intention to submit form.
- [x] Find out the latest date a thesis can be submitted in 2014.

MGSR replied:

> We will be accepting submissions up until around 11am on 24 December 2014, however, we strongly suggest submitting at least a week prior so that your thesis can be sent to examiners, if they have been nominated, prior to the Christmas break.

So, I have nominated 17/12/2014 as the date that I will submit my thesis.

### [Thesis submission](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#submission)

> Three copies of the thesis should be submitted for examination with the candidate's name clearly marked on the spine, with the family name in capitals.

> The form also includes a section where the candidate (in consultation with the supervisor) may request that the examiners maintain the confidentiality of the thesis contents. Such a request may be required where there are research contracts with outside organisations or other issues related to intellectual property.

- [ ] Check with Sue Clark and Emma Whitelaw whether sections of my thesis related to their work need to remain confidential.

### [Final form of the thesis](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#finalform)

> On completion of the examination two copies of the thesis in permanent hard cover binding, incorporating any necessary amendments or revisions, must be submitted to the Chair of Examiners to be approved. Candidates who commenced from January 2007 are also required to submit an electronic copy of the thesis to the University of Melbourne instutional repository (refer to Making your thesis available on the University of Melbourne institutional repository p. 34). Addenda of any length are not accepted. One copy must be printed on archival quality paper (paper should be approximately 80–100gsm) available from the Co-op Bookshop (in the Baillieu Library Building) or stationery suppliers. The words "Produced on archival quality paper" should be printed on the title page of this copy.

> The University logo is not permitted to be used in the thesis.

> The following must be printed on the spine of the thesis:
* Full name of the author
* Title of the thesis (abbreviated if necessary)
* Degree (ie PhD)
* Year of initial submission, or in the case of rewrites the year of resubmission.

> There are no colour restrictions for the thesis cover. The archival quality copy is for the Baillieu Library and the second copy for the Department Library.

### [Making your thesis available on the University of Melbourne institutional repository](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#eprints)

> A digital copy of a completed higher degree thesis must be submitted to the repository (via the Library).

> Candidates can elect to have the entirety of their thesis on open access, or only the citation and metadata. Candidates are required to discuss with their supervisor any Intellectual Property implications prior to lodging their thesis online.

> If the candidate agrees to allow the full text of their thesis to be made available, the necessary permissions for third party copyright material must be dealt with (refer to Copyright section of this handbook and the Copyright Office webpage).

- [ ] Select open access option (provided Sue Clark and Emma Whitelaw don't require confidentiality).
- [ ] Check rules on self-publishing.

### [Bibliography format](http://gradresearch.unimelb.edu.au/handbooks/phd/thesis.html#resources)

> [See the Citing Styles tab](http://www.unimelb.libguides.com/content.php?pid=87165&sid=3005272) for style guides and manuals, and for general referencing information.

- [ ] What style should I use for my bibliography?

### [Amendment to thesis](http://gradresearch.unimelb.edu.au/handbooks/phd/examination.html#amendment)

> Candidates are not permitted to make any changes to their thesis not specified by the examiners (or implied in their reports). Candidates are required to consult with their supervisor and the Chair of Examiners in this matter.

## 3.5 year candidature review

TODOs for 3.5 year candidature review, which is due by November 15.

- [x] Complete progress review
- [x] Organise progress review meeting __URGENT__
  - [x] Emailed Terry
  - [x] Emailed Peter
  - [x] Emailed Gordon

## Fellowship applications

TODOs for fellowship applications for 2015.

- [ ] Check out the [Joint Academic Scholarships Online Network](http://jason.edu.au/)
- [x] __URGENT:__ Consult notes from meeting with Lynne Hartley