### 2015-04-24 Rmarkdown/concluding_remarks.Rmd
Update thesis structure.

---
title: "Concluding remarks"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

\chapter{Concluding remarks}\label{chap:concluding_remarks}

__TODO: Review each chapter__