### 2015-05-02 Rmarkdown/comethylation_review.Rmd
Goknur's suggested changes.

---
author: "Peter Hickey"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
  html_document:
    keep_md: true
bibliography: ../latex/phd_thesis.bib
---

\chapter{A critical review of co-methylation}\label{chap:co-methylation_review}

\begin{chapabstract}
Co-methylation is the dependence structure of DNA methylation data. This chapter critically reviews different definitions of co-methylation and puts them in the statistical framework of Chapter \ref{chap:wgbs_statistical_framework}. I highlight various shortcomings of existing methods for quantifying co-methylation, which led me to develop the methods and software described in  Chapter \ref{chap:co-methylation}. I also review how methods to detect differential methylation have sought to account for, and exploit, co-methylation.
\end{chapabstract}

\section{What is co-methylation and why study it?}\label{subsec:co-methylation_definition}

I define co-methylation as the dependence structure of DNA methylation data. This broad definition encompasses several similar concepts previously described in the literature. These include: "the presence of methylation over a stretch of neighbouring CpG positions" \citep{Schatz:2004cha}; "the relationship between the degree of methylation over distance" \citep{Eckhardt:2006gh}; the "correlation of two [loci] across many samples" \citep{Akulenko:2013fl}; and the "_vertical_ (i.e., progenitor to descendant) correlation between the same CpG site in neighbouring cell types" \citep{Capra:2014uh}.

It is important to understand that there exists both within-sample and between-sample co-methylation.

Within-sample co-methylation is the spatial dependence of DNA methylation along the genome within a single sample. Many researchers have observed that there is a strong spatial dependence of DNA methylation along the genome, that is, methylation loci near to one another in the genome tend to be similarly methylated \citep[e.g.,][]{Eckhardt:2006gh, Cokus:2008fc, Lister:2009hy, Lacey:2009by, Li:2010fb, Lyko:2010dr, Landan:2012kp}. These studies also found that the strength of this dependence itself depends on the distance between the loci and on the genomic context.

The fundamental level of within-sample co-methylation is the dependence structure of DNA methylation events occurring on the same DNA fragment, which I call _within-fragment co-methylation_. At a higher level is the _correlation of aggregate methylation levels_, such as the spatial correlation of $\beta$-values within a sample.

Between-sample co-methylation is the relationship between methylation levels at a pair of loci across a set of samples[^concordance]. For example, \citet{Akulenko:2013fl} reported 187 pairs of genes whose methylation level was highly correlated (Pearson $|r| \geq 0.75$) across more than 300 breast cancer samples. It is worth noting that measures of within-sample co-methylation can themselves be compared between samples.

[^concordance]: One measure of between-sample correlation of aggregate methylation levels that does not fall under my definition of co-methylation is the following: Consider a pair of samples, $j$ and $j'$, and compute the correlation between the set of $\beta$-values for sample $j$ against sample $j'$, i.e., $cor(\{\beta_{i, j}\}_{i = 1}^{i = N_{loci}}, \{\beta_{i, j'}\}_{i = 1}^{i = N_{loci}})$. This measure is often reported as evidence for the "concordance" of methylation levels between replicates, but is not what I consider a form of co-methylation.

There have been several attempts to analyse co-methylation, most with the aim of better understanding the biological processes that lay down and regulate DNA methylation. An increased understanding of co-methylation could also lead to cheaper and more efficient assays of DNA methylation. For example, we might identify methylation loci whose methylation status is highly predictive of the methylation level of a surrounding region and design an assay to interrogate those highly predictive loci. This idea is analogous to the use of tag-SNPs on genotyping microarrays.

There are also reasons for studying co-methylation that are less to do with improving our biological understanding and more to do with improving statistical techniques for analysing DNA methylation data. If co-methylation can be better estimated and understood, then statistical techniques can be developed that explicitly account for and even exploit these dependencies to create more powerful and efficient analysis methods.

A common difficulty in translating biological observations into mathematical or statistical language are vague, ambiguous or otherwise insufficiently detailed descriptions of the analyses performed. This is frequently coupled with a lack of software to implement the proposed method. These deficiencies can bring into question the results and certainly hinder the development of new analysis methods and software. Many of the papers reviewed in this chapter suffer from these deficiencies and in these cases I have made my best attempt at deciphering the methods, often via correspondence with the authors.

\section{Correlations of aggregate methylation levels}
\label{sec:previous_cor_of_agg_methylation}

To date, most analyses of co-methylation have been based on analysing correlatiosn of aggregate methylation levels, such as $\beta$-values, within individual samples \citep{Eckhardt:2006gh, Cokus:2008fc, Li:2010fb, Lyko:2010dr, Lacey:2013iy}. These analyses have an implicit assumption of stationarity, that is, that the correlation between two loci depends only on the distance between them and not the actual position of the loci in the genome.

\subsection{\citet{Eckhardt:2006gh}}\label{sec:eckhardt}

\citet{Eckhardt:2006gh} is one of the first publications to apply bisulfite-sequencing to a large number of regions and samples. Of particular relevance to a discussion of co-methylation, \citet{Eckhardt:2006gh} is often cited as evidence of co-methylation. For example, \citet{Hebestreit:2013ko} cite \citet{Eckhardt:2006gh} as evidence that "methylation levels are strongly spatially correlated" and \citet{Hansen:2011gu} cite \citet{Eckhardt:2006gh} as evidence that"proximal CpGs [have] similar methylation levels".

\citet{Eckhardt:2006gh} performed PCR-amplicon bisulfite-sequencing of $2,524$ amplicons on human chromosomes $6$, $20$ and $22$. PCR-amplicon bisulfite-sequencing is a labour-intensive assay and so it was a considerable effort to produce these data. Nonetheless, it is important to bear in mind that these data are nowhere near as comprehensive as data generated using modern whole-genome bisulfite-sequencing assays. The samples are a variety of human tissue types from multiple donors of different ages, both male and female.

As part of their analysis, the authors explore co-methylation, which they define as "the relationship between the degree of methylation over distance". They state that "we were able to establish a significant correlation for comethylation over short distances ($\leq 1,000$ bp), it deteriorated rapidly for distances $>2,000$ bp" and reference the plot reprinted in Figure \ref{fig:Eckhardt} in support of this claim. In Figure \ref{fig:Eckhardt}, the authors plot the "percentage identical methylation" against "distance in bp", which ranges from $0$ to $20,000$ bp. This is not a correlation in the usual sense of the word. Nowhere in the original paper is it clearly described what the "percentage identical methylation" is nor how it is computed[^beck].

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/ng1909-F3a.jpg}
\caption[Co-methylation figure from \citet{Eckhardt:2006gh}]{Correlation of DNA methylation with spatial distance. (a) Correlation between comethylation and spatial distance. Orange dots represent CpG methylation values aggregated and averaged over $25,000$ individual measurements. Gray dots represent CpG methylation values based on resampling of random CpG positions. Blue dots indicate CpG methylation values based on resampling of amplicon positions. At distances $>1,000$ bp, we did not detect any correlation between CpG methylation and spatial distance. Adapted by permission from from Macmillan Publishers Ltd: [Nature Genetics] (\citet{Eckhardt:2006gh}), copyright (2006)}
\label{fig:Eckhardt}
\end{figure}

[^beck]: An email discussion with the senior author, Professor Stephan Beck (University College London), was unable to clarify the definition or method.

The average amplicon length is reported as $411$ bp with a standard deviation of $71$ bp. Therefore, it is impossible for the "percentage identical methylation" to have been obtained from methylation events co-occuring on the same amplicon, at least for those pairs separated by more than approximately $600$ bp.

My best guess as to the definition of "percentage identical methylation" is that it is the proportion of pairs of identical methylation calls at different CpGs from (generally) different amplicons, that is, $Pr(Z_{h, i} = Z_{h', i'})$. I will refer to my best guess of "percentage identical methylation" as $PIM^{\ast}$, and now turn to some of its properties.

Let $0 \leq p_{h} \leq 1$ be the genome-wide average CpG methylation level for the sample from which $Z_{h, i}$ is sampled, that is, $Pr(Z_{h, i} = 1) = p_{h}$. For two distinct CpGs sequenced on two (potentially) distinct amplicons, $Z_{h, i}$ and  $Z_{h', i'}$, we want to compute the probability that the methylation states are identical, that is, $Pr(Z_{h, i} = Z_{h', i'})$. Not only may the two CpGs be from distinct amplicons, but in the original analysis the amplicons themselves may be from distinct samples with different genome-wide average methylation levels. This most general setting is described by $Z_{h, i} \eqd \text{Bernoulli}(p_{h})$ and $Z_{h', i'} \eqd \text{Bernoulli}(p_{h'})$, with $cor(Z_{h, i}, Z_{h', i'}) = c$. In this setting we can compute the theoretical $PIM^{\ast}$ as follows:

\begin{align*}
PIM^{\ast} &= Pr(Z_{h, i} = Z_{h', i'}) \\
        &= Pr(Z_{h, i} = 0, Z_{h', i'} = 0) + Pr(Z_{h, i} = 1, Z_{h', i'} = 1) \\
        &= Pr(Z_{h', i'} = 0 | Z_{h, i} = 0) Pr(Z_{h, i} = 0) +
           Pr(Z_{h', i'} = 1 | Z_{h, i} = 1) Pr(Z_{h, i} = 1) \\
        &= \big [ (1 - p_{h'}) + c p_{h} \big ] (1 - p_{h}) +
           \big [ p_{h'} + c (1 - p_{h}) \big ] p_{h} \\
        &= (1 - p_{h}) (1 - p_{h'}) + p_{h} p_{h'} + 2 c p_{h} (1 - p_{h})
\end{align*}

The proof of this result is given in Appendix \ref{app:bernoulli_equal}.

In general, $PIM^{\ast} \in {[0, 1]}$ since it is a probability, which is already unlike a correlation coefficient that takes values in ${[{-1}, 1]}$. There are a few important special cases that are worth discussing:

1. $PIM^{\ast}_{inid} = (1 - p_{h}) (1 - p_{h'}) + p_{h} p_{h'}$: Independent and non-identically distributed (_inid_), i.e., $Z_{h, i} \eqd Bernoulli(p_{h})$ and $Z_{h', i'} \eqd Bernoulli(p_{h'})$ and with $cor(Z_{h, i}, Z_{h', i'}) = c = 0$.
2. $PIM^{\ast}_{did} = (1 - p)^{2} + p^{2} + 2 c p ( 1 - p)$: Dependent and identically distributed (_did_), i.e., $Z_{h, i}, Z_{h', i'} \eqd Bernoulli(p)$ and $cor(Z_{h, i}, Z_{h', i'}) = c \neq 0$.
3. $PIM^{\ast}_{iid} = (1 - p)^{2} + p^{2}$: Independent and identically distributed (_iid_), i.e., $Z_{h, i}, \allowbreak Z_{h', i'} \eqd Bernoulli(p)$ and $cor(Z_{h, i}, Z_{h', i'}) = c = 0$.

As for $PIM^{\ast}$, $PIM^{\ast}_{inid} \in [0, 1]$. In contrast, it is easy to show by differentiation that $PIM^{\ast}_{iid} \geq 0.5$ and that $PIM^{\ast}_{did} \geq 0.5 (c + 1)$, which is $> 0.5$ (resp. $< 0.5$) for $c > 0$ (resp. $c < 0$).

To return to the plot shown in Figure 3a; the "percentage identical methylation" asymptotes at $40 \%$, or $0.4$, which is less than the theoretical lower-bound of $PIM^{\ast}_{iid}$. If we believed that all samples had the same average CpG methylation levels, $p_{h}$, then we would conclude that the methylation states of two CpGs separated by a large distance are in fact __anti-correlated__ ($c < 0$) for CpGs separated by approximately $1,000$ bp. However, since we know that there is sample-to-sample variability in the average level of CpG methylation, the more likely explanation for this result is that the $p_{h}$ vary between samples and therefore $PIM^{\ast} < 0.5$ are possible, even supposing $c > 0$.

Of course, the above discussion pre-supposes that what is labelled "percentage identical methylation" in Figure \ref{fig:Eckhardt} is mathematically equivalent to $PIM^{\ast}$. I believe this to be true but due to the aforementioned inadequacies of the method description it is not possible to verify this derivation.

Quite apart from what exactly is plotted in Figure \ref{fig:Eckhardt}, the analysis is based on data from only $2,524$ amplicons. These limited data cannot give a genome-wide picture of co-methylation. Furthermore, the data come from a variety of tissues and donors, which makes it difficult to know how applicable these results are for a particular tissue. Caution is warranted in extrapolating these results to general statements about the correlation of aggregate methylation levels.

\subsection{\citet{Cokus:2008fc}}\label{sec:Cokus}

\citet{Cokus:2008fc} published the BS-seq protocol for performing whole-genome bisulfite-sequencing. They demonstrated this technique by generating a genome-wide map of DNA methylation in wildtype  _Arabidopsis thaliana_. They performed an autocorrelation analysis of the $\beta$-values for CG, CHG and CHH methylation[^arabidopsis]. The autocorrelation computes, for a set of distances ($0-5,000$ bp), the (Pearson) correlation coefficient of methylation levels for all pairs of methylation loci separated by each distance. To emphasise, I believe this analysis uses all pairs of methylation loci, that is, regardless of how many intervening methylation loci there are for each pair.

[^arabidopsis]: Many plant species, such as _A. thaliana_, have much higher levels of non-CG methylation than do mammals.

\citet{Cokus:2008fc} found that methylation levels are highly correlated, particularly for CG methylation levels, and that the correlation decays as a function of genomic distance between methylation loci. From this they concluded that the "significant correlation between methylated cytosines for distances up to 5,000 nucleotides or more [were] probably a reflection of regional foci of methylation throughout the genome and of large blocks of pericentromeric heterochromatin". They also identified correlations between different methylation contexts, e.g., CG vs. CHG, which "suggest[s] complex interactions between the different types of methylation".

The autocorrelation plots showed two periodicities, one of $\approx 167$ bp and the other of $10$ bp (CHH only). Both periodicities are visible in plots of the raw autocorrelations and are confirmed by a Fourier analysis of the autocorrelation signal.

[^CHH_only]: The $10$ bp period was later observed for CG and CHG methylation as well, at least for nucleosome bound DNA \citep{Chodavarapu:2010iqa}

The $167$ bp period "is similar to, but slightly shorter than, estimates of the average spacing of nucleosomes in plant chromatin" and Cokus et al. hypothesise that this periodicity is due to histones dictating access to DNA by the DNA methyltransferases.

The $10$ bp period is equal to the length of one helical DNA turn. From this result Cokus et al. hypothesise that the DOMAINS REARRANGED METHYLASE 2 (_DRM2_), "the main enzyme controlling asymmetric methylation in _Arabidopsis_", contains two active sites that methylate sites $8$ to $10$ bp apart[^comethylation_def2], like its mammalian homologue, DNA methyltransferase 3 (_DNMT3_). The $10$ bp CHH-only period in the autocorrelation analysis of the $\beta$-values is also observed in a _within-fragment analysis_ (discussed in \ref{sec:previous_within-fragment_analyses})

[^comethylation_def2]: This is also referred to as "co-methylation" by \citet{Jurkowska:2011dt}, here meaning the methylation of multiple loci in a single event. \citet{Jurkowska:2011dt} review a plausible biochemical model for how _DNMT3a_ could co-methylate two CpGs separated by $8$ to $10$ bp; this model includes both the scenario where both CpGs are on the same DNA strand and the scenario where the CpGs are on opposite strands.

\subsection{\citet{Lister:2009hy}}\label{sec:lister_2009}

\citet{Lister:2009hy} report an $8$ to $10$ bp period in the occurrence "methylcytosines". Recall that \citeauthor{Lister:2009hy} define a "methylcytosine" as a cytosine that has a non-zero $\beta$-value, as determined by a binomial test (see Section \ref{sec:beta-value_interpretation} for details and a criticism of this concept). With its reference to "methylcytosines", this may first appear to be a within-fragment analysis of co-methylation, but it is in fact based on $\beta$-values. However, unlike other methods described in this section, the analysis is not based on a correlation of these $\beta$-values, but rather on patterns in the spacing of "methylcytosines" along the genome. This leads to problems.

\citet{Lister:2009hy} tabulate the number of neighbouring[^non_redundant] pairs of "methylcytosines" as a function of the $IPD$. These counts are plotted as a bar chart, with $IPD$ along the x-axis and count on the y-axis, along with a cubic spline fit to the counts. This is done separately for each combination of methylation context (CG, CHG and CHH) and genomic context (exonic, intronic or random) for a total of nine plots (reproduced in Figure \ref{fig:Lister_2009_Sup_Fig_9}).

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/Lister_2009_Sup_Fig_9.pdf}
\caption[Co-methylation figure from \citet{Lister:2009hy}]{Spacing of Adjacent Methylcytosines in Different Contexts. Prevalence of mCHG/mCHH sites (y-axis) as a function of the number of bases between adjacent mCHG/mCHH sites (x-axis) based on all non redundant pair-wise distances up to $50$ nt in exons, introns and random sequences. The blue line represents smoothing by cubic splines. Adapted by permission from from Macmillan Publishers Ltd: [Nature Genetics] (\citet{Lister:2009hy}), copyright (2009)}
\label{fig:Lister_2009_Sup_Fig_9}
\end{figure}

[^non_redundant]: I think that "neighbouring" is what is meant by "non-redundant pair-wise distances" in the caption of Supplementary Figure 9, but I am not certain.

An $8$ to $10$ bp period in the fitted cubic spline is observed in most, but not all, of these nine plots. This periodicity is very weak in several contexts and is absent for exonic CpGs. The weakness of the signal is partially attributable to the small number of "methylcytosines" in the CHG and CHH contexts in human samples. More importantly, the reported $8$ to $10$ bp periods are weaker than a $3$ bp period for "methylcytosines" at exonic CpGs. Unlike the $8$ to $10$ bp period, the $3$ bp period can clearly be observed in the raw counts, although the cubic spline smoother seemingly ignores it. \citeauthor{Lister:2009hy} do not comment on the $3$ bp period.

Based on these nine plots, \citet{Lister:2009hy} draw parallels to the $10$ bp period in the autocorrelation of CHH $\beta$-values observed by \citet{Cokus:2008fc} and speculate that "_DNMT3A_  may be responsible for catalysing the methylation at non-CG sites". However, I suspect that the results in these nine plots are in fact driven by periodicities in where cytosines are located in the genome rather than any periodicities in the actual methylation of these cytosines. That is not to say that there aren't periodicities in the methylation of cytosines, but that is not what is being measured in this analysis. To explain, these plots use the raw counts of "methylcytosines", which are not adjusted for the frequency of pairs of cytosines with a given $IPD$. If pairs of cytosines are observed more often at certain $IPD$s then even if methylation was totally independent of $IPD$ we would expect more pairs of "methylcytosines" at these same $IPD$s. Therefore, at a minimum, it is necessary to normalise the number of "methylcytosines" by the number of cytosines.

In fact, there are well known periodicities in the distribution of bases in genomes. For example, in eukaryotic genomes there is a $10$ bp period of G+C dinucleotides that acts out of phase with a $10$ bp period of A+T dinucleotides for nucleosome-bound DNA \citep{Segal:2006iz} and a "well-known period-3 oscillation in coding regions" of the genome \citep{Li:1997ts}; the former likely explains the $8$ to $10$ bp period and the latter the $3$ bp period for exonic CpGs.

\subsection{\citet{Chodavarapu:2010iqa}}\label{sec:chodavarapu}

\citet{Chodavarapu:2010iqa} use MNase-seq to study nucleosome positioning and the relationship to DNA methylation, including co-methylation, in _Arabidopsis_ and an embryonic human stem cell line (_HSF1_). The authors plot a "weighted average methylation", which I presume to be a weighted version of the $\beta$-values, against the distance from predicted nucleosome start sites, which are obtained via MNase-seq. The nucleosome start sites anchor the methylation observations to a common grid and this plot is used to show that nucleosome bound DNA has a greater level of methylation than DNA that is not bound to nucleosomes. A $10$ bp period is readily observed in these plots, consistent with the discovery of $10$ bp period in the autocorrelation of $\beta$-values published by \cite{Cokus:2008fc}. Again, the observed periodicity is confirmed by analysing a Fourier transform of this correlation signal.

In addition to the genome-wide analyses, the same method is used to examine promoters, genes, repeats, euchromatic regions and centromeric regions of the _Arabidopsis_ sample and promoters, genes, repeats and CpG islands of the _HSF1_ sample, which confirmed the $10$ bp periodicity. In fact, these results revealed that the $10$ bp period is common to all methylation types, including symmetric CpG methylation, at least for nucleosome-bound DNA. This led \citeauthor{Chodavarapu:2010iqa} to discard their earlier hypothesis[^same_authors] as to the source of the $10$ bp period \citep{Cokus:2008fc} in favour of one that "nucleosomes are to some extent dictating access to the DNA and therefore setting the register of methylation __for all DNA methyltransferases__ [emphasis added]" and not just for the DRM2 methyltransferase.

[^same_authors]: Many of the same authors contributed to \citet{Cokus:2008fc} and \citet{Chodavarapu:2010iqa}.

\subsection{\citet{Li:2010fb}}\label{sec:li}

\citet{Li:2010fb} compute "the correlation of methylation level of any two nearby CpGs and the relationship between spatial distance (from one CpG to another) and strength of this correlation". \citeauthor{Li:2010fb} conclude that "co-methylation deteriorates over distance and becomes nearly undetectable at distances $>1,000$ bp", and that their results are consistent with those in \citet{Eckhardt:2006gh}. In addition to this genome-wide analysis, \citet{Li:2010fb} perform the same analysis for CpGs within 19 different genomic contexts such as upstream of a gene, untranslated regions, coding DNA sequences and various repetitive elements. Somewhat surprisingly, they do not specifically compare CpG islands to non CpG islands, although some of these differences will be captured by the analysis of regions upstream of a gene, which include many CpG island promoters.

Unfortunately, just as in \citet{Eckhardt:2006gh}, \citet{Li:2010fb} never define exactly what is being computed in their analysis of co-methylation and no software is available. All the co-methylation results are presented as supplementary figures, which lack figure legends and are supported by only very brief captions. So, once again, I have been forced to make my best guess as to what these figures show and how it was computed.

Figure \ref{fig:Li6} plots the "correlation" against "distance of CpG" ($0 - 1000$), which I believe are estimates of the correlation of aggregate methylation levels over the range $0$ to $1000$ bp. Based on the "$n = 18,936,995$" quoted in the header of panel (a) of this figure, which I interpret as the number of pairs used in this plot of "correlation" against "distance of CpG" ($0 - 1000$), I believe that what is shown are the correlations of pairs of CpG $\beta$-values where the pairs are separated by a common distance and have no intervening CpGs. If all pairs were used, $n$ would be much larger since there are more than twenty million CpGs on one strand of the haploid human genome. To emphasise, the correlation computed by \citet{Li:2010fb} is not the same as the autocorrelation computed in \citet{Cokus:2008fc} since \citet{Li:2010fb} only use pairs of adjacent CpGs whereas \citet{Cokus:2008fc} use all pairs of CpGs, regardless of the number of intervening CpGs.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/{journal.pbio.1000533.s006}.pdf}
\caption[Co-methylation by genomic feature from \citet{Li:2010fb}]{Co-methylation patterns for different genomic features. A reproduction of Supplementary Figure 6 from \citet{Li:2010fb} under the Creative Common Attribution (CCBY) licence.}
\label{fig:Li6}
\end{figure}

The plot of genome-wide correlations (panel (a) in Figure \ref{fig:Li6}) contains a period of approximately $170$ bp, very similar to the $167$ bp period identified in _Arabidopsis_ by \citet{Cokus:2008fc}. This is further evidence that nucleosome spacing effects the strength of co-methylation. However, \citet{Li:2010fb} found no evidence of higher frequency periods, such as the $10$ bp period reported by \citet{Cokus:2008fc} and \citet{Chodavarapu:2010iqa}.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/Figure_S7.pdf}
\caption[Co-methylation figure from \citet{Li:2010fb}]{Transformation of methylation correlation at nearby CpG cytosines. A reproduction of Supplementary Figure 7 from \citet{Li:2010fb} under the Creative Common Attribution (CCBY) licence.}
\label{fig:Li7}
\end{figure}

\subsection{\cite{Lyko:2010dr}}\label{sec:lyko}

In their study of differential methylation in queen and worker honey bees (_Apis mellifera_), \citet{Lyko:2010dr} perform an autocorrelation analysis of CpG methylation. They report that the "correlation of methylation status of neighbouring CpGs increases sharply between 1 bp and 20 bp, then drops rapidly between 40 bp and 100 bp, and then slowly fades away". However, this analysis is fundamentally flawed.

Figure \ref{fig:Lyko} reproduces the results for the co-methylation analysis. Note that the maximum value of the "CpG autocorrelation" is less than $0.015$ (panel (A) of Figure \ref{fig:Lyko}). I strongly suspect that the autocorrelation has been incorrectly computed. Specifically, what I believe has been computed is the autocorrelation of all positions in the genome, not just CpGs, where non-CpGs have been artificially assigned a $\beta$-value of zero. It does not make sense to include non-CpG positions, particularly non-cytosines that can never be methylated, in an analysis of CpG co-methylation. I do not have the data used in \citet{Lyko:2010dr}, but the following example supports my conclusion. This example uses the chromosome 22 data from the _FF_ sample in the _Lister_ dataset (see Section \ref{sec:Lister_dataset} for full details of this sample).

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/Lyko_Figure_S7.pdf}
\caption[Co-methylation figure from \citet{Lyko:2010dr}]{Periodicity of methylation patterns. (A) Autocorrelation of CpG methylation status over 1 kb. (B) Autocorrelation over 100 bp. Figures A and B show that the correlation of methylation status of neighboring CpGs increases sharply between 1 bp and 20 bp, then drops rapidly between 40 bp and 100 bp, and then slowly fades away. CpGs within a neighborhood of 2 bp to 100 bp are thus more likely to share the same methylation status than more distant CpGs. (C) Fourrier transform of autocorrelation showing a clear periodicity peak at 33 cycles per 100 bp (every 3 bp). (D) Distribution of codon position of mCs, and distribution of methylation level depending on the position. These two panels indicate that the distance between methylated CpGs is often a multiple of three and that the methylated cytosine corresponds most frequently to the first nucleotide of an arginine codon. A reproduction of Supplementary Figure 7 from \citet{Lyko:2010dr} under the Creative Common Attribution (CCBY) licence.}
\label{fig:Lyko}
\end{figure}

Let $\bm{\beta^{\ast}}$ be the augmented vector of $\beta$-values, where:

\begin{equation*}
\beta^{\ast}_{i'} = \left\{
  \begin{array}{ll}
  0 & \text{if } i' \text{ is not a CpG}\\
  \beta_{i} & \text{if } i' \text{ is a CpG}
  \end{array}
  \right.
\end{equation*}

Figure \ref{fig:FF_chr21_acf_beta_star} is a plot of the autocorrelation of the augmented vector, $\bm{\beta^{\ast}}$ for the chromosome 22 data from the _FF_ sample. We see that it is remarkably similar to that panel (a) of Figure \ref{fig:Lyko}; the correlation first increases then decreases and the maximum correlation is very small, close to zero.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/FF_chr21_acf_beta_star.pdf}
\caption[Autocorrelation of incorrectly defined $\beta$-values]{Autocorrelation of $\bm{\beta^{\ast}}$ for the chromosome 21 data of the \emph{FF} sample.}
\label{fig:FF_chr21_acf_beta_star}
\end{figure}

Furthermore, if we compute the Fourier transform of this autocorrelation we find a period of approximately 3 bp, identical to that reported by \citet{Lyko:2010dr} for honey bees (compare panel (c) of Figure \ref{fig:Lyko} and Figure \ref{fig:FF_chr21_acf_beta_star_fourier}). Like my earlier criticism of the co-methylation results of \citet{Lister:2009hy}, I believe the apparent $3$ bp period is more likely due to the spacing of bases in the genome than anything to do with their methylation state.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/FF_chr21_acf_beta_star_fourier.pdf}
\caption[Fourier transformation of incorrectly defined $\beta$-values]{Fourier transformation of the autocorrelation of $\bm{\beta^{\ast}}$ for the chromosome 21 data of the \textit{FF} sample. Fourier transformation computed and plotted using the \texttt{spectrum} function in R. The period is defined as 1 / frequency. The maximum frequency is approximately $0.33$, which corresponds to a period of approximately $3$ bp.}
\label{fig:FF_chr21_acf_beta_star_fourier}
\end{figure}

All of this casts serious doubt on the validity and interpretation of the co-methylation results in \citet{Lyko:2010dr}.

\subsection{\citet{Lacey:2013iy}}\label{sec:lacey_2013}

\citet{Lacey:2013iy} explore co-methylation in their paper on modelling and analysing reduced representation bisulfite-sequencing data (RRBS). To do so, \citeauthor{Lacey:2013iy} fit a Gaussian variogram model to the CpG $\beta$-values from chromosome 11 of a normal myotube cell line (_MTCTL2_). The empirical and fitted variograms asymptote for CpGs separated by approximately $3000$ bp (Figure 1d of \citet{Lacey:2013iy}, unable to reproduce here due to rights restrictions) and so they conclude that the methylation level of CpGs "show a strong correlation for sites in close proximity, decaying to near independence at distances beyond 3000 bp".

Caution is warranted in extrapolating these results to statements about genome-wide co-methylation. Firstly, this analysis uses only approximately $60,000$ CpGs[^ncpgs], far fewer than the $25$ million CpGs in the haploid human genome. Secondly, these data are from a single sample from a single tissue, which cannot capture the sample-to-sample nor tissue-to-tissue variability in co-methylation. Finally, RRBS targets CpG dense regions of the genome with low levels of methylation and so these data do not capture co-methylation outside of these regions, where the majority of the CpGs are found and which are highly methylated.

[^ncpgs]: The data for MTCTL2 are not publicly available so I estimated the number of sequenced CpGs on chromosome 11 from three comparable RRBS libraries (see Appendix \ref{app:lacey}).

\section{Within-fragment co-methylation}
\label{sec:previous_within-fragment_analyses}

As discussed in Section \ref{sec:m-tuples}, bisulfite-sequencing allows for the study of methylation patterns at m-tuples from individual DNA fragments. Since each read comes from a single haplotype from a single cell[^chimera], these data can be used to study co-methylation at the level of individual DNA molecules. I argue that this is the more biologically meaningful level at which to study co-methylation, since this is the same scale as the physical process that lays down and maintains DNA methylation.

[^chimera]: Here I ignore the possibility of chimeric reads that may be artificially produced by DNA fragments ligating to one another during the library preparation or sequencing process.

I call this _within-fragment co-methylation_, where 'fragment' refers to a fragment of DNA. More precisely, within-fragment co-methylation is the dependence of DNA methylation at methylation loci that occur on the same DNA fragment. It can be thought of as the dependence structure of the binary stochastic process $\mathbf{Z_{h}} = (Z_{h, 1}, Z_{h, 2}, \ldots, Z_{h, N_{loci}})$. There is less previous research studying within-fragment co-methylation than there is studying correlations of aggregate methylation levels. In Chapter \ref{chap:co-methylation} I extensively explore within-fragment co-methylation and how different variables affect this dependence, such as the distance between methylation loci and the genomic context of the methylation loci.

If we could sequence entire chromosomes by single reads, then we could analyse co-methylation between any set of methylation loci. However, the most commonly used technology for high-throughput bisulfite-sequencing, which is based on Illumina's sequencing technology, can only generate reads that span approximately $200$ to $250$ bp. We are therefore limited to studying within-fragment co-methylation between methylation loci separated by at most $200$ to $250$ bp[^assembly]. Long-read sequencing technologies (e.g., Pacific Biosciences, Roche 454 and Oxford Nanopore) would overcome this limitation but the lower throughput of these technologies, and the reliability with which they can identify methylcytosines, create new complications. An alternative solution, although completely hypothetical, is to use a technique that generates synthetic long reads, similar to those generated by Illumina's TruSeq Synthetic Long-Read technique[^moleculo].

[^assembly]: NB Assembling these reads into longer contigs will not solve this problem since the assembly does not guarantee that the reads assembled into longer fragments actually originated from the same cell or haplotype.

[^moleculo]: My understanding is that Illumina's TruSeq Synthetic Long-Read technique cannot be directly applied to bisulfite-sequencing for a number of technical reasons, including the fragility of large DNA fragments treated with sodium bisulfite.

To summarise, current technologies mean that we are limited to studying within-fragment co-methylation at loci that are within regions of approximately $200-250$ bp.

\subsection{\citet{Cokus:2008fc}}\label{sec:cokus_within_fragment}

As discussed in Section \ref{sec:Cokus}, \citet{Cokus:2008fc} identify a $10$ bp period in the autocorrelation of $\beta$-values for CHH methylation from a whole-genome bisulfite-sequencing experiment of _Arabidopsis thaliana_. They also report that they find this same period "when individual reads are examined directly".

The sequencing technology at the time of this publication produced very short reads, on average only $31$ bp of usable sequence. For the within-fragment co-methylation results, \citeauthor{Cokus:2008fc} first identify all reads containing multiple CHH loci. Then, for a pair of CHH loci in the same read, they estimate the probability that the second CHH in the pair is methylated given that the first CHH in the pair is methylated. This is estimated by simply counting the number of times this occurs and dividing by the number of pairs of CHH loci that occur within the same read[^cokus_clarification]. They estimate this probability separately for each distance between CHH loci and separately for each of the five _Arabidopsis_ chromosomes. \citeauthor{Cokus:2008fc} then plot the average across the five chromosomes as a function of the number of bp between the CHH loci. The plot includes pointwise $95\%$ confidence intervals, computed under an assumption of normality, and a running average of these averages. This plot, reproduced in Figure \ref{fig:Cokus_2008_Sup_Fig_14}, shows evidence of an approximately $10$ bp period, supporting the interpretation of the results observed in the autocorrelation of the $\beta$-values.

[^cokus_clarification]: A read may contain multiple pairs of CHH loci and it is not clear whether \citeauthor{Cokus:2008fc} consider all pairs of CHH loci or only the first pair of CHH loci in a read. The number of pairs of CHH loci that occur within the same read is not the same as the number of reads containing multiple CHH loci. Either method should give similar results since there will be few $31$ bp reads containing multiple pairs of CHH loci.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{../figures/Cokus_2008_Sup_Fig_14.pdf}
\caption[Co-methylation figure from \citet{Cokus:2008fc}]{Within-read probability of additional methylation of CHH sites within a given distance from a methylated CHH site. Data were derived from individual BS-Seq reads. The x-axis indicates the distance between the two cytosines. The y-axis indicates the probability of methylation of CHH sites within the given distance from another methylated CHH site. Each point is the mean value from averaging the probability from each of the five Arabidopsis chromosomes, and the blue line is a running average of these mean values. Error bars represent $95\%$ confidence intervals via critical values of Student $t$ distributions. Adapted by permission from from Macmillan Publishers Ltd: [Nature Genetics] (\citet{Cokus:2008fc}), copyright (2008)}
\label{fig:Cokus_2008_Sup_Fig_14}
\end{figure}

I consider these results as based on a restricted definition of within-fragment co-methylation. To explain, these results tell us nothing about the probability that both CHH loci are _unmethylated_ nor the joint probability distribution of methylation for two CHH loci on the same fragment; they tell us only about the probability that both CHH loci are methylated.

Specifically, \citet{Cokus:2008fc} only estimate the marginal probability that a CHH is methylated, $Pr(Z_{h, i} = 1)$, and the probability that two CHH loci on the same fragment are both methylated, $Pr(Z_{h, i'} = 1 | Z_{h, i} = 1)$, where $i < i'$. From these two quantities we can only estimate two of the four joint probabilities, $Pr(Z_{h, i} = x, Z_{h, i} = y), x, y \in \{0, 1\}$, namely, $Pr(Z_{h, i} = 1, Z_{h, i'} = 1)$ and $Pr(Z_{h, i} = 1, Z_{h, i'} = 0)$. It does not allow estimation of $Pr(Z_{h, i} = 0, Z_{h, i'} = 0)$, nor of $Pr(Z_{h, i} = 0, Z_{h, i'} = 1)$, since these require estimates of $Pr(Z_{h, i'} = 0 | Z_{h, i} = 0)$. While it is trivial to additionally estimate $Pr(Z_{h, i'} = 0 | Z_{h, i} = 0)$, I argue in Section \ref{subsec:wf_cometh} that it is more useful and simpler to estimate the odds ratio, $\frac{Pr(Z_{h, i} = 1, Z_{h, i'} = 1) Pr(Z_{h, i} = 0, Z_{h, i'} = 0)}{Pr(Z_{h, i} = 1, Z_{h, i'} = 0) Pr(Z_{h, i} = 0, Z_{h, i'} = 1)}$, rather than the conditional probabilities, $Pr(Z_{h, i'} = 1 | Z_{h, i} = 1)$ and $Pr(Z_{h, i'} = 0 | Z_{h, i} = 0)$.

Furthermore, by aggregating all pairs of CHH loci separated by a common distance and then estimating these probabilities, there is an implicit assumption that all such pairs on the same chromosome have the same conditional probability, $Pr(Z_{h, i'} = 1 | Z_{h, i} = 1)$. There is good reason to suspect that this is not true and that there exists significant pair-to-pair variability. For example, we know for CpGs that those within CpG islands have a vastly different probability of being methylated than do those outside of islands; it is reasonable to suspect that this is also true for CHH methylation in different regions of the _Arabidopsis_ genome and hence for the co-methylation of these same loci. This aggregation may also introduction artificial associations into the $2 \times 2$ contingency table from which these probabilities are estimated \citep{Good:1987ce}.

\subsection{\citet{Ball:2009kh}}

\citet{Ball:2009kh} use bisulfite padlock probes with $36$ bp single-end reads to profile CpG methylation at approximately $7,000$ CpGs in an Epstein-Barr virus transformed B-lymphocyte cell line. As part of their study, \citeauthor{Ball:2009kh} sought to investigate whether the co-methylation reported by \citet{Eckhardt:2006gh} occurs at the single-molecule level.

To do so, \citeauthor{Ball:2009kh} took all reads containing multiple CpGs and mapping to positions with intermediate $\beta$-values ($0.2 < \beta < 0.8$) with at least $100\times$ sequencing coverage. From these reads they computed the correlation of within-read methylation states. \citeauthor{Ball:2009kh} found that the distribution of these correlations is centred around $0.5$ with very few values below zero, evidence that methylation levels at CpGs on the same DNA fragment "are generally positively correlated".

The obvious limitation to this analysis is the small number of CpG pairs interrogated by the assay. The short reads also only allow for the exploration of within-fragment co-methylation at very close CpGs and \citeauthor{Ball:2009kh} do not explore the effect of distance between CpGs ($IPD$) on co-methylation.

\subsection{\citet{Lacey:2009by}}\label{sec:lacey_2009}

\citet{Lacey:2009by} develop stochastic models of DNA methylation replication during mitosis (cell division) for a single double-stranded DNA molecule in humans. The most complex model, called the _neighbouring sites model_, uses a one-step Markov random field. Under this model, the methylation state of one locus, $Z_{h, i}$, depends on the methylation states of the adjacent loci, $Z_{h, i - 1}$ and $Z_{h, i + 1}$. The physical distances between these neighbours, the $IPD$s, are not considered.

The neighbouring sites model requires an analysis of within-fragment co-methylation in order to estimate the transition probabilities. They estimate these parameters from two small regions of 10 human ovarian carcinoma samples and a pool of somatic controls, each region approximately $200$ bp in length and containing 13 and 14 CpGs, respectively. These regions were sequenced using hairpin-bisulfite PCR, an assay that measures DNA methylation on both strands of a double-stranded DNA molecule[^hairpin_bs]. For each region, $12$ to $16$ clones were sequenced per sample[^clones]

[^hairpin_bs]: This is unlike traditional bisulfite-sequencing, which only measures methylation along a single DNA strand.

[^clones]: Essentially, $12-16$ reads were generated per sample.

For each of these regions, \citeauthor{Lacey:2009by} found that the methylation level of a CpG strongly depends on that of its neighbours. In other words, they found significant within-fragment co-methylation. They also found that the overall level of methylation of the CpGs in a region influenced the strength of the within-fragment co-methylation.

In summary, this analysis uses high resolution measurements from a moderate number of clones from a moderate number of samples for a minuscule fraction of the human genome. While the methods and models are quite interesting, the data are so limited that the results obtained from these analyses cannot represent a genome-wide picture of within-fragment co-methylation.

\subsection{\citet{Landan:2012kp}}\label{sec:landan}

\citet{Landan:2012kp} is notable for its use of the within-read information available from bisulfite-sequencing data. The authors explore the within-sample heterogeneity of methylation patterns at 4-tuples and 6-tuples of CpGs from ultra-deep bisulfite-sequencing ($> 10,000 \times$ median sequencing coverage of each CpG in the target) of 45 cancer-related CpG islands and study how methylation dynamics at this regions evolves over time. \citeauthor{Landan:2012kp} also use these data to estimate co-methylation.

As part of their analysis, \citeauthor{Landan:2012kp} construct "methylation linkage diagrams" based on a Pearson correlation estimate of within-fragment co-methylation. For each CpG 2-tuple in their targeted regions, they create a $2 \times 2$ contingency table. The general form of the table is shown below in Table \ref{tab:landan}. They then compute the Pearson correlation of each table[^cor_equivalence], $r = \frac{n_{mm} \times n_{uu} - n_{mu} \times n_{um}}{\sqrt{n_{u+} \times n_{m+} \times n_{+u} \times n_{+m}}}$. Importantly, $r$ is based on within-read measurements from a single sample and so these $r$ are estimates of within-read, within-sample co-methylation.

[^cor_equivalence]: The formula I give for $r$ is also known as the $\phi$ coefficient of a $2 \times 2$ table. The $\phi$ coefficient of a $2 \times 2$ table of binary variables, such as Table \ref{tab:landan}, is mathematically equivalent to the Pearson correlation coefficient of the same data.

\begin{table}[h]
\centering
\caption[Counts of reads at a CpG 2-tuple]{Summary of read counts at a CpG 2-tuple. For example, $n_{um}$ is the number of reads methylated at the first CpG and unmethylated at the second CpG. We use the 'plus' notation to denote the sum over the index it replaces, e.g., $n_{u+} = n_{uu} + n_{um}$.}
\label{tab:landan}
\begin{tabular}{@{}cccc@{}}
\cmidrule(l){2-4}
 & \multicolumn{3}{c}{Second CpG} \\
 \cmidrule(l){2-4}
\multirow{3}{*}{First CpG} & & & \\
 & $n_{uu}$       & $n_{um}$       & $n_{u+}$ \\
 & $n_{mu}$       & $n_{mm}$       & $n_{m+}$ \\ \cmidrule(l){2-4}
 & $n_{+u}$  & $\cdot n_{+m}$  & $n_{++} = n$      \\
\end{tabular}
\end{table}

So-called "methylation linkage diagrams" are created by plotting the Pearson correlation of each 2-tuple, $r_{(i, i')}$ as a heatmap (_à la_ heatmaps of linkage disequilibrium) to identify blocks of CpGs with significant co-methylation. It should be noted that the concepts of 'subject' and 'population' are rather different for "methylation linkage" than they are for genetic linkage (see Table \ref{tab:linkage}).

\begin{table}[h]
\centering
\caption[Notions of `subject' and `population' in methylation linkage]{Notions of `subject' and `population' in traditional genetic linkage and the "methylation linkage" of \citet{Landan:2012kp}. NB `person' could be any organism, e.g, `mouse', `dog', etc.}
\label{tab:linkage}
\begin{tabular}{@{}lll@{}}
\toprule
                    & Subject               & Population \\ \midrule
Genetic linkage     & Person                   & Population \\
Methylation linkage & (Haploid) genome of cell & Person     \\ \bottomrule
\end{tabular}
\end{table}

From the methylation linkage analysis of these 45 regions, \citeauthor{Landan:2012kp} conclude that:

> "[the] correlation between the methylation states of pairs of CpGs was generally very low. This lack of correlation suggests that methylation dynamics are typically independent for different CpGs, making the methylation state of one CpG (whether high or low) uninformative on the methylation state of nearby CpGs."

As we will see in Chapter \ref{chap:co-methylation}, this result is contradicted by my analyses of within-fragment co-methylation using whole-genome data.

\section{Between-sample co-methylation}
\label{sec:between_sample_cometh}

As is now abundantly clear, there are multiple definitions of within-sample co-methylation. An almost orthogonal concept is between-sample co-methylation, the correlation of methylation levels for a single pair of loci between a set of samples.

Since it is a loci-specific, population-level correlation, between-sample co-methylation is somewhat analogous to the idea of linkage disequilibrium between genetic loci. Briefly, for two loci, $(i, i')$, the correlation of aggregate methylation levels is computed across samples using, for example, the Pearson correlation coefficient, $r = cor(\bm{\beta_{i}}, \bm{\beta_{i'}})$, where $\bm{\beta_{i}} = (\beta_{i, 1}, \ldots, \beta_{i, N_{samples}})$.

The concept of between-sample co-methylation leads to the attractive idea of identifying "methylation tag-loci", whose methylation level is a good proxy for the methylation level across a larger region. These tag-loci could act like tag-SNPs used in genome-wide association studies, which would allow researchers to assay a small subset of loci in the methylome and capture much of the methylation variability. This idea has also been explored by \citet{Barrera:2012cq} with respect to the use of _HpaII_ sites as proxies for the methylation level of CpG islands.

Unfortunately, unlike linkage disequilibrium, we cannot assume that between-sample co-methylation is identical between different tissues from the same individuals since DNA methylation, unlike DNA sequence, varies between tissues. Therefore, at best, we might hope to generate tissue-specific maps of between-sample co-methylation.

In order to exploit between-sample co-methylation in this way, we also need to know the size of these co-methylation blocks, that is, the length of the regions over which between-sample co-methylation is sufficiently strong. The longer these blocks are, the more sparsely we can afford to sample individual methylomes.

\citet{Bell:2011fm} identify evidence of between-sample co-methylation using the Illumina 27k microarray to measure DNA methylation in lymphoblastoid cell lines from 77 HapMap Yoruba individuals. Of note, they found that between-sample co-methylation decays as a function of the distance between CpGs ($IPD$) and is stronger within CpG islands than outside of CpG islands. However, it must be recognised that the Illumina 27k microarray measures DNA methylation at less than $1\%$ of all CpGs as is biased towards CpG islands.

Another example of between-sample co-methylation is given by \citet{Liu:dy}. The authors compute the correlation of pairs of $\beta$-values between samples in three studies ($n = 247, n = 91$ and $n= 305$, respectively) to identify clusters of correlated CpGs and relate methylation changes at these clusters to nearby SNP genotypes. The resolution of their data is better than \citet{Bell:2011fm} since they use the Illumina 450k microarray, however it still only captures less than $2\%$ of all CpGs in the genome. \citeauthor{Liu:dy} conclude that "DNA methylation is correlated over regions with a median length of 274 bp from our data set". They also note that, "methylation codependency [co-methylation] itself varies across the genome", as is to be expected.

A simple way to estimate between-sample co-methylation is to create a matrix of $\beta$-values, $\bm{\beta}$, where the $(i, j)$ entry is $\beta_{i, j}$. Then, compute $R = cor(\bm{\beta})$, whose $(k, l)$ entry is $cor(\bm{\beta_{., k}}, \bm{\beta_{., l}})$, where $\bm{\beta_{., k}}$ is the column vector of $\beta$-values for the $k^{th}$ methylation loci. This is what the R package, `coMET`   ([http://bioconductor.org/packages/devel/bioc/html/coMET.html](http://bioconductor.org/packages/devel/bioc/html/coMET.html)), does to estimate the between-sample co-methylation of Illumina HumanMethylation450 array data. As far as I am aware, `coMET` is the only software specifically designed to estimate between-sample co-methylation and integrate this with other genomic annotations.

\section{Exploiting co-methylation in downstream analyses}
\label{sec:exploiting_cometh}

The variety of definitions, rigour, and reproducibility of the published analyses of co-methylation, not to mention the lack of software implementing proposed methods, has made it difficult to properly account for co-methylatoin in downstream analyses. Ideally, co-methylation would not only be _accounted for_ in downstream analyses but _exploited_ to leverage the power of the correlation structure of DNA methylation data. In fact, some authors have already sought to do this.

For example, the R/Bioconductor packages `bsseq` \citep{Hansen:2011gu, Hansen:2012gr} and `BiSeq` \citep{Hebestreit:2013ko} both cite \citet{Eckhardt:2006gh} as providing evidence of co-methylation, which can be exploited by spatial smoothing of the $\beta$-values prior to testing for differential methylation. Specifically, by assuming that CpGs close to one another have similar methylation levels, the smoothed $\beta$-values are, on average, more accurate than the raw $\beta$-values. Both `bsseq` and `BiSeq` smooth $\beta$-values on a per-sample basis.

Smoothing is particularly powerful for experiments with low sequencing coverage since the error in estimating the raw $\beta$-values is inversely proportional to the sequencing coverage and smoothing reduces this error. An important parameter when smoothing is the choice of window size over which to smooth, also known as the _bandwidth_ of the smoother. The default window size in `bsseq` is defined as one that contains at least 70 CpGs and is at least 2000 bp wide. `BiSeq` uses a much smaller default window size of 80 bp, a decision that is  driven by its focus on RRBS data which has a higher CpG density than whole-genome bisulfite-sequencing data. Ideally, a data-driven bandwidth would be used, one driven by the strength of co-methylation in the region, but this is difficult to implement in practice. A simpler but potentially useful hybrid would be to define different window sizes for genomic features with generally different co-methylation structures, e.g., CpG islands vs. the rest of the genome (see Chapter \ref{chap:co-methylation}).

In addition to smoothing the $\beta$-values, `BiSeq` also estimates and tries to account for the spatial correlation of test statistics when testing for differential methylation at neighbouring loci. This is not the same as looking for correlations between the $\beta$-values, but is rather based on looking for correlations amongst the _differences_ in $\beta$-values, since these differences are what the test statistics of differential methylation are based upon. Specifically, `BiSeq` fits a semivariogram to the test statistics of differentially methylated CpGs. The correlations estimated using the semivariogram are in turn used in estimates of the standard deviations when identifying differentially methylated regions.

More generally, methods such as `SLIM` \citep{Wang:2011cw} and `comb-p` \citep{Pedersen:2012vl} attempt to account for the correlation amongst test statistics in genomics data. They do this by analysing the distribution of P-values rather than the distribution of the test statistics themselves or the underlying data. This makes these methods quite general but application-specific methods may be more powerful. `comb-p` has been used in the analysis of DNA methylation data from the Illumina microarray platforms \citep{Pedersen:2012vl}.

These methods exploiting correlations amongst the test statistics or their P-values are not directly based on co-methylation estimates. However, what drives the correlations amongst the test statistics is a complex mixture of within-sample and between-sample co-methylation.

`Aclust` \citep{Sofer:2013bk} is notable for its use of between-sample co-methylation to identify differentially methylated regions. Unlike other methods for detecting DMRs, which test for DMCs and then cluster these DMCs into DMRs, `Aclust` clusters loci prior to testing for differential methylation. That is, it only tests for DMRs and not DMCs. This has the advantage of reducing the dimensionality of the data. Clusters are formed using a modified form of agglomerative nested clustering, where the modification restricts clusters to comprise neighbouring loci. The distance metric used in the clustering is based on an estimate of the between-sample co-methylation for each pair of adjacent loci.

\citet{Capra:2014hx} also make use of between-sample co-methylation, albeit a specialised form of it, to study the dynamics of DNA methylation in differentiating cell lineages. For each locus there is a correlation between its methylation state in the precursor cell and its dependent cell types, which they call "vertical" correlations (in contrast to the "horizontal" correlations of within-sample co-methylation). In this specialised setting, where 'between-samples' means "between samples in a lineage", they show that a method based on "vertical" correlations is better at imputing the methylation level of CpGs than a method based on "horizontal" within-sample co-methylation.

\section{Summary}

While it has long been observed that DNA methylation at nearby loci is highly correlated, pinning down exactly what this means, and its implications, is surprisingly difficult. As I have shown, there are problems and limitations with most, if not all, previous analyses. These problems range from a lack of methodological detail and limited data, through to serious flaws in the methodology or its implementation. Moreover, despite numerous papers including an analysis of co-methylation, none of these have includes software to reproduce their method, further complicating the assessment of these methods.

It is clearly inappropriate to treat methylation at nearby loci as independent and methods have been proposed to account for or exploit this in various ways.  These methods would be better informed by a more rigorous and reproducible analysis of co-methylation, which I give in Chapter \ref{chap:co-methylation}.